### 2020-10-12 presentation/presentation.tex
slide: ~model

\ifcase 0  % choose 0=slides, 1=article, 2=refart
   \documentclass[aspectratio=169,ignorenonframetext,9pt]{beamer}
\or\documentclass[a4paper,11pt]{article}
   \usepackage{url,beamerarticle}
\or\documentclass[a4paper,11pt]{refart}
   \let\example\relax
   \usepackage{url,beamerarticle}
\fi

\ifcase 0  % choose a theme like these
    % \usetheme{boxes}
    \usetheme{Boadilla}
    % \usetheme{Goettingen}% I recommend
    % \usetheme{Singapore}
    % \usetheme{Pittsburgh}
    % \usetheme{Madrid}
    % \usetheme{Warsaw} % common choice, but often poor
\fi

\usepackage{graphicx,pgfplots,parskip}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,url,array}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]

\title{BSTERGM: Bayesian Separable-Temporal Exponential family Graph Models}
\author{Choi Seokjun}
\date{16 Oct. 2020}

\begin{document}

\begin{frame}
\maketitle
\end{frame}

% \begin{abstract}
% This abstract, being outside the frame environment, does not appear in the presentation.  Your outline will be the basis for a couple of sentences of talk for each of the following questions:
% \begin{itemize}
% \item What was done?
% \item Why do it?
% \item What were the results?
% \item What do the results mean in theory and/or practice?
% \item What is the reader's benefit?
% \item How can the readers use this information for themselves? 
% \end{itemize}
% \end{abstract}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{definitions and terminologies}
\begin{frame}{Random graphs}
    \begin{defn}[Random graphs and related terminologies]
    Let $\mathcal{B}=\{0,1\}$. For given $n\in \mathbb{N}$,
    \begin{itemize}
        \item The set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes (without weights for each nodes and edges.)
        \item Let $\Omega$ be an event set. We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
        \item For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$, \\
            satisfying $Y_{ij}=1$ if the edge is connected. Otherwise, $Y_{ij}=0$.
        \item If edges of $Y$ have directions, then $Y$ is called a directed graph. Otherwise, $Y$ is called a undirected graph.
    \end{itemize}
    \end{defn}
    Let me notate a realization of random graph by $y$ and its edges by $y_{ij}$ for $i,j=1,2,...,n$.

    Here is a remark. These are obvious that, for given $n\in \mathbb{N}$,
    \begin{itemize}
        \item $|\mathcal{Y}|=2^{n(n-1)}$ if $\mathcal{Y}$ is the set of all directed graphs not permitting self-connecting edges.
        \item $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
    \end{itemize}
    ;thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
\end{frame}

\begin{frame}{ERGM: Exponential family Random Graphs Models}
    \begin{defn}[ERGM: Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes and $Y$ be a random graph (on $\mathcal{Y}$.)
        Set a distribution on $\mathcal{Y}$ to
        \[P(Y=y;\theta) = \frac{exp(\theta^{T}s(y))}{c(\theta)}\]
        for some $\theta\in\mathbb{R}^p$,
        where $s(y)\in\mathbb{R}^p$ is a vector which is part of y's sufficient network statistics,
        \\ and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
        \\ Models which have a such form called the ERGM.
    \end{defn}

    \begin{defn}[TERGM: Temporal Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) to
        \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta) = \frac{exp(\theta^{T}s(y_t, y_{t-1}))}{c(\theta, y_{t-1})}\]
        for $\theta\in\mathbb{R}^p$ and with the first-order Markov assumption $P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})$,
        \\where $s(y_t, y_{t-1})\in\mathbb{R}^n$ is a part of sufficient network statistics,
        and $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$ is a normalizing constant.
        Models which have a such form called the TERGM.
    \end{defn}
\end{frame}

\begin{frame}{STERGM: Separable-Temporal Exponential family Random Graphs Models}
    \begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) by following way:

        Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
        Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
        Next, to $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
        \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})},
        P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
        for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
        and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1})), c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.

        Then, defining operations $+,-$ on $\mathcal{Y}$ following the boolean algebra edgewise-ly, set $y_t$ to
        \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]

        Additionally, assume that
        \begin{itemize}
            \item The first-order Markov assumption: $P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})$
            \item The separability: the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$;
                thus, \(P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+)P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\)
        \end{itemize}
        Models which have a such form called the SERGM.
    \end{defn}
\end{frame}
\end{document}

### 2020-10-12 presentation/presentation.tex
tex: ~fin

@@
 
+\usepackage{algorithm,algorithmicx}
 \usepackage{graphicx,pgfplots,parskip}
@@
 \newtheorem{exmp}{Example}[section]
+\newtheorem{algo}{Algorithm}[section]
 
@@
 
-\section{definitions and terminologies}
+\section{Model definitions and related terminologies}
 \begin{frame}{Random graphs}
@@
         Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
-        Next, to $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
+        Next, for $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
         \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})},
@@
         \end{itemize}
-        Models which have a such form called the SERGM.
+        Models which have a such form called the STERGM.
     \end{defn}
 \end{frame}
+
+\begin{frame}{BSTERGM: Bayesian STERGM}
+    To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
+    Then, the posterior of $\theta^+,\theta^-$ becomes
+    \[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
+    where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.
+
+    Remarks:
+    \begin{itemize}
+        \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
+        \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
+            Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Network Statistics for BSTERGM}
+    As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
+    A common candidates of $s'$ are:
+    \begin{itemize}
+        \item the number of edges
+        \item node degree distribution (each order)
+        \item edgewise shared partner distribution (each order)
+        \item dyadwise shared partner distribution (each order)
+        \item k-star distribution
+        \item triangle distribution
+    \end{itemize}
+    In fact, a bundle of all order of node degree distribution and all order of edgewise shared partner distribution form sufficient statistics of a graph.
+    Also, we can set $s'$ by a function of these statistics.
+\end{frame}
+
+\section{Model fitting algorithm}
+\begin{frame}{Fitting the BSTERGM}
+    Suppose that a sequence of graph samples $y_1,...,y_T$ is observed
+    \\ (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined,) 
+    \\ and we want to fit the observation using BSTERGM.
+
+    How do we find the posterior distribution of $\theta^+,\theta^-$?
+
+    Note that an ordinary MCMC algorithm does not work in our situation, because the constant part remains when calculating the MCMC ratio.
+    To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
+\end{frame}
+
+\begin{frame}{Fitting the BSTERGM}
+    Here is the algorithm of our main chain for the exchange algorithm.
+    \begin{algo}[the main chain]
+    Let $y_1,...,y_T$ be given. For $m=1,...,M$,
+    \begin{enumerate}
+        \item Propose candidates $\theta_*^+,\theta_*^-$ from $\epsilon(.|\theta_{m-1}^+,\theta_{m-1}^-)$.
+        \item Select a lag $(t-1,t)$ randomly on $2 \leq t \leq T$.
+        \item Generate an exchange graph $y_{ex} \in\mathcal{Y}|_t$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
+        \item Calculate the exchange MCMC ratio $\pi$ at the lag,
+            \[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+,\theta_*^-)}
+                {P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+,\theta_{m-1}^-)}
+                \frac{P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}{P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
+        \item With probability $min(\pi,1)$, accept the proposal and put $(\theta_m^+,\theta_m^-) = (\theta_*^+,\theta_*^-)$.\\
+            Otherwise, reject the proposal and put $(\theta_m^+,\theta_m^-) = (\theta_{m-1}^+,\theta_{m-1}^-)$.
+    \end{enumerate}
+    \end{algo}    
+    Observe that, the $\pi$ has no normalizing constant terms because they are canceled out by added exchange terms.
+    \(log\pi = (\theta_*^+-\theta_{m-1}^+)(s(y_{t}^+,y_{t-1})-s(y_{ex}^+,y_{t-1}))
+    +(\theta_*^- -\theta_{m-1}^-)(s(y_{t}^-,y_{t-1})-s(y_{ex}^-,y_{t-1}))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
+    \( = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
+    +(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
+\end{frame}
+
+\begin{frame}{Fitting the BSTERGM}
+    We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
+    we should have a generative algorithm at a given parameter points.
+    However, we still do not know the normalizing constant. Thus, I use one more MCMC chain.
+    \begin{algo}[the auxiliary chain]
+    Let $\theta^+,\theta^-,y_{t-1}$ be given. For $k=1,...,K$,
+    \begin{enumerate}
+        \item Select one edge randomly from the $y_{k-1}$, say, $y_{k-1;ij}$.
+        \item Propose a new graph $y_*$ with switching the $y_{ij}$ value from $y_{k-1}$: 
+        \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$(dissolution case.) Oterwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$(formation case.)
+        \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
+        \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
+        \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= \frac{exp(\theta^T s(y_*,y_{t-1}))}{exp(\theta^T s(y_{k-1},y_{t-1}))} = exp(\theta^T (s'(y*)-s'(y_{k-1})))\]
+        \item With probability $min(\phi,1)$, accept the proposal and put $y_k=y_*$.\\
+            Otherwise, reject the proposal and put $y_k=y_{k-1}$.
+    \end{enumerate}
+    \end{algo}
+    After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
+\end{frame}
+
+\begin{frame}{MCMC Diagnosis}
+    Since we run two kinds of MCMC chains, we should proceed two diagnostic task.
+
+    The main chain produces the posterior samples of parameter, so the procedure is same as ordinary MCMC case.
+    \begin{itemize}
+        \item Cut burn-in period. Do thinning if it is needed.
+        \item Depict traceplots of each parameter chain to check the convergence and the mixing.
+        \item Depict autocorrelation plot. Calculate ESS if it is needed.
+    \end{itemize}
+
+    Next, checking all auxiliary chains is practically irritating one. 
+    In general, it is suffice to check the auxiliary chain of the last iteration (of the main chain)
+    with statistics included in the model.
+    \begin{itemize}
+        \item Calculate network statistics of all graphs produced by the last auxiliary chain.
+        \item Depict traceplots of each statistics to check the convergence and the mixing.
+    \end{itemize}
+\end{frame}
+
+\section{Model inference}
+\begin{frame}{Inference and Prediction}
+    Since we have posterior samples by running the main chain as many as we want, we can do basic inference procedure.
+    \begin{itemize}
+        \item A outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
+        \item An approximated summary statistics: mean, mode, variance, ...
+        \item An approximated quantile and probability interval
+    \end{itemize}
+
+    Moreover, we already have a generative algorithm at the specific parameter point,
+    we can predict the form of network at $T+1,T+2,...$ using posterior sample following standard Bayesian method.
+    For example, for predicting $T+1$,
+    \begin{itemize}
+        \item Run K iteration using auxiliary chain at $y_T$ with each posterior sample points.
+        \item Take the each network as a predicted result.
+    \end{itemize}
+    If you need, calculate some network statistics for the results and get a summary statistics of them.
+\end{frame}
+
+\begin{frame}{Goodness of Fit}
+    To evaluate the goodness of fit for the posterior $\theta^+,\theta^-$,
+    we can use the auxiliary chain algorithm once again.
+    \begin{algo}[GOF procedure]
+        For t=2,...,T
+        \\For s=1,...,S
+        \begin{enumerate}
+            \item sample $\theta_s^+,\theta_s^-$ from the estimate of posterior.
+            \item simulate $y_s$ using the auxiliary chain under $y_{t-1}$.
+            \item calculate $g(y_s)$, some higher degree statistics (eg. Node-degree dist \& Edgewise Shared Partner dist)
+        \end{enumerate}
+        Draw the box-plot of $g(y_s)$ and compare with $g(y_t)$.
+    \end{algo}
+\end{frame}
+
+\begin{frame}{Supplements}
+    You can find the C++ implementation (using Armadilo: see http://arma.sourceforge.net/)
+    \\ of BSTERGM fitting, diagnostic, and GOF algorithms
+    \\ at my Github page: https://github.com/letsjdosth/BayesianSTERGM.
+\end{frame}
+
 \end{document}

### 2020-10-12 BayesianSTERGM/BSTERGM_MCMC.h
BSTERGM_MCMC : log r modification (exclude a calculation of obsSeq.netstat

@@ -279,20 +279,20 @@ private:
         // now model
         // n_Edge, k_starDist(2)
         model_delta_exchangeFormation_1 += {
-            (double)exchange_Formation.get_n_Edge() - observedSeq[lag_startTimePoint].get_n_Edge(),
-                (double)exchange_Formation.get_k_starDist(2) - observedSeq[lag_startTimePoint].get_k_starDist(2)
+            (double)exchange_Formation.get_n_Edge(), // - observedSeq[lag_startTimePoint].get_n_Edge(),
+                (double)exchange_Formation.get_k_starDist(2) // - observedSeq[lag_startTimePoint].get_k_starDist(2)
         };
         model_delta_exchangeDissolution_2 += {
-            (double)exchange_Dissolution.get_n_Edge() - observedSeq[lag_startTimePoint].get_n_Edge(),
-                (double)exchange_Dissolution.get_k_starDist(2) - observedSeq[lag_startTimePoint].get_k_starDist(2)
+            (double)exchange_Dissolution.get_n_Edge(), // - observedSeq[lag_startTimePoint].get_n_Edge(),
+                (double)exchange_Dissolution.get_k_starDist(2) // - observedSeq[lag_startTimePoint].get_k_starDist(2)
         };
         model_delta_obsFormation_3 += {
-            (double)observed_FormationSeq[lag_startTimePoint+1].get_n_Edge() - observedSeq[lag_startTimePoint].get_n_Edge(),
-                (double)observed_FormationSeq[lag_startTimePoint+1].get_k_starDist(2) - observedSeq[lag_startTimePoint].get_k_starDist(2)
+            (double)observed_FormationSeq[lag_startTimePoint+1].get_n_Edge(), // - observedSeq[lag_startTimePoint].get_n_Edge(),
+                (double)observed_FormationSeq[lag_startTimePoint+1].get_k_starDist(2) // - observedSeq[lag_startTimePoint].get_k_starDist(2)
         };
         model_delta_obsDissolution_4 += {
-            (double)observed_DissolutionSeq[lag_startTimePoint+1].get_n_Edge() - observedSeq[lag_startTimePoint].get_n_Edge(),
-                (double)observed_DissolutionSeq[lag_startTimePoint+1].get_k_starDist(2) - observedSeq[lag_startTimePoint].get_k_starDist(2)
[…]

### 2020-10-26 presentation_simple/presentation_simple.tex
add data and simple version tex

--- 2020-10-12 presentation/presentation.tex
@@
 
-\section{Model definitions and related terminologies}
-\begin{frame}{Random graphs}
-    \begin{defn}[Random graphs and related terminologies]
-    Let $\mathcal{B}=\{0,1\}$. For given $n\in \mathbb{N}$,
-    \begin{itemize}
-        \item The set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes (without weights for each nodes and edges.)
-        \item Let $\Omega$ be an event set. We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
-        \item For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$, \\
-            satisfying $Y_{ij}=1$ if the edge is connected. Otherwise, $Y_{ij}=0$.
-        \item If edges of $Y$ have directions, then $Y$ is called a directed graph. Otherwise, $Y$ is called a undirected graph.
-    \end{itemize}
-    \end{defn}
-    Let me notate a realization of random graph by $y$ and its edges by $y_{ij}$ for $i,j=1,2,...,n$.
-
-    Here is a remark. These are obvious that, for given $n\in \mathbb{N}$,
-    \begin{itemize}
-        \item $|\mathcal{Y}|=2^{n(n-1)}$ if $\mathcal{Y}$ is the set of all directed graphs not permitting self-connecting edges.
-        \item $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
-    \end{itemize}
-    ;thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
-\end{frame}
-
+\section{Model definition}
 \begin{frame}{ERGM: Exponential family Random Graphs Models}
     \begin{defn}[ERGM: Exponential family Random Graphs Models]
-        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes and $Y$ be a random graph (on $\mathcal{Y}$.)
+        Let $\mathcal{Y} \subset \{0,1\}^{n^2}$ be a set of graphs with $n$ nodes and $Y$ be a random graph (on $\mathcal{Y}$.)
+
         Set a distribution on $\mathcal{Y}$ to
@@
         \\ and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
-        \\ Models which have a such form called the ERGM.
-    \end{defn}
-
-    \begin{defn}[TERGM: Temporal Exponential family Random Graphs Models]
-        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
-        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) to
-        \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta) = \frac{exp(\theta^{T}s(y_t, y_{t-1}))}{c(\theta, y_{t-1})}\]
-        for $\theta\in\mathbb{R}^p$ and with the first-order Markov assumption $P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})$,
-        \\where $s(y_t, y_{t-1})\in\mathbb{R}^n$ is a part of sufficient network statistics,
-        and $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$ is a normalizing constant.
-        Models which have a such form called the TERGM.
+        \\ Models which have a such form are called the ERGM.
     \end{defn}
@@
     \begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
-        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
-        Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) by following way:
+        Let $Y_{t-1}=y_{t-1} \in \mathcal{Y}$ be given and $Y_t \in \mathcal{Y}$.
+        Let $\mathcal{Y}^+ \subset \mathcal{Y}$ be a set of all graphs having equal or additional edges comparing to $y_{t-1}$.
+Likewise, define $\mathcal{Y}^-\subset\mathcal{Y}$ having equal or sparser edges.
+        For $Y_t^+ \in \mathcal{Y}^+$, $Y_t^- \in \mathcal{Y}^-$, set
 
-        Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
-        Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
-        Next, for $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
-        \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})},
-        P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
-        for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
+        Formation model:
+        \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+        Dissolution model:
+        \[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
+        for $\theta^+,\theta^-\in\mathbb{R}^p$, where $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$ are (functions of) sufficient network statistics,
         and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1})), c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
-        Then, defining operations $+,-$ on $\mathcal{Y}$ following the boolean algebra edgewise-ly, set $y_t$ to
+        Then, defining operations $+,-$ on $\mathcal{Y}$ by the boolean algebra edgewisely, set $y_t$ to
         \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
@@
         \end{itemize}
-        Models which have a such form called the STERGM.
+        Models which have a such form are called the STERGM.
     \end{defn}
@@
 \begin{frame}{BSTERGM: Bayesian STERGM}
-    To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
+    To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes rule as the inference rule.
     Then, the posterior of $\theta^+,\theta^-$ becomes
@@
     As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
-    A common candidates of $s'$ are:
+    For example, some candidates of $s'$ are:
     \begin{itemize}
         \item the number of edges
-        \item node degree distribution (each order)
-        \item edgewise shared partner distribution (each order)
-        \item dyadwise shared partner distribution (each order)
-        \item k-star distribution
-        \item triangle distribution
+        \item k-stars
+        \item triangles
     \end{itemize}
-    In fact, a bundle of all order of node degree distribution and all order of edgewise shared partner distribution form sufficient statistics of a graph.
-    Also, we can set $s'$ by a function of these statistics.
+    Note that we can set $s'$ by a function of sufficient network statistics.
 \end{frame}
@@
     How do we find the posterior distribution of $\theta^+,\theta^-$?
-
-    Note that an ordinary MCMC algorithm does not work in our situation, because the constant part remains when calculating the MCMC ratio.
-    To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
 \end{frame}
@@
 \begin{frame}{Fitting the BSTERGM}
-    Here is the algorithm of our main chain for the exchange algorithm.
     \begin{algo}[the main chain]
@@
         \item Select a lag $(t-1,t)$ randomly on $2 \leq t \leq T$.
-        \item Generate an exchange graph $y_{ex} \in\mathcal{Y}|_t$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
+        \item Generate an exchange graph $y_{ex} \in\mathcal{Y}$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
         \item Calculate the exchange MCMC ratio $\pi$ at the lag,
@@
     \end{algo}    
-    Observe that, the $\pi$ has no normalizing constant terms because they are canceled out by added exchange terms.
-    \(log\pi = (\theta_*^+-\theta_{m-1}^+)(s(y_{t}^+,y_{t-1})-s(y_{ex}^+,y_{t-1}))
-    +(\theta_*^- -\theta_{m-1}^-)(s(y_{t}^-,y_{t-1})-s(y_{ex}^-,y_{t-1}))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
-    \( = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
+    Observe that the $\pi$ has no normalizing constant terms because they are canceled out by added exchange terms.
+    \(log\pi = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
     +(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
+
+    Moreover, it is noteworthy that for the second step, an auxiliary MCMC chain which generate an exchange sample is used.
 \end{frame}
 
-\begin{frame}{Fitting the BSTERGM}
-    We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
-    we should have a generative algorithm at a given parameter points.
-    However, we still do not know the normalizing constant. Thus, I use one more MCMC chain.
-    \begin{algo}[the auxiliary chain]
-    Let $\theta^+,\theta^-,y_{t-1}$ be given. For $k=1,...,K$,
-    \begin{enumerate}
-        \item Select one edge randomly from the $y_{k-1}$, say, $y_{k-1;ij}$.
-        \item Propose a new graph $y_*$ with switching the $y_{ij}$ value from $y_{k-1}$: 
-        \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$(dissolution case.) Oterwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$(formation case.)
-        \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
-        \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
-        \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= \frac{exp(\theta^T s(y_*,y_{t-1}))}{exp(\theta^T s(y_{k-1},y_{t-1}))} = exp(\theta^T (s'(y*)-s'(y_{k-1})))\]
-        \item With probability $min(\phi,1)$, accept the proposal and put $y_k=y_*$.\\
-            Otherwise, reject the proposal and put $y_k=y_{k-1}$.
-    \end{enumerate}
-    \end{algo}
-    After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
-\end{frame}
+% \begin{frame}{Fitting the BSTERGM}
+%     We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
+%     we should have a generative algorithm at a given parameter points.
+%     However, we still do not know the normalizing constant. Thus, I use one more MCMC chain.
+%     \begin{algo}[the auxiliary chain]
+%     Let $\theta^+,\theta^-,y_{t-1}$ be given. For $k=1,...,K$,
+%     \begin{enumerate}
+%         \item Select one edge randomly from the $y_{k-1}$, say, $y_{k-1;ij}$.
+%         \item Propose a new graph $y_*$ with switching the $y_{ij}$ value from $y_{k-1}$: 
+%         \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$(dissolution case.) Oterwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$(formation case.)
+%         \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
+%         \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
+%         \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= \frac{exp(\theta^T s(y_*,y_{t-1}))}{exp(\theta^T s(y_{k-1},y_{t-1}))} = exp(\theta^T (s'(y*)-s'(y_{k-1})))\]
+%         \item With probability $min(\phi,1)$, accept the proposal and put $y_k=y_*$.\\
+%             Otherwise, reject the proposal and put $y_k=y_{k-1}$.
+%     \end{enumerate}
+%     \end{algo}
+%     After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
+% \end{frame}
 
@@
 \begin{frame}{Inference and Prediction}
-    Since we have posterior samples by running the main chain as many as we want, we can do basic inference procedure.
+    Since we have posterior samples by running the main chain as many as we want, we can follow the common inference procedure for the Bayes models.
     \begin{itemize}
-        \item A outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
+        \item An outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
         \item An approximated summary statistics: mean, mode, variance, ...
@@
 \begin{frame}{Goodness of Fit}
-    To evaluate the goodness of fit for the posterior $\theta^+,\theta^-$,
+    To evaluate the goodness of fit for the posterior of $\theta^+,\theta^-$,
     we can use the auxiliary chain algorithm once again.
@@
         \end{enumerate}
-        Draw the box-plot of $g(y_s)$ and compare with $g(y_t)$.
+        Draw a box-plot of $g(y_s)$s and compare with $g(y_t)$.
     \end{algo}

### 2020-12-06 doc/doc.tex
BSTERGM_onlyedge fit / conv varrate set

--- 2020-10-12 presentation/presentation.tex
@@
-\ifcase 0  % choose 0=slides, 1=article, 2=refart
+\ifcase 1  % choose 0=slides, 1=article, 2=refart
    \documentclass[aspectratio=169,ignorenonframetext,9pt]{beamer}
@@
 \usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,url,array}
+\usepackage{cite}
 
@@
 \title{BSTERGM: Bayesian Separable-Temporal Exponential family Graph Models}
-\author{Choi Seokjun}
-\date{16 Oct. 2020}
+\author{Choi, Seokjun}
+\date{14 Dec. 2020}
 
@@
 
-% \begin{abstract}
-% This abstract, being outside the frame environment, does not appear in the presentation.  Your outline will be the basis for a couple of sentences of talk for each of the following questions:
-% \begin{itemize}
-% \item What was done?
-% \item Why do it?
-% \item What were the results?
-% \item What do the results mean in theory and/or practice?
-% \item What is the reader's benefit?
-% \item How can the readers use this information for themselves? 
-% \end{itemize}
-% \end{abstract}
+\begin{abstract}
+Separable Temporal Exponential Random Graph Models (STERGM) are useful models to describe changes in dynamic networks' structures evolving over time. 
+The Bayesian-inference based approach for the model, however, had not been developed yet. 
+In this paper, I propose the Bayesian STERGM (BSTERGM): STERGM with the Bayesian setting, and I illustrate the computational algorithm that generates posterior samples from the model dealing with the doubly-intractable constant problem.
+Furthermore, I show the interpretability of the model in the Bayesian context.
 
-\begin{frame}{Outline}
-\tableofcontents
-\end{frame}
+\end{abstract}
 
-\section{Model definitions and related terminologies}
-\begin{frame}{Random graphs}
-    \begin{defn}[Random graphs and related terminologies]
-    Let $\mathcal{B}=\{0,1\}$. For given $n\in \mathbb{N}$,
-    \begin{itemize}
-        \item The set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes (without weights for each nodes and edges.)
-        \item Let $\Omega$ be an event set. We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
-        \item For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$, \\
+% \begin{frame}{Outline}
+% \tableofcontents
+% \end{frame}
+
+\section{Introduction}
+
+\section{Background}
+\subsection{Random graphs}
+For statistical analysis on relations between subjects of data,
+I will introduce terminologies and notations for denoting components related to networks, 
+which are a representation of these type of data.
+
+We say each subject on the data to a node. They are same as a vertex in graph theory.
+In addition, we say a tie joining two nodes to an edge. A tie can be directed or not.
+These two components have fundamental role describing the data.
+Define a graph or a network to $y=(V,E)$, where $V$ is a set of node and $E$ is a set of edges of a network.
+
+For simplicity, I will deal with a situation without weight on each node or edge.
+Then, a graph $y$ can be represented by simpler form.
+Assume that the number of node $n\in \mathbb{N}$ on data is given.
+Then, for $\mathcal{B}=\{0,1\}$,
+the set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes.
+Now, a graph $y \in \mathcal{Y}$ can be represented by a $n \times n$ binary matrix
+with elements $y_{ij}=1$ if there is a tie from $i$-th node to $j$-th node,
+otherwise $y_{ij}=0$.
+
+If edges of $y$ have directions, then $y$ is called a directed graph. Otherwise, $y$ is called a undirected graph.
+These are obvious that, for given $n\in \mathbb{N}$,
+$|\mathcal{Y}|=2^{n(n-1)}$ if $\mathcal{Y}$ is the set of all directed graphs not permitting self-connecting edges,
+and $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
+Thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
+
+I extend the concept of graphs to be able to describe a situation that connections among nodes are random.
+Let $\Omega$ be an event set of all possible connected situations among given nodes.
+We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
+For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$
 satisfying $Y_{ij}=1$ if the edge is connected. Otherwise, $Y_{ij}=0$.
-        \item If edges of $Y$ have directions, then $Y$ is called a directed graph. Otherwise, $Y$ is called a undirected graph.
-    \end{itemize}
-    \end{defn}
+Note that the $Y_{ij}$s are also random variables.
+
 Let me notate a realization of random graph by $y$ and its edges by $y_{ij}$ for $i,j=1,2,...,n$.
 
-    Here is a remark. These are obvious that, for given $n\in \mathbb{N}$,
-    \begin{itemize}
-        \item $|\mathcal{Y}|=2^{n(n-1)}$ if $\mathcal{Y}$ is the set of all directed graphs not permitting self-connecting edges.
-        \item $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
-    \end{itemize}
-    ;thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
-\end{frame}
+\subsection{Exponential random graph model: ERGM}
+From the definition of random graph, one of the natural questions is 
+how we can model a distribution on $\mathcal{Y}$.
+As a answer, the Exponential random graph models (ERGM) (Wasserman and Pattison(1996) and Robins et al.(2007)) %ref
+gives a classical but 
+flexible scheme to model for time-invariant networks represented as a single directed or undirected graph.
+Moreover, ERGM provide a statistical approach to the systematic exploration of several local features simultaneously and 
+how these local structures interact to form a global network.
 
-\begin{frame}{ERGM: Exponential family Random Graphs Models}
-    \begin{defn}[ERGM: Exponential family Random Graphs Models]
-        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes and $Y$ be a random graph (on $\mathcal{Y}$.)
+Specifically, following the notation of previous section,
+let $\mathcal{Y}$ be a set of random graphs with $n$ nodes and $Y \in \mathcal{Y}$ be a random graph.
 Set a distribution on $\mathcal{Y}$ to
@@
 where $s(y)\in\mathbb{R}^p$ is a vector which is part of y's sufficient network statistics,
-        \\ and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
-        \\ Models which have a such form called the ERGM.
-    \end{defn}
+and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
+Models which have such a form called the ERGM.
 
-    \begin{defn}[TERGM: Temporal Exponential family Random Graphs Models]
-        Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-        Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
+Note that the form of the right-hand side is the canonical form of exponential family distributions.
+That is why the model is called exponential. Moreover, observe that the normalizing constant cannot be directly calculated
+because the size of $\mathcal{Y}$ grows exponentially as $n$ grows.
+This situation is said to doubly-intractable constant problem, and it causes serious challenge to estimate the parameter vector $\theta$
+in real data analysis. For more detail on fitting algorithm using MCMCMLE, see Hunter and Handcock(2006).% ref
+
+\subsection{BERGM: Bayesian ERGM}
+Caimo and Friel(2012) %ref
+extend ERGM to bayesian setting. Putting prior $p(\theta)$ to the parameter $\theta$, the model becomes
+\[p(\theta|y)=\frac{p(y|\theta)p(\theta)}{p(y)}=\frac{exp(\theta^T s(y))}{c(\theta)}\frac{p(\theta)}{p(y)}\]
+Since not only the doubly-intractable constant problem still exists but also the $p(y)$ cannot be evaluated directly,
+the estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
+The solution of Caimo and Friel is using the exchange algorithm.
+
+%exchange algorithm 설명
+
+\subsection{TERGM: Temporal Exponential family Random Graphs Models}
+Let's consider a sequence of random graph $\{Y_1, Y_2, ..., Y_T\} (T\in\mathbb{N})$ and assume that
+we are interested in the dynamics or evolution of the network sequence.
+The temporal Temporal Exponential family Random Graphs Models(TERGM)
+proposed by Hanneke, Fu and Xing(2009)%ref
+is a model for our interest if we permit a Markov assumption
+\[P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})\]
+and focus on the transition probability along index of the sequence.
+For convenience, I refer the index as a time.
+
+Specifically, let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
+Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs in $\mathcal{Y}$.
 Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) to
 \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta) = \frac{exp(\theta^{T}s(y_t, y_{t-1}))}{c(\theta, y_{t-1})}\]
-        for $\theta\in\mathbb{R}^p$ and with the first-order Markov assumption $P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})$,
-        \\where $s(y_t, y_{t-1})\in\mathbb{R}^n$ is a part of sufficient network statistics,
-        and $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$ is a normalizing constant.
+where $t=2,...,T$, $\theta\in\mathbb{R}^p$,
+sufficient network statistics $s(y_t, y_{t-1})\in\mathbb{R}^n$,
+and a normalizing constant $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$.
 Models which have a such form called the TERGM.
-    \end{defn}
-\end{frame}
 
-\begin{frame}{STERGM: Separable-Temporal Exponential family Random Graphs Models}
+The estimation of $\theta$ can be performed by MCMCMLE,
+but the estimation suffers severe degeneracy problem.
+Furthermore, When we interpret the estimated TERGM parameters,
+we cannot separate the generation and duration because
+the higher value of $\theta$ increases both.
+These raise demand for modified models to remedy the degeneracy and enable separable interpretation.
+
+\subsection{STERGM: Separable-Temporal Exponential family Random Graphs Models}
 \begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
@@
     Next, for $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
-        \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})},
-        P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
+    \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+    \[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
     for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
@@
 \end{defn}
-\end{frame}
 
-\begin{frame}{BSTERGM: Bayesian STERGM}
-    To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
-    Then, the posterior of $\theta^+,\theta^-$ becomes
-    \[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
-    where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.
-
-    Remarks:
-    \begin{itemize}
-        \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
-        \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
-            Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Network Statistics for BSTERGM}
+\subsection{Network Statistics for BSTERGM}
 As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
@@
 Also, we can set $s'$ by a function of these statistics.
-\end{frame}
 
-\section{Model fitting algorithm}
-\begin{frame}{Fitting the BSTERGM}
+\section{Bayesian STERGM}
+
+To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
+Then, the posterior of $\theta^+,\theta^-$ becomes
+\[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
+where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.
+
+Remarks:
+\begin{itemize}
+    \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
+    \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
+        Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
+\end{itemize}
+
+\section{Bayesian Estimation}
+
 Suppose that a sequence of graph samples $y_1,...,y_T$ is observed
@@
 To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
-\end{frame}
 
-\begin{frame}{Fitting the BSTERGM}
 Here is the algorithm of our main chain for the exchange algorithm.
@@
 +(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
-\end{frame}
 
-\begin{frame}{Fitting the BSTERGM}
 We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
@@
 After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
-\end{frame}
 
-\begin{frame}{MCMC Diagnosis}
+\section{Diagnostics, inference, and Bayesian GOF}
+
+\subsection{Model diagnostics}
 Since we run two kinds of MCMC chains, we should proceed two diagnostic task.
@@
 \end{itemize}
-\end{frame}
 
-\section{Model inference}
-\begin{frame}{Inference and Prediction}
+\subsection{Model inference}
+
 Since we have posterior samples by running the main chain as many as we want, we can do basic inference procedure.
@@
 If you need, calculate some network statistics for the results and get a summary statistics of them.
-\end{frame}
 
-\begin{frame}{Goodness of Fit}
+\subsection{Goodness of Fit}
 To evaluate the goodness of fit for the posterior $\theta^+,\theta^-$,
@@
 \end{algo}
-\end{frame}
+
+\section{Example: Scholars and faculty members' experiences
+in online social networks}
+\section{Summary and Discussion}
 
@@
 
+testcite \cite{RN100}
+
+\nocite{*}
+\bibliographystyle{plain}
+\bibliography{doc.bib}
+
 \end{document}

### 2020-12-06 paper_format/CSY.tex
BSTERGM_onlyedge fit / conv varrate set

\documentclass[a4paper, 11pt]{report}
\usepackage{times,epsfig,calc,subfigure}
\usepackage[authoryear,round]{natbib}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
%\usepackage[tablesfirst, nolists]{endfloat}
\usepackage{amsbsy, epsfig, epsf, psfrag, booktabs, graphicx, amssymb, enumerate}
\usepackage[hidelinks,bookmarks=false]{hyperref}
\usepackage{lscape}
\usepackage{bm}
\usepackage{float}
\floatplacement{table}{!htbp}
\usepackage{adjustbox}
\usepackage{color, colortbl}

\definecolor{Gray}{gray}{0.9}

\usepackage{makecell}
\usepackage{changepage}%adjust
\usepackage{setspace}%stretch
\usepackage{booktabs} 
\usepackage{multirow}
%\usepackage{hypernat}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{multicol}
\usepackage{epstopdf}
\usepackage{tabularx, caption}
\usepackage[flushleft]{threeparttable}
\usepackage{grad}

\usepackage{kotex}

\usepackage{placeins}

\makeatletter
\newcommand{\distas}[1]{\mathbin{\overset{#1}{\kern\z@\sim}}}%
\newsavebox{\mybox}\newsavebox{\mysim}
\newcommand{\distras}[1]{%
	\savebox{\mybox}{\hbox{\kern3pt$\scriptstyle#1$\kern3pt}}%
	\savebox{\mysim}{\hbox{$\sim$}}%
	\mathbin{\overset{#1}{\kern\z@\resizebox{\wd\mybox}{\ht\mysim}{$\sim$}}}%
}
\makeatother

\renewcommand{\baselinestretch}{1}
\linespread{1.4}
\renewcommand{\arraystretch}{1.3}
\newcommand{\cov}{\mathop{\rm cov}\nolimits} %covariance
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defn}[thm]{Definition}
\newcommand\defeq{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
\DeclareMathOperator*{\argmin}{argmin}

%%%%%%%%%%%%%%%%%%%%%%%
%   TITLE
%%%%%%%%%%%%%%%%%%%%%%%

\title{Longitudinal Latent Space Item Response Model with the Applications to U.S. Congressional Votes}
\author{Su Young Choi}
\date{Dec 2020}

\begin{document}	
\maketitle \pagestyle{plain}

\newpage
\thispagestyle{empty}

  \null
  \begin{center}
    \vskip -1.4cm
    {\fontsize{14pt}{14pt}\selectfont This certifies that Master's thesis of Su Young Choi is approved.}
    \vskip 4cm%
    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Thesis Supervisor: Prof. Ick Hoon Jin\\
    \vskip 2cm
    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Committee Member: Prof. Jongho Im\\
    \vskip 2cm
    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Committee Member: Prof. Jaewoo Park\\
    \vskip 3cm%
    {\fontsize{14pt}{14pt}\selectfont {The Graduate School \vskip 1.0em Yonsei University \vskip 1.0em Dec 2020}}%
  \end{center}
  \par
\clearpage

\pagestyle{plain}

\clearpage \pagenumbering{roman} \tableofcontents \clearpage

\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
	\hspace*{-0.29in} List of Figures}
\listoffigures

\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
    \hspace*{-0.29in} List of Tables}
\listoftables

\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
    \hspace*{-0.29in} Abstracts}

%%%%%%%%%%%%
% ABSTRACT %
%%%%%%%%%%%%

\begin{abstract}
\begin{center}
{\LARGE Longitudinal Latent Space Item Response Model  with the Applications to U.S. Congressional Votes}
\end{center}
\indent
\vskip 0.8cm

\noindent

\vskip 0.2cm

\vspace{\stretch{1}} \noindent
\hrulefill\\
{\bf Key words : }
\parbox[t]{0.8\textwidth}
{}
\end{abstract}

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\clearpage
\chapter{Introduction} \label{Intro}
\pagenumbering{arabic}

The main idea of this paper is applying LSRM to political data and showing the advantage of the model in aspect of visualization, interpretation, and usability. As for congressional votes data, it is suitable for LSRM because it consists of respondents (politician) and items (bill). Politicians' voting results are expressed as binary data.

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

\newcommand{\indep}{\raisebox{0.05em}{\rotatebox[origin=c]{90}{$\models$}}}
%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Data Description} \label{Chapter2}
\section{U.S. Congressional Votes Data} \label{data}

We chose the votes data of the US house congress for our model. The data is collected from congressional votes database of govtrack.us. We selected the votes of the first quarter in 2016, 2018, and 2020. The reason of 2 years interval is that the election of the house of representatives are held in every even-numbered year. Since the data of each time points are made by new house members, it is meaningful to analysis this time series data. Among the bills, we deducted the bills which are passed or rejected by the overwhelming majority, over 95\% of the all seats, because it might induce biased results. Calculating the proportion of the vote result, we transformed the raw data to binary data consisting of 1 and 0. Considering that the vote of politicians are composed of agreement(yes), rejection(no), and abstention(Not voting), it is natural to set an agreement as 1, and, a rejection and an abstention as 0. Although an abstention is not a strong declaration of own intention, it can be interpreted that the agreement is not acceptable to them. As for each vote, the number of participants differs because of absent members. Then, identifying the seat number should be preceded. After this process, absents of representatives and vacant seats are detected and we regard it as a missing value.

\quad Investigating politicians and congressional district, we found that Pennsylvania was under redistricting, and decided to remove Pennsylvania districts from the data. Redistricting is the process of redrawing electoral district boundaries, and because of this act, the congressional districts of Pennsylvania are not the same at three-time points. In consequence, objects of analysis are not identical in Pennsylvania, and it might distort the result. Thus, it is suitable to preclude the 18 districts of Pennsylvania from the analysis.

\quad The number of politicians is 417 at all time points, and it is $n$, the number of respondents, in our model. The number of votes is 128, 114, and 95 at each 2016, 2018, and 2020. These are $p$, the number of items, in our model. Finally, we obtain 417 by 128, 417 by 114, 417 by 95 binary data matrices.

\section{Data Collecting Method} \label{crawling}
We used Selenium package of Python for crawling congressional data. Because there is one CSV file for one bill, we had to collect about 300 files. To make it work automatically, we built the crawling code with Selenium. Thanks to the crawling tool, the collecting process only took about 30 minutes. After that, we cleansed the collected data and integrated into one CSV file for each year using R program.
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
\chapter{Background} \label{Cahpter3}
\section{Item Response Theory}

\section{Latent Space Model}

\section{Latent Space Item Response Model}\label{network}

Longitudinal latent space item response model is the time series expansion of Latent space Rasch model. Ahead of formulating the longitudinal version, we must define the standard LSRM. Latent space rasch model is based on the distance model of network analysis, in which the probability of a tie depends on the Euclidean distance between latent positions of a respondent and an item. There is the two types of nodes, the respondents and items, in the network of Rasch model while the original distance model only has one kind of node, actors.
%% Notation
\begin{itemize}
\item Notation
\begin{itemize}
    \item ${\bf X}_t$: A binary item response dataset at time $t$.
    \item $\boldsymbol\beta_t$ and $\boldsymbol\theta_t$: Item and respondent intercept parameters at time $t$, respectively.
    \item $\gamma_t$: A scale parameter for the respondent and item latent positions.
    \item ${\bf Z}_{t,n \times d}$: Latent positions for respondents at time $t$.
    \item ${\bf W}_{t,p \times d}$: Latent positions for items at time $t$.
\end{itemize}
%% Rasch model
We formulate the latent space Rasch model with respondent and item attribute effects as follows:
\begin{equation}
\label{lsrm}
    \mbox{logit}\Big(X_{t,ki} = 1 \mid \beta_{t,i}, \theta_{t,k}, \gamma_t, {\bf z}_{t,k}, {\bf w}_{t,i}\Big)
    = \beta_{t,i} + \theta_{t,k} - \gamma_t \Big|\Big|{\bf z}_{t,k} - {\bf w}_{t,i}\Big|\Big|
\end{equation}
where $k$ and $i$ are indexes for respondents and items, respectively. $\theta_{t,k}$ implies the willingness of the respondent to make a response as 1 for the item. $\beta_{t,i}$ means the easiness of the item to be responded as 1 by the respondent. We treats $\theta_{t,k}$ as a random effect but $\beta_{t,i}$ as a fixed effect, following the convention of item response modeling. For this model, we specify prior distributions for the model parameters as follows:
\[\begin{split}
    \beta_{t,i} \mid \tau_{\beta}^2 &\sim \mbox{N}\big(0, \tau_{\beta}^2\big), \quad \tau_{\beta}^2 > 0 \\
    \theta_{t,k} \mid \sigma_t^2 &\sim \mbox{N}\big(0, \sigma_t^2\big), \quad \sigma_t^2 > 0 \\
    \sigma_t^2 &\sim \mbox{Inv-Gamma}\big(a_{\sigma}, b_{\sigma}\big), \quad a_{\sigma} > 0, b_{\sigma} > 0 \\
    {\bf z}_k &\sim \mbox{MVN}_d\big({\bf 0}, {\bf I}_d\big)\\
    {\bf w}_i &\sim \mbox{MVN}_d\big({\bf 0}, {\bf I}_d\big)\\
    \log \gamma_t &\sim \mbox{N}\big(\mu_{\gamma}, \tau_{\gamma}^2\big), \quad \mu_{\gamma} \in \mathbb{R}, \tau_{\gamma}^2 > 0,
\end{split}\]
where ${\bf 0}$ is a $d$-vector of zeros and ${\bf I}_d$ is the $d \times d$ identity matrix. Then, the full posterior of the proposed model is proportional to 
\begin{equation}
\label{lsrm.likelihood}
\begin{split}
&\pi\Big(\boldsymbol\beta_t, \boldsymbol\theta_t, \sigma_t^2, {\bf Z}_t, {\bf W}_t, \gamma_t \mid {\bf X}_t\Big)\\
&\propto \pi\big(\sigma_t^2\big) \pi\big(\gamma_t\big) \prod_{k}\pi\big(\theta_{t,k} \mid \sigma_t^2\big) \prod_{i}\pi\big(\beta_{t,i}\big) \prod_{k}\pi\big({\bf z}_{t,k}\big) \prod_{i}\pi\big({\bf w}_{t,i}\big)\\
&\times \prod_k \prod_i P\Big(x_{t,ki} \mid {\bf z}_{t,k}, {\bf w}_{t,i}, \theta_{t,k}, \beta_{t,i}\Big)^{x_{ki}}\\
&\Big(1-P\big(x_{t,ki} \mid {\bf z}_{t,k}, {\bf w}_{t,i}, \theta_{t,k}, \beta_{t,i}\big)\Big)^{1-x_{ki}}
\end{split}
\end{equation}

%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Model} \label{Chapter4}
\section{Longitudinal Latent Space Item Response Model}\label{llsrm}
When we deal with longitudinal data with a discrete time sequence and there is not any time dependency, we can easily define longitudinal latent space item response model by multiplying the posterior of LSRM $\it{t}$ times where there are $\it{t}$ time periods in the data. The full posterior of the longitudinal LSRM without any time-dependent assumption is proportional to 
\begin{equation}
\label{longitudinal}
\begin{split}
&\pi\Big(\boldsymbol\beta, \boldsymbol\theta, \boldsymbol\sigma^2, {\bf Z}, {\bf W}, \boldsymbol\gamma \mid {\bf X}\Big)\\
&= \prod_{t=1}^T \pi\Big(\boldsymbol\beta_t, \boldsymbol\theta_t, \sigma_t^2, {\bf Z}_t, {\bf W}_t, \gamma_t \mid {\bf X}_t\Big)\\ 
&\propto \prod_{t=1}^T \Bigg\{\pi\big(\sigma_t^2\big) \pi\big(\gamma_t\big) \prod_{k}\pi\big(\theta_{t,k} \mid \sigma^2\big) \prod_{i}\pi\big(\beta_{t,i}\big)
\prod_{k}\pi\big({\bf z}_{t,k}\big) \prod_{i}\pi\big({\bf w}_{t,i}\big)\\
&\times \prod_k \prod_i P\Big(x_{t,ki} \mid {\bf z}_{t,k}, {\bf w}_{t,i}, \theta_{t,k}, \beta_{t,i}\Big)^{x_{ki}}\Big(1-P\big(x_{t,ki} \mid {\bf z}_{t,k}, {\bf w}_{t,i}, \theta_{t,k}, \beta_{t,i}\big)\Big)^{1-x_{ki}}\Bigg\}
\end{split}
\end{equation}

where $\it{T}$ is the number of discrete time periods, and subscript $\it{t}$ of each variable represents corresponding time period.

%% Longitudinal LSRM with Political data
\section{Longitudinal Latent Space Item Response Model for Political Data}

\quad Longitudinal LSRM described above assumes that nodes from each time point are independent. However, this assumption might be a obstacle to describing correct relations between nodes when there is a time dependency in reality. As for U.S. congressional data, it is the longitudinal data which has 3 time periods and a dependency structure in the time.

\quad According to the U.S. congressional data, there are several types of politicians in 417 districts. First of all, there are 271 re-elected and unchanged senators during 3 time periods. This type occupies the largest proportion. Secondly, 81 senators were re-elected in November 2016, but the result could not be the same in November 2018. Third, someone could lose the election in November 2016, but win the following election and get back to the position. However, this third situation did not occur in our data and, then, we can ignore it in the estimation step. Fourth, there can be the situation that a new candidate was elected in November 2016, and keep the position in the next election. 53 districts are under this kind of situation. Finally, a senator might be changed in every time periods. 12 districts changed their representatives every election. (say about vacancy? move to data section?)

\quad It is natural to say that the respondents(politicians) are time-dependent considering these changes of senators through 3 time periods. Therefore, we should take  time dependency into account for modeling. In the beginning, we need to split the whole politician into 5 cases under $T=3$ condition.
%% Grouping
\begin{itemize}
    \item A=\{Same politician at $t=1,2,3$\}. It means ${\bf z}_{1,k}={\bf z}_{2,k}={\bf z}_{3,k}$, and, a prior is $\pi({\bf z}_{k})$.
    \item B=\{Same politician at $t=1,2$\}. It means ${\bf z}_{1,k}={\bf z}_{2,k}\neq {\bf z}_{3,k}$, and, a prior is $\pi({\bf z}_{1,k})\pi({\bf z}_{3,k})$.
    \item C=\{Same politician at $t=1,3$\}. It means ${\bf z}_{1,k}={\bf z}_{3,k}\neq{\bf z}_{2,k}$, and, a prior is $\pi({\bf z}_{1,k})\pi({\bf z}_{2,k})$.
    \item D=\{Same politician at $t=2,3$\}. It means ${\bf z}_{2,k}={\bf z}_{3,k}\neq{\bf z}_{1,k}$, and, a prior is $\pi({\bf z}_{1,k})\pi({\bf z}_{2,k})$.
    \item E=\{In all time period, different politicians are elected\}. It means ${\bf z}_{1,k}\neq{\bf z}_{2,k}\neq{\bf z}_{3,k}$, and, a prior is $\pi({\bf z}_{1,k})\pi({\bf z}_{2,k})\pi({\bf z}_{3,k})$.
\end{itemize}

\quad This grouping method assumes that the same person has the same latent position, and if person changes, the latent position also changes. Of course, a different latent variable has a different prior.

\quad Taking this grouping and the posterior of longitudinal LSRM into account together, the full posterior is proportional to
%% Full posterior
\begin{equation}
\label{full.posterior}
\begin{split}
\pi\Big(\boldsymbol\beta, \boldsymbol\theta, \boldsymbol\sigma^2, &{\bf Z}, {\bf W}, \boldsymbol\gamma \mid {\bf X}\Big) = \prod_{t=1}^T \pi\Big(\boldsymbol\beta_t, \boldsymbol\theta_t, \sigma_t^2, {\bf Z}_t, {\bf W}_t, \gamma_t \mid {\bf X}_t\Big)\\ 
&\propto \prod_{t=1}^T \Bigg\{\pi\big(\sigma_t^2\big) \pi\big(\gamma_t\big) \prod_{k}\pi\big(\theta_{t,k} \mid \sigma^2\big) \prod_{i}\pi\big(\beta_{t,i}\big)
\prod_{i}\pi\big({\bf w}_{t,i}\big) \Bigg\}\\
&\times \prod_{K\in A}\Big \{\pi({\bf z}_{k}) L_{11} L_{12} L_{13} \Big \}\\
&\times \prod_{K\in B}\Big \{\pi({\bf z}_{1,k}) \pi({\bf z}_{3,k}) L_{11} L_{12} L_{33}\Big \}\\
&\times \prod_{K\in C}\Big \{\pi({\bf z}_{1,k}) \pi({\bf z}_{2,k}) L_{11} L_{13} L_{22} \Big \}\\
&\times \prod_{K\in D}\Big \{\pi({\bf z}_{1,k}) \pi({\bf z}_{2,k}) L_{11} L_{22} L_{23} \Big \}\\
&\times \prod_{K\in E}\Big \{\pi({\bf z}_{1,k}) \pi({\bf z}_{2,k}) \pi({\bf z}_{3,k}) L_{11} L_{22} L_{33} \Big \},
\end{split}
\end{equation}
where
\begin{equation}\notag
\begin{split}
&L_{t_1,t_2} = L({\bf z}_{t_1,k}, {\boldsymbol \phi}_{t_2,k,i})\\
&= \prod_{k=1}^n\prod_{i=1}^p P\Big(x_{t_2,ki} \mid {\bf z}_{t_1,k}, {\bf w}_{t_2,i}, \theta_{t_2,k}, \beta_{t_2,i}\Big)^{x_{t_2,ki}}
\Big(1-P\big(x_{t_2,ki} \mid {\bf z}_{t_1,k}, {\bf w}_{t_2,i}, \theta_{t_2,k}, \beta_{t_2,i}\big)\Big)^{1-x_{t_2,ki}}
\end{split}
\end{equation}

\quad $\it{K}$ is the set of respondents in the certain group among the five. $\boldsymbol{\phi}_{t_2,k,i}$ represents the parameters, such as $x_{t_2,ki}, {\bf w}_{t_2,i}, \theta_{t_2,k},$ and $\beta_{t_2,i}$, except ${\bf z}_{t_1,k}$ . The subscripts of likelihood $L_{t_1,t_2}$ indicate the time points to which each parameter corresponds. $t_1$ and $t_2$ indicate the time of ${\bf z}_{t_1,k}$ and $\boldsymbol{\phi}_{t_2,k,i}$ each. Now, we can see that the prior part of ${\bf z}$ and the likelihood part are devided in five according to the politician grouping. $\prod_{K\in A}\Big \{\pi({\bf z}_{k}) L_{11} L_{12} L_{13} \Big \}$ is the prior and likelihood of the respondents in a group A. Politicians of this group keep own position all the time. Therefore, the latent position is unchanged, and it means there is the only one $\pi({\bf z}_{k})$ regardless of time. As for $L_{11} L_{12} L_{13}$, the parameter ${\bf z}_{t_1,k}$ is the same with ${\bf z}_{1,k}$ at the all three times when the subscript 1 of ${\bf z}_{1,k}$ indicates the first time point, 2016, but the other parameters change along with the time sequence. In case of group B, a person in 2016 and 2018 is the same, and only 2020's person is different. The former person has the ${\bf z}_{1,k}$ as a latent position, and the later has the ${\bf z}_{3,k}$ when subscript 3 indicates the third time point, 2020. During the first two sequences, the latent position ${\bf z}_{1,k}$ is unchanged, therefore, the first subscript of $L_{11}$ and $L_{12}$, indicating the time points of the latent position, is also unchanged. Unlike the latent position ${\bf z}$, the other parameters always changes in the all time points. The second subscript of all three $L$ is different, and, the same thing happens in the all group. In the same way, the likelihoods of group C and D also can be defined. In the last, group E is the simplest case. Politicians change every time, and it means that the latent positions also change. Therefore, the function contains all three prior, $\pi({\bf z}_{1,k})$, $\pi({\bf z}_{2,k})$, and $\pi({\bf z}_{3,k})$. Furthermore, the time of latent position and other parameters matches. Then, the function is like $\prod_{K\in E}\Big \{\pi({\bf z}_{1,k}) \pi({\bf z}_{2,k}) \pi({\bf z}_{3,k}) L_{11} L_{22} L_{33} \Big \}$.

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Estimation} \label{Chapter5}
We used a fully Bayesian approach based on Markov chain Monte Carlo (MCMC) sampling to estimate the proposed model using the Gibbs sampling and the Metropolis-Hastings algorithm. We specify prior distributions for the model parameters as follows:
\[\begin{split}
    \beta_{t,i} \mid \tau_{\beta}^2 &\sim \mbox{N}\big(0, \tau_{\beta}^2\big), \quad \tau_{\beta}^2 > 0 \\
    \theta_{t,k} \mid \sigma^2_t &\sim \mbox{N}\big(0, \sigma_t^2\big), \quad \sigma_t^2 > 0 \\
    \sigma_t^2 &\sim \mbox{Inv-Gamma}\big(a_{\sigma}, b_{\sigma}\big), \quad a_{\sigma} > 0, b_{\sigma} > 0 \\
    \boldsymbol{z}_{t,k} &\sim \mbox{MVN}_d\big({\bf 0}, {\bf I}_d\big)\\
    \boldsymbol{w}_{t,i} &\sim \mbox{MVN}_d\big({\bf 0}, {\bf I}_d\big),\\
\end{split}\]
where ${\bf 0}$ is a $d$-vector of zeros and ${\bf I}_d$ is the $d \times d$ identity matrix. We set the prior parameters like the following choices: $\tau_{\beta}^2 = 4$, $a_{\sigma} = b_{\sigma} = 0.001$. Also, we regard $\gamma_t$, a scale parameter for the respondent and item latent positions, as a constant 1 for the computational easiness.

\quad The full posterior of the proposed model is proportional to (\ref{full.posterior}). Our MCMC sampler iterates over the model parameters with the priors given above. At iteration $l$, the MCMC sampling can be described as follows:

\begin{enumerate}
%% beta
    \item Use Metropolis-Hastings steps to sample $\beta_{t,i}$ for each $i$ and $t$. Propose $\beta^\prime_{t,i}$ from a proposal distribution $\varphi(\cdot)$ and accept with probability equal to
\begin{equation*}
    r_\beta \Big( \beta^\prime_{t,i},\beta^{(l)}_{t,i} \Big) = \frac{\pi \Big( \beta^\prime_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)}{\pi \big( \beta^{(l)}_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big)}\frac{\varphi \Big(\beta^\prime_{t,i}\rightarrow \beta^{(l)}_{t,i} \Big)}{\varphi \Big(\beta^{(l)}_{t,i}\rightarrow \beta^\prime_{t,i} \Big)}
\end{equation*}
where $\pi \Big( \beta_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)$ is the full conditional posterior for $\beta_{t,i}$.
%% theta
\item Propose $\theta^\prime_{t,k}$ from a proposal distribution $\varphi(\cdot)$ and accept the proposed value with probability
\begin{equation*}
    r_\theta \Big( \theta^\prime_{t,k},\theta^{(l)}_{t,k} \Big) = \frac{\pi \Big( \theta^\prime_{t,k} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \beta}_t,\sigma^2_t \Big)}{\pi \big( \theta^{(l)}_{t,k} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \beta}_t,\sigma^2_t\Big)}\frac{\varphi \Big(\theta^\prime_{t,k}\rightarrow \theta^{(l)}_{t,k} \Big)}{\varphi \Big(\theta^{(l)}_{t,k}\rightarrow \theta^\prime_{t,k} \Big)}
\end{equation*}
where $\pi \Big( \theta_{t,k} \mid {\bf X}_t,{\bf Z}_t,{\bf W}_t,{\boldsymbol \beta}_t,\sigma^2_t \Big)$ is the full conditional posterior for $\theta_{t,k}$.
%% W
\item Propose ${\bf w}^\prime_{t,i}$ from a proposal distribution $\varphi(\cdot)$ and accept with probability equal to
\begin{equation*}
    r_w \Big( {\bf w}^\prime_{t,i},{\bf w}^{(l)}_{t,i} \Big) = \frac{\pi \Big( {\bf w}^\prime_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)}{\pi \big( {\bf w}^{(l)}_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big)}\frac{\varphi \Big({\bf w}^\prime_{t,i}\rightarrow {\bf w}^{(l)}_{t,i} \Big)}{\varphi \Big({\bf w}^{(l)}_{t,i}\rightarrow {\bf w}^\prime_{t,i} \Big)}
\end{equation*}
where $\pi \Big( {\bf w}_{t,i} \mid {\bf X}_t,{\bf Z}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)$ is the full conditional posterior for ${\bf w}_{t,i}$.
%% Z
\item Propose ${\bf z}^\prime_{t,k}$ from a proposal distribution $\varphi(\cdot)$ and accept the proposed value with probability equal to

\begin{equation*}
    r_z \Big( {\bf z}^\prime_{t,k},{\bf z}^{(l)}_{t,k} \Big) = \frac{\pi \Big( {\bf z}^\prime_{t,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)}{\pi \big( {\bf z}^{(l)}_{t,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big)}\frac{\varphi \Big({\bf z}^\prime_{t,k}\rightarrow {\bf z}^{(l)}_{t,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{t,k}\rightarrow {\bf z}^\prime_{t,k} \Big)}
\end{equation*}
where $\pi \Big( {\bf z}_{t,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)$ is the full conditional posterior for ${\bf z}_{t,k}$. However, the MH ratio differs dependently on the group to which each ${\bf z}_{t,k}$ belongs.

\begin{enumerate}
%% Z_A
    \item  If ${\bf z}_{t,k} \in A$, the acceptance rate is equal to
    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{1,k},{\bf z}^{(l)}_{1,k} \Big) = \frac{\prod^3_{t=1} \Big[ \pi \Big( {\bf z}^\prime_{1,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)\Big]}{\prod^3_{t=1} \Big[ \pi \big( {\bf z}^{(l)}_{1,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big) \Big]}\frac{\varphi \Big({\bf z}^\prime_{1,k}\rightarrow {\bf z}^{(l)}_{1,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{1,k}\rightarrow {\bf z}^\prime_{1,k} \Big)},
    \end{equation*}
    where ${\bf z}_{1,k} = {\bf z}_{2,k} = {\bf z}_{3,k}$.

    %% Z_B
    \item If ${\bf z}_{t,k} \in B$, the acceptance rate is equal to
    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{1,k},{\bf z}^{(l)}_{1,k} \Big) = \frac{\prod^2_{t=1} \Big[ \pi \Big( {\bf z}^\prime_{1,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)\Big]}{\prod^2_{t=1} \Big[ \pi \big( {\bf z}^{(l)}_{1,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big) \Big]}\frac{\varphi \Big({\bf z}^\prime_{1,k}\rightarrow {\bf z}^{(l)}_{1,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{1,k}\rightarrow {\bf z}^\prime_{1,k} \Big)},
    \end{equation*}

    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{3,k},{\bf z}^{(l)}_{3,k} \Big) = \frac{\pi \Big( {\bf z}^\prime_{3,k} \mid {\bf X}_3,{\bf W}_3,{\boldsymbol \beta}_3,{\boldsymbol \theta}_3,\sigma^2_3 \Big)}{\pi \big( {\bf z}^{(l)}_{3,k} \mid {\bf X}_3,{\bf W}_3,{\boldsymbol \beta}_3,{\boldsymbol \theta}_3,\sigma^2_3\Big) }\frac{\varphi \Big({\bf z}^\prime_{3,k}\rightarrow {\bf z}^{(l)}_{3,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{3,k}\rightarrow {\bf z}^\prime_{3,k} \Big)},
    \end{equation*}
    where ${\bf z}_{1,k} = {\bf z}_{2,k} \neq {\bf z}_{3,k} $.

    %% Z_D
    \item If ${\bf z}_{t,k} \in D$, the acceptance rate is equal to
    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{1,k},{\bf z}^{(l)}_{1,k} \Big) = \frac{ \pi \Big( {\bf z}^\prime_{1,k} \mid {\bf X}_1,{\bf W}_1,{\boldsymbol \beta}_1,{\boldsymbol \theta}_1,\sigma^2_1 \Big)}{ \pi \big( {\bf z}^{(l)}_{1,k} \mid {\bf X}_1,{\bf W}_1,{\boldsymbol \beta}_1,{\boldsymbol \theta}_1,\sigma^2_1\Big) }\frac{\varphi \Big({\bf z}^\prime_{1,k}\rightarrow {\bf z}^{(l)}_{1,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{1,k}\rightarrow {\bf z}^\prime_{1,k} \Big)},
    \end{equation*}

    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{2,k},{\bf z}^{(l)}_{2,k} \Big) = \frac{\prod^3_{t=2} \Big[ \pi \Big( {\bf z}^\prime_{2,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)\Big]}{\prod^3_{t=2} \Big[ \pi \big( {\bf z}^{(l)}_{2,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big) \Big]}\frac{\varphi \Big({\bf z}^\prime_{2,k}\rightarrow {\bf z}^{(l)}_{2,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{2,k}\rightarrow {\bf z}^\prime_{2,k} \Big)},
    \end{equation*}
    where ${\bf z}_{1,k} \neq {\bf z}_{2,k} = {\bf z}_{3,k} $.

    %% Z_E
    \item If ${\bf z}_{t,k} \in E$, the acceptance rate is equal to
    \begin{equation*}
    r_z \Big( {\bf z}^\prime_{t,k},{\bf z}^{(l)}_{t,k} \Big) = \frac{\pi \Big( {\bf z}^\prime_{t,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t \Big)}{\pi \big( {\bf z}^{(l)}_{t,k} \mid {\bf X}_t,{\bf W}_t,{\boldsymbol \beta}_t,{\boldsymbol \theta}_t,\sigma^2_t\Big)}\frac{\varphi \Big({\bf z}^\prime_{t,k}\rightarrow {\bf z}^{(l)}_{t,k} \Big)}{\varphi \Big({\bf z}^{(l)}_{t,k}\rightarrow {\bf z}^\prime_{t,k} \Big)},
    \end{equation*}
    where ${\bf z}_{1,k} \neq {\bf z}_{2,k} \neq {\bf z}_{3,k} $.

\end{enumerate}

%% sigma
\item Update $\sigma^2_t$ from the Inverse Gamma distribution
\begin{equation*}
    \sigma^2_t \sim \text{Inv-Gamma} \Bigg( a_\sigma + \frac{N}{2}, b_\sigma + \frac{\sum_k\theta^2_{t,k}}{2} \Bigg)
\end{equation*}

%% imputation
\item Impute missing value $x_{t,ki}$ from the binomial distribution with the probability of
\begin{equation*}
    \mbox{logit}\Big(X_{t,ki} = 1 \mid \beta_{t,i}^{(l+1)}, \theta_{t,k}^{(l+1)}, {\bf z}_{t,k}^{(l+1)}, {\bf w}_{t,i}^{(l+1)}\Big)
    = \beta_{t,i}^{(l+1)} + \theta_{t,k}^{(l+1)} - \Big|\Big|{\bf z}_{t,k}^{(l+1)} - {\bf w}_{t,i}^{(l+1)}\Big|\Big|.
\end{equation*}

\end{enumerate}

\quad All proposed values are drawn from normal distributions centered around the parameter values at iteration $l$ with tunable variance parameters. For instance, $\varphi \big(\theta_{t,i}^{(l)} \rightarrow \theta_{t,i}^{\prime}\big) \sim N\big(\theta_{t,i}^{(l)}, \sigma^2_{t,prop} \big)$. The convergence of the parameters is checked using trace plots.

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Results} \label{Chapter6}
\section{Estimated Latent Space}
\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{LS_ggplot_2.png}
\caption{Latent positions of politicians and bills.}
\label{fig:ls}
\end{center}
\end{figure}

\quad After computation, we have had the estimates of $\bf{z}$ and $\bf{w}$ for 2016, 2018, and 2020. Now we can show the latent position of three periods of politicians and bills in a two-dimensional plane at once like Figure \ref{fig:ls}. In Figure \ref{fig:ls}, red spots represent the latent positions of republicans, and blue ones do those of democrats. Black spots are the latent positions of bills. If looking at the pattern of the latent positions,
we can easily know that republicans and democrats are separated clearly. Furthermore, we can also see that the bills are grouped in three parts, one of which is close to the republican party, the other one close to the democrat party, and the another located between the two parties. According to the latent space item response model, a distance between latent positions is the measure of the correlation between them. Then, we can interpret that the bills positioned in the territory of a certain political party is only supported by that party, but not by the opposite party. On the other hand, the bills in the middle area are correlated to both parties at the same level. It means that most politicians supported or rejected the mid-area bills in the same regardless of their political stance. As for politicians comparatively located in the mid-area, we can say that they did not follow their party blindly.

\begin{figure}
  \begin{center}
    \subfigure[2016]{\label{fig:annual-a}\includegraphics[scale=0.25]{LS_2016.png}}
    \subfigure[2018]{\label{fig:annual-b}\includegraphics[scale=0.25]{LS_2018.png}}
    \subfigure[2020]{\label{fig:annual-c}\includegraphics[scale=0.25]{LS_2020.png}}
  \end{center}
  \caption{Latent positions in 2016, 2018, and 2020. }
  \label{fig:annual}
\end{figure}

\quad We depict the time of each point in Figure \ref{fig:ls} by differentiating the shape of a spot. A circle, a triangle, and a rectangle mean that the latent position belongs to 2016, 2018, and 2020. There is an overlapped politician's latent position through several time-periods, and it means that the politician took office multiple times. In other words, the same person takes the same latent position in the latent space. Because the same bill is not usually put to the vote twice at the congress, there are not identical latent positions among those of the bills.

\quad As shown in Figure \ref{fig:ls}, our model succeeded in catching the political correlations between politicians and bills. The house of representatives is split in two parties, as we expected, and bills also show the correlation with the parties. Moreover, the time characteristic of the latent position is also included.

\section{Politicians in the middle area}

\quad Considering U.S. politicians' high loyalty toward their own political party, it is not hard to understand that politicians are clustered in a clean pattern.Therefore, it is meaningful to investigate the politicians who are positioned in the middle area, relatively far from their cluster. They are in such an area because of voting the opposite party's bill more frequently than other comrades. It means they have a moderate political attitude, and in U.S. politics, a politician who has this kind of stance is called "Bipartisan." Among the bipartisan, we pick the most evident people and look into the characteristics of their congressional districts and themselves. We chose 4 Democrats and 2 Republicans in the middle area.

\begin{table}[h!]
\centering
    \begin{tabular}{c | c | c | c | c}
        ID & State & District & Name & Party  \\
        \hline \hline
        400316 & Minnesota & 7 & Collin Peterson & Democrat \\
        412420 & Illinois & 10 & Bob Dold & Republican \\
        412796 & New Jersey & 2 & Jefferson H. Van Drew & Republican  \\
        412808 & Oklahoma & 5 & Kendra Horn & Democrat  \\
        412814 & South Carolina & 1 & Joe Cunningham & Democrat  \\
        412829 & Utah & 4 & Ben McAdams & Democrat  \\
    \end{tabular}
    \caption{Bipartisan far from the party}
    \label{tab:data}
\end{table}

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{districts.png}
  \end{center}
  \caption{Congressional districts of U.S. }
  \label{fig:districts}
\end{figure}

\quad In the analysis, we accept several beliefs in political tendency. First, Republican is conservative, and Democrat is progressive. Second, an urban region, including the big city, is more politically progressive. Third, each state excepting the swing state supports a certain political party. Taking it into consideration, the analysis is undertaken.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{mn.png}
  \end{center}
  \caption{Congressional districts of Minnesota }
  \label{fig:mn}
\end{figure}

\quad Minnesota, having Minneapolis as a capital city, is a relatively progressive state. Among the districts of Minnesota, the 7th district covers almost all of the western side of Minnesota. In this region, the proportion of the rural area is larger than that of the urban area. Moreover, the district borders with North Dakota and South Dakota, which are conservative states. Considering this geographic characteristic, the 7th district is likely to be conservative. Therefore, it is natural that the district's representative, Collin Peterson, is a conservative Democrat, reflecting both the state's progressive personality and the conservative aspect of the district. Peterson is one of the founders of the Blue Dog Coalition, a group of moderate and conservative Democrats. Peterson has split from his party on issues such as gay marriage, healthcare, the estate tax, gun control, the environment, and abortion. It is no doubt that he successfully represented the district's opinion considering that he was first elected in 1990 and kept re-elected until 2018.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{il.png}
  \end{center}
  \caption{Congressional districts of Illinois}
  \label{fig:il}
\end{figure}

\quad Illinois is known as one of the progressive states, placing the large city of Chicago. Unlike our anticipation, the Republican politician was elected in the district of the progressive state. Bob Dold, the representative for Illinois's 10th congressional district in 2016, is a moderate Republican who supports gun control measures. Additionally, his predecessor, Brad Schneider, is also known as moderate Democrat. From the characteristic of representatives, it is possible to say that 10th district of Illinois prefers a moderate politician. We can deduce the reasons for the political position of this district from geographic factors as well. 10th district lies in the northeast corner of the state, and mostly comprises northern suburbs of Chicago. In this region, the district covers Lake county which is the second most wealthy county in Illinois. When it is believed that wealthy people tend to be more conservative in common sense, it is natural thinking that the 10th district is politically moderate despite it is near Chicago.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{nj.png}
  \end{center}
  \caption{Congressional districts of New Jersey }
  \label{fig:nj}
\end{figure}

\quad New Jersey is also a progressive state influenced by an adjacent state, New York. However, the 2nd district elected Republican representative. The district, based in Southern New Jersey, is a swing region with a history of supporting different parties. The proportion of the rural area is relatively high, and there is no metropolis in this district. Moreover, it is located most far from New York and less influenced by New York City. For this reason, the 2nd district seems to be more conservative than the average of New Jersey.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{ok.png}
  \end{center}
  \caption{Congressional districts of Oklahoma }
  \label{fig:ok}
\end{figure}

\quad Oklahoma is a state in the South Central region of the United States and it is a conservative state. However, the 5th district includes Oklahoma City, and it is likely to have a moderate political stance. The result of the election of Kendra Horn, the representative of Democrat, demonstrates this enough. The district's preference is obvious, seeing the fact that Horn was joining the Blue Dog Coalition, a group of moderate and conservative Democrats.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{sc.png}
  \end{center}
  \caption{Congressional districts of South Carolina }
  \label{fig:sc}
\end{figure}

\quad South Carolina is known as a conservative state. Conversely, the 1st district elected the Democrat politician. The 1st district has historically been based in Charleston, the largest city in South Carolina. The urban characteristic may influence the political position.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{utah.png}
  \end{center}
  \caption{Congressional districts of Utah }
  \label{fig:utah}
\end{figure}

\quad Utah is famous as conservative state because a little more than half of all Utahns are Mormons. The 4th district representative, Ben McAdams is the only Democratic member of Utah's congressional delegation, but a member of the Blue Dog Coalition, a group of a moderate politician. Despite the Mormons' headquarter, Salt Lake City, the most populous city in Utah, tends to affect progressive in a political position.

\quad Interpreting the patterns found above, the commonality can be seen that Bipartisan is elected in the district of a moderate political position compared with that of the whole state. Moreover, we infer that the degree of the district's political orientation varies with the geographical feature. For the districts of the progressive state, the district with the broad rural area or covering wealthy county tends to be more conservative than other districts in the state. As for the districts of the conservative state, the district containing a large city is more progressive than others. In addition, the political atmosphere of neighboring states affects the residents' political ideas of the district.

\section{The Bills in the Middle Area}

\begin{figure}[h]
  \begin{center}
    \subfigure[2016]{\label{fig:bill-a}\includegraphics[scale=0.18]{vote_2016.png}}
    \subfigure[2018]{\label{fig:bill-b}\includegraphics[scale=0.18]{vote_2018.png}}
    \subfigure[2020]{\label{fig:bill-c}\includegraphics[scale=0.18]{vote_2020.png}}
  \end{center}
  \caption{Latent positions of bills and their numbers in 2016, 2018, and 2020. }
  \label{fig:bill}
\end{figure}

\section{Outliers}

\end{itemize}

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Conclusion} \label{Chapter7}

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%
\chapter{Appendix}\label{Appendix}

\section{Trace Plots}
\subsection{$\theta_{t,k}$}
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.4]{ts_theta_1.png}
  \end{center}
  \caption{Trace plot of $\theta_{t,1}$}
  \label{fig:ts_theta}
\end{figure}

%----------------------------------------------------------
\newpage

\section*{국 문 요 약}
\setstretch{1.8}
\par
\bigskip

\begin{center}
\large
종단 잠재 공간 항목 반응 모형의 응용,\\
미 하원 의회 투표 자료 분석을 중심으로
\end{center}
\setstretch{1.8}
\par
\bigskip
\noindent

\bigskip
\noindent \hrulefill\\
핵심 용어 :  \par

%\pagebreak

\newpage

%--------------------------------------------------------------------------------%
%--------------------------------------------------------------------------------%

%\nocite{*}
\bibliographystyle{apalike}
\bibliography{References}

\end{document}

### 2020-12-06 BayesianSTERGM/BSTERGM_MCMC.h

@@ -54,8 +54,13 @@ private:
         }
     }
     double log_paramPriorPDF(Col<double> param_formation, Col<double> param_dissolution) {
-        //NOW: model parameter prior : 1
-        return 0;
+        double log_prior_val;
+        double sigma2 = 100;
+        Mat<double> CovInv = eye(n_paramDim, n_paramDim)/(2*sigma2);
+        Col<double> formation_part = param_formation.t() * CovInv * param_formation;
+        Col<double> dissolution_part = param_dissolution.t() * CovInv * param_dissolution;
+        log_prior_val = formation_part(0) + dissolution_part(0);
+        return log_prior_val;
     }
     Col<double> proposeParam(Col<double> lastParam, double varRate) {
         Mat<double> proposalCov(n_paramDim, n_paramDim, fill::eye);
@@ -99,14 +104,13 @@ private:
         log_r_val -= log_paramPriorPDF(param_lastFormation, param_lastDissolution);
         return log_r_val;
     }
-    void sampler(int num_exchangeMCiter) {
+    void sampler(int num_exchangeMCiter, double proposalVarRate) {
         Col<double> param_lastFormation = paramVec_formation.back();
         Col<double> param_lastDissolution = paramVec_dissolution.back();
 
         //proposal
-        double varRate = 0.01;
-        Col<double> param_newFormation = proposeParam(param_lastFormation, varRate);
-        Col<double> param_newDissolution = proposeParam(param_lastDissolution, varRate);
+        Col<double> param_newFormation = proposeParam(param_lastFormation, proposalVarRate);
+        Col<double> param_newDissolution = proposeParam(param_lastDissolution, proposalVarRate);
 
         //select timeLag
         int lag_startTimePoint = randi<int>(distr_param(0, T_time - 2));
@@ -157,12 +161,12 @@ public:
         dissociateObservedSeq();
     }
 
-    void generateSample(int num_mainMCiter, int num_exchangeMCiter) {
+    void generateSample(int num_mainMCiter, int num_exchangeMCiter, double proposalVarRate) {
         for (int i = 0; i < num_mainMCiter; i++) {
             if (i % 200 == 0) {
                 cout << "MCMC : " << n_iterated << "/" << num_mainMCiter << endl;
             }
-            sampler(num_exchangeMCiter);
+            sampler(num_exchangeMCiter, proposalVarRate);
         }
[…]

### 2020-12-06 BayesianSTERGM/main.cpp

@@ -669,12 +669,12 @@ int main()
     //=================================================================================================
     //BSTERGM test
     
-    Col<double> initFormationParam = { 0.01 };
-    Col<double> initDissolutionParam = { -0.01 };
-    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_1_seq);
-    Bstergm.generateSample(50000, 20);
-    Bstergm.cutBurnIn(10000);
-    Bstergm.thinning(50);
+    Col<double> initFormationParam = { -0.01 };
+    Col<double> initDissolutionParam = { 0.01 };
+    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_2_seq);
+    Bstergm.generateSample(10000, 500, 0.001);
+    Bstergm.cutBurnIn(1000);
+    Bstergm.thinning(20);
 
     //BSTERGM posterior sample diagnostics
     Diagnostics_MCParamSample BstergmDiag1(Bstergm.getPosteriorSample_formation());

### 2020-12-07 BayesianSTERGM/BSTERGM_MCMC.h
model specify : node + GWESP

@@ -83,19 +83,19 @@ private:
         // now model
         model_delta_exchangeFormation_1 += {
             (double)exchange_Formation.get_n_Edge(),
-                // (double)exchange_Formation.get_undirected_geoWeightedESP(0.5)
+                (double)exchange_Formation.get_undirected_geoWeightedESP(0.5)
         };
         model_delta_exchangeDissolution_2 += {
             (double)exchange_Dissolution.get_n_Edge(),
-                // (double)exchange_Dissolution.get_undirected_geoWeightedESP(0.5)
+                (double)exchange_Dissolution.get_undirected_geoWeightedESP(0.5)
         };
         model_delta_obsFormation_3 += {
             (double)observed_FormationSeq[lag_startTimePoint + 1].get_n_Edge(),
-                // (double)observed_FormationSeq[lag_startTimePoint + 1].get_undirected_geoWeightedESP(0.5)
+                (double)observed_FormationSeq[lag_startTimePoint + 1].get_undirected_geoWeightedESP(0.5)
         };
         model_delta_obsDissolution_4 += {
             (double)observed_DissolutionSeq[lag_startTimePoint + 1].get_n_Edge(),
-                // (double)observed_DissolutionSeq[lag_startTimePoint + 1].get_undirected_geoWeightedESP(0.5)
+                (double)observed_DissolutionSeq[lag_startTimePoint + 1].get_undirected_geoWeightedESP(0.5)
         };
 
         log_r_val += dot(param_lastFormation - param_newFormation, model_delta_exchangeFormation_1 - model_delta_obsFormation_3);

### 2020-12-07 BayesianSTERGM/BSTERGM_emp_data_run/make_plots.R

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-07 BayesianSTERGM/BSTERGM_emp_data_run_edgeonly/make_plots.R

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run_edgeonly/BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run_edgeonly/BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/1_seq2_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run_edgeonly/BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-07 BayesianSTERGM/NetMCMCSampler_STERGM.h

@@ -53,8 +53,8 @@ private:
 
     double log_r(Network& last, Network& proposed, bool isDissolution) {
         //NOW MODEL
-        Col<double> model_delta = { (double)proposed.get_n_Edge() - last.get_n_Edge(), };
-                                    // (double)proposed.get_undirected_geoWeightedESP(0.5) - last.get_undirected_geoWeightedESP (0.5) }; // <-model specify
+        Col<double> model_delta = { (double)proposed.get_n_Edge() - last.get_n_Edge(),
+                                     (double)proposed.get_undirected_geoWeightedESP(0.5) - last.get_undirected_geoWeightedESP (0.5) }; // <-model specify
         Col<double> param;
         if (isDissolution) {
             param = dissolution_Param;

### 2020-12-07 BayesianSTERGM/main.cpp

@@ -669,12 +669,12 @@ int main()
     //=================================================================================================
     //BSTERGM test
     
-    Col<double> initFormationParam = { -0.01 };
-    Col<double> initDissolutionParam = { 0.01 };
+    Col<double> initFormationParam = { -2, 1 };
+    Col<double> initDissolutionParam = { -1, 2 };
     BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_2_seq);
-    Bstergm.generateSample(10000, 500, 0.001);
-    Bstergm.cutBurnIn(1000);
-    Bstergm.thinning(20);
+    Bstergm.generateSample(25000, 300, 0.001);
+    //Bstergm.cutBurnIn(1000);
+    //Bstergm.thinning();
 
     //BSTERGM posterior sample diagnostics
     Diagnostics_MCParamSample BstergmDiag1(Bstergm.getPosteriorSample_formation());
@@ -682,17 +682,17 @@ int main()
     Col<double> quantilePts = { 0.1, 0.25, 0.5, 0.75, 0.9 };
     BstergmDiag1.print_quantile(0, quantilePts);
     BstergmDiag1.print_autoCorr(0, 30);
-    /*BstergmDiag1.print_mean(1);
+    BstergmDiag1.print_mean(1);
     BstergmDiag1.print_quantile(1, quantilePts);
-    BstergmDiag1.print_autoCorr(1, 30);*/
+    BstergmDiag1.print_autoCorr(1, 30);
 
     Diagnostics_MCParamSample BstergmDiag2(Bstergm.getPosteriorSample_dissolution());
     BstergmDiag2.print_mean(0);
     BstergmDiag2.print_quantile(0, quantilePts);
     BstergmDiag2.print_autoCorr(0, 30);
-    /*BstergmDiag2.print_mean(1);
+    BstergmDiag2.print_mean(1);
     BstergmDiag2.print_quantile(1, quantilePts);
-    BstergmDiag2.print_autoCorr(1, 30);*/
+    BstergmDiag2.print_autoCorr(1, 30);
 
     BstergmDiag1.writeToCsv_Sample("BSTERGM_formation.csv");
     BstergmDiag2.writeToCsv_Sample("BSTERGM_dissolution.csv");

### 2020-12-07 BayesianSTERGM/BSTERGM_emp_data_run/make_plots.R
seq1 chain1

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain1_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-07 BayesianSTERGM/main.cpp

@@ -669,10 +669,10 @@ int main()
     //=================================================================================================
     //BSTERGM test
     
-    Col<double> initFormationParam = { -2, 1 };
-    Col<double> initDissolutionParam = { -1, 2 };
-    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_2_seq);
-    Bstergm.generateSample(25000, 300, 0.001);
+    Col<double> initFormationParam = { -0.2, 0.01 };
+    Col<double> initDissolutionParam = { -0.2, 0.01 };
+    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_1_seq);
+    Bstergm.generateSample(30000, 400, 0.001);
     //Bstergm.cutBurnIn(1000);
     //Bstergm.thinning();
 

### 2020-12-08 BayesianSTERGM/BSTERGM_emp_data_run/make_plots.R
seq1chain2

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-08 BayesianSTERGM/BSTERGM_emp_data_run/make_plots.R
seq3chain1

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain2_BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain2_BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain1_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq2_chain2_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-08 BayesianSTERGM/main.cpp

@@ -669,10 +669,11 @@ int main()
     //=================================================================================================
     //BSTERGM test
     
-    Col<double> initFormationParam = { -0.2, 0.01 };
-    Col<double> initDissolutionParam = { -0.2, 0.01 };
-    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_1_seq);
-    Bstergm.generateSample(30000, 400, 0.001);
+    Col<double> initFormationParam = { -0.1, 0.01 };
+    Col<double> initDissolutionParam = { -0.1, 0.01 };
+    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_3_seq);
+    //아래 GOF seq도 바꿀 것!
+    Bstergm.generateSample(20000, 400, 0.001);
     //Bstergm.cutBurnIn(1000);
     //Bstergm.thinning();
 
@@ -698,12 +699,12 @@ int main()
     BstergmDiag2.writeToCsv_Sample("BSTERGM_dissolution.csv");
     
     //BSTERGM GOF
-    GoodnessOfFit_STERGM BstergmGoF0 (Bstergm.getPosteriorSample_formation(), Bstergm.getPosteriorSample_dissolution(), net_1_seq);
+    GoodnessOfFit_STERGM BstergmGoF0 (Bstergm.getPosteriorSample_formation(), Bstergm.getPosteriorSample_dissolution(), net_3_seq);
     cout << "t=0 to t=1" << endl;
     BstergmGoF0.undirected_run(0, 100, 100);
     BstergmGoF0.undirected_printResult(0);
 
-    GoodnessOfFit_STERGM BstergmGoF1(Bstergm.getPosteriorSample_formation(), Bstergm.getPosteriorSample_dissolution(), net_1_seq);
+    GoodnessOfFit_STERGM BstergmGoF1(Bstergm.getPosteriorSample_formation(), Bstergm.getPosteriorSample_dissolution(), net_3_seq);
     cout << "\n\nt=1 to t=2" << endl;
     BstergmGoF1.undirected_run(1, 100, 100);
     BstergmGoF1.undirected_printResult(1);

### 2020-12-09 BayesianSTERGM/BSTERGM_MCMC.h
seq3 chain4

@@ -210,7 +210,7 @@ public:
 
 
 
-// old code block
+// old block
 //class BSTERGM_MCMC {
 //private:
 //    vector<Network> observedSeq;

### 2020-12-09 BayesianSTERGM/BSTERGM_emp_data_run/make_plots.R

@@ -1,6 +1,6 @@
 #traceplot
 
-formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_formation.csv", header=FALSE)
+formation_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq3_chain3_BSTERGM_formation.csv", header=FALSE)
 mean(formation_posterior_sample$V1)
 mean(formation_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -13,7 +13,7 @@ acf(formation_posterior_sample$V2)
 
 
 
-dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_dissolution.csv", header=FALSE)
+dissolution_posterior_sample = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq3_chain3_BSTERGM_dissolution.csv", header=FALSE)
 mean(dissolution_posterior_sample$V1)
 mean(dissolution_posterior_sample$V2)
 par(mfrow=c(1,2))
@@ -25,7 +25,7 @@ acf(dissolution_posterior_sample$V2)
 
 
 
-network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq1_chain2_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
+network_diag_stats = read.csv("C:/gitProject/BayesianSTERGM/BayesianSTERGM/BSTERGM_emp_data_run/seq3_chain3_BSTERGM_lastExNetSamplerNetworkStats.csv", header=TRUE)
 
 # for(i in 1:length(network_diag_stats)){
 length(network_diag_stats) #11

### 2020-12-10 doc/doc.tex
model specify : node + GWESP
tex 1211

@@
 \begin{abstract}
-Separable Temporal Exponential Random Graph Models (STERGM) are useful models to describe changes in dynamic networks' structures evolving over time. 
+Separable Temporal Exponential Random Graph Models (STERGMs) are useful models to describe changes in dynamic networks' structures evolving over time. 
 The Bayesian-inference based approach for the model, however, had not been developed yet. 
-In this paper, I propose the Bayesian STERGM (BSTERGM): STERGM with the Bayesian setting, and I illustrate the computational algorithm that generates posterior samples from the model dealing with the doubly-intractable constant problem.
+In this paper, I propose the Bayesian STERGMs (BSTERGMs): STERGMs with the Bayesian setting, and I illustrate the computational algorithm that generates posterior samples from the model dealing with the doubly-intractable constant problem.
 Furthermore, I show the interpretability of the model in the Bayesian context.
@@
 \section{Introduction}
+Nowadays, new kinds of data having special forms emerge in many field.
+These new data are not confined by a series of values of real numbers as traditional data are in the past. 
+The data may have non-number type of value, special structure, or even integrated different type of values with model-specific relation.
+These data are called complex data, and an attention of statistical analysis on the data increases over time.
+
+As one form of the complex data, networks get highlighted attention among complex data for statistical analysis.
+Many phenomena can be represented by some relational structure, and the structure can be expressed effectively
+by a form of a network. For example, social networks including both offline and online can be thought as a sort of networks.
+In this view, some social aspects of relation between actors can be modeled with an object of social science
+including an analysis of power structure and a summary of intimacy structure on given group.
+Not only that, application to a transmitted disease leads a modeling using real data in epidemiology.
+
+Since statistical models on network provide not only lots of possibility of application but also
+interesting and useful features on observed network, statistical models have important role.
+The statistical model gives information about local relations like formation or dissolution between
+two actors as well as global features of the whole network such as probability that a specific structure
+realizes. Not only that, the statistical models investigate a complex behavior of network including
+how the global features of an observed network may be related to local network structure.
+During the investigation, statistical model can capture existing dependency on a given network
+and provide a tool to understand what certain structural features are more common on rare.
+
+Exponential-family random graph models(ERGMs) may be the most useful tools to analyze cross-sectional networks and their dependencies.
+\cite{RN117}, \cite{RN123}, \cite{RN124} and \cite{RN122} proposed the models and apply them to analyze dependent structures,
+and the Bayesian approaches of the models were started by \cite{RN115} and \cite{RN116}.
+In the ERGMs, The dependencies are captured using a series of local features called network statistics from classical graph-theory.
+Since these models play important role in this paper, I will introduce more precisely.
+
+Next natural question is a way to model a network sequence, or the evolution of network over time.
+For instance, one may want to model social network structures for 1 year, or
+simulation of a disease infection based on the sequential observations in the past.
+The first exploration applying ERGMs was \cite{RN109}. 
+After them, a notable contributions were of \cite{RN125} and \cite{RN126}, the Temporal ERGMs(TERGMs).
+In the TERGM, ERGMs are used by modelling the transition probability between times.
+Moreover, \cite{RN93} extend the TERGMs to the Separable Temporal ERGMs(STERGMs), considering an
+separation between duration - how long ties tend to last - and incidence - the rate at which new ties are formed -
+from prevalence - for example, the total number of ties - of the network properties.
+Imposing a conditional independence condition between the duration and incidence,
+the models can be extended effectively using similar fitting methods with TERGMs
+without adding much complexity.
+
+Although the STERGMs give us strength when it comes to interpret the fitted parameter,
+a Bayesian approach for the STERGMs has not been conducted. It is partly because of a hardship
+coming from the huge computation in a parameter estimation method for the model.
+Nonetheless, the Bayesian approach of STERGMs provides the statistical inference method
+following the standard Bayesian way, so the approach has huge advantage of convenience
+for statistical works for fitted parameter and predictions to next time network structure.
+Thus, in this paper, I specify the bayesian separable temporal exponential random graph models(BSTERGMs)
+and propose a fitting algorithm for the model parameter. 
+I also discuss ways of inference, the diagnostics procedure, and checking the goodness of fit.
+
+In chapter 2, I review the background meterials including ERGMs, BERGMs, TERGMs, and STERGMs,
+and in chapter 3, I show the detail setting of BSTERGMs. In chapter 4, I propose the fitting method for BSTERGM.
+In chapter 5, I explain ways to conduct diagnostics, statistical inference, and goodness of fit check.
+In chapter 6, I illustrate an application of BSTERGMs to network data of scholars and faculty
+from the UCLA medical working project team. Finally, I discuss some extension and 
+ways to supplement the BSTERGM further for future research.
 
@@
 
-\subsection{Exponential random graph model: ERGM}
+\subsection{ERGMs: Exponential random graph models}
 From the definition of random graph, one of the natural questions is 
 how we can model a distribution on $\mathcal{Y}$.
-As a answer, the Exponential random graph models (ERGM) (Wasserman and Pattison(1996) and Robins et al.(2007)) %ref
+As a answer, the Exponential random graph models (ERGMs) (\cite{RN123} and \cite{RN122})
 gives a classical but 
 flexible scheme to model for time-invariant networks represented as a single directed or undirected graph.
-Moreover, ERGM provide a statistical approach to the systematic exploration of several local features simultaneously and 
+Moreover, ERGMs provide a statistical approach to the systematic exploration of several local features simultaneously and 
 how these local structures interact to form a global network.
@@
 and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
-Models which have such a form called the ERGM.
+Models which have such a form called the ERGMs.
 
@@
 because the size of $\mathcal{Y}$ grows exponentially as $n$ grows.
-This situation is said to doubly-intractable constant problem, and it causes serious challenge to estimate the parameter vector $\theta$
-in real data analysis. For more detail on fitting algorithm using MCMCMLE, see Hunter and Handcock(2006).% ref
+Thus in real data analysis, the calculation of $c(\theta)$ should be avoided.
+For more detail of fitting algorithm, see \cite{RN104} the detail on fitting algorithm using MCMCMLE.
+You can also see \cite{RN100} addressing the ergm package in R with the MCMCMLE or the classic method using psudolikelihood.
 
-\subsection{BERGM: Bayesian ERGM}
-Caimo and Friel(2012) %ref
-extend ERGM to bayesian setting. Putting prior $p(\theta)$ to the parameter $\theta$, the model becomes
+\subsection{BERGMs: Bayesian ERGMs}
+\cite{RN115} and \cite{RN116} extend ERGMs to bayesian setting. Putting prior $p(\theta)$ to the parameter $\theta$, the model becomes
 \[p(\theta|y)=\frac{p(y|\theta)p(\theta)}{p(y)}=\frac{exp(\theta^T s(y))}{c(\theta)}\frac{p(\theta)}{p(y)}\]
-Since not only the doubly-intractable constant problem still exists but also the $p(y)$ cannot be evaluated directly,
-the estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
-The solution of Caimo and Friel is using the exchange algorithm.
 
-%exchange algorithm 설명
+Observe that not only $c(\theta)$ but also the $p(y)$ cannot be evaluated directly
+because of the difficulty of summation when $n$ grows and of integration of $p(y)$.
+Moreover, the constant depends on the parameter $\theta$, so ordinary MCMC technique cannot be used.
+This situation is said to doubly-intractable constant problem(\cite{RN119}), and causes serious challenge to estimate the parameter vector $\theta$.
+Thus, The estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
+The solution of Caimo and Friel is using the exchange algorithm. See \cite{RN115} for details of fitting methods of BERGMs
+and \cite{RN127} for the exchange algorithm.
 
-\subsection{TERGM: Temporal Exponential family Random Graphs Models}
+\subsection{TERGMs: Temporal Exponential family Random Graphs Models}
 Let's consider a sequence of random graph $\{Y_1, Y_2, ..., Y_T\} (T\in\mathbb{N})$ and assume that
 we are interested in the dynamics or evolution of the network sequence.
-The temporal Temporal Exponential family Random Graphs Models(TERGM)
-proposed by Hanneke, Fu and Xing(2009)%ref
+The temporal Temporal Exponential family Random Graphs Models(TERGMs)
+proposed by \cite{RN125} and \cite{RN126}
 is a model for our interest if we permit a Markov assumption
@@
 and a normalizing constant $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$.
-Models which have a such form called the TERGM.
+Models which have a such form called the TERGMs.
 
@@
 but the estimation suffers severe degeneracy problem.
-Furthermore, When we interpret the estimated TERGM parameters,
+Furthermore, When we interpret the estimated TERGMs parameters,
 we cannot separate the generation and duration because
@@
 
-\subsection{STERGM: Separable-Temporal Exponential family Random Graphs Models}
-\begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
-    Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-    Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
-    Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) by following way:
+\subsection{STERGMs: Separable-Temporal Exponential family Random Graphs Models}
+One solution to remedy the problems of TERGMs is 
+the Separable-Temporal Exponential family Random Graphs Models, or simply STERGMs. %ref
+The key idea is to separate the formation models and dissolution model
+for decoupling an effect on edge's duration of model parameter from edge's incidence
+when setting a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds, $T\in\mathbb{N}$),
+where $\mathcal{Y}$ is a set of graphs with given $n$ nodes.
 
+Precisely,
+Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (on $\mathcal{Y}$).
 Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
@@
 for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
-    and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1})), c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
+and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
+and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
-    Then, defining operations $+,-$ on $\mathcal{Y}$ following the boolean algebra edgewise-ly, set $y_t$ to
+Then, defining operations $+,-$ on $\mathcal{Y}$ following the edgewise boolean algebra,
+(For detail, Define binary operations on the edge values $\{0,1\}$ by $+$ and $-$ 
+by $1+1=1, 1+0=0+1=1, 0+0=0$ and $1-1=0, 1-0=1, 0-1=0, 0-0=0$) 
+set the combined network $y_t$ to
 \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
 
-    Additionally, assume that
-    \begin{itemize}
-        \item The first-order Markov assumption: $P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})$
-        \item The separability: the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$;
-            thus, \(P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+)P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\)
-    \end{itemize}
-    Models which have a such form called the STERGM.
-\end{defn}
+Additionally, assume the first-order Markov assumption
+\[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
+and the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$, called the separability.
+Then we get
+\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+Models which have a such form called the STERGMs.
 
-\subsection{Network Statistics for BSTERGM}
-As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
-A common candidates of $s'$ are:
-\begin{itemize}
-    \item the number of edges
-    \item node degree distribution (each order)
-    \item edgewise shared partner distribution (each order)
-    \item dyadwise shared partner distribution (each order)
-    \item k-star distribution
-    \item triangle distribution
-\end{itemize}
-In fact, a bundle of all order of node degree distribution and all order of edgewise shared partner distribution form sufficient statistics of a graph.
-Also, we can set $s'$ by a function of these statistics.
+\subsection{Network Statistics for ERGMs and derived models}
+Now, I briefly discuss the terms of network statistics.
+As a usual, $s(y)$ of ERGMs or $s(y_{t}^.,y_{t-1})$ of TERGMs and STERGMs would 
+reflect the structure of given network without identifying each node index.
+Common candidates of the statistics are here.
 
-\section{Bayesian STERGM}
+Let $y\in\mathcal{Y}$ and $n\in\mathbb{N}$ be the number of nodes of $y$.
+The simplest candidate is the number of edges, denoted by $|y|$.
+Following \cite{RN104}, common candidates are distributions defined by some sense of structure of given graph.
+For a given $i, 1\leq i \leq n-1$, $D_i(y)$ is defined to be a number of nodes in y whose the number of edges incident to the node equals $i$,
+and called $i$-th order node degree distribution. 
+For a given $i, 1\leq i \leq n-2$, $P_i(y)$ is defined to be a number of dyads $(j,k), 1\leq j,k \leq n$
+in y whose $j$ and $k$ have nodes connected by an edge each other and they share exactly $i$ connected nodes in common,
+and called $i$-th order edgewise shared partner distribution.
+Note that, the bundle of all order of node degree distribution and all order of edgewise shared partner distribution,
+\[(D_1(y),D_2(y),...,D_{n-1}(y),P_1(y),P_2(y),...,P_{n-2}(y))\] 
+forms sufficient statistics of a graph.
+And there are relation that
+\[n=D_0(y)+D_1(y)+...+D_{n-1}(y)\]
+and
+\[|y|=P_0(y)+P_1(y)+...+P_{n-2}(y)=\frac{1}{2}\sum_{i=1}^{n-1} iD_i(y)\]
 
-To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
-Then, the posterior of $\theta^+,\theta^-$ becomes
-\[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
-where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.
+A function of the sufficient statistics can be candidates for the model statistics.
+\cite{RN128} proposed graph statistics including $k$-star $S_k(y)$ for $1\leq k \leq n-1$ and $k$-triangles $T_k$ for $1\leq k \leq n-2$.
+The $k$-star consists of a node together with a set of $k$ of its connected nodes,
+and the $k$-triangle consists of $k$ triangles that share one common edge.
+They can be expressed by the definition of node degree distributions and edgewise shared partner distributions,
+\[S_1 = |y|, 
+S_k = \sum_{i=1}^{n-1} \binom{i}{k} D_i(y), k\geq 2\]
+and
+\[T_1 = \frac{1}{3}\sum_{i=1}^{n-2} iP_i(y),
+T_k = \sum_{i=1}^{n-2} \binom{i}{k} P_i(y), k\geq 2\]
 
-Remarks:
-\begin{itemize}
-    \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
-    \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
-        Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
-\end{itemize}
+Another interesting statistics are geometrically weighted value of distributions defined by
+\cite{RN118},\cite{RN104},\cite{RN128} and \cite{RN103}.
+For example, geometrically weighted (node) degree with parameter $\tau$ is defined by
+\[GWD(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)D_k(y)\]
+and geometrically weighted shared partner distribution with parameter $\tau$ is defined by
+\[GWESP(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)P_k(y)\]
 
-\section{Bayesian Estimation}
+In the context of temporal models, parameters in the model take two network as a arguments like $s(y_t,y_{t-1})$.
+One of common choices for these statistics is the bundle of sufficient statistics of first argument $y_t$.
+Another common choice is using a difference, $s(y_t,y_{t-1})=s'(y_t)-s'(y_{t-1})$ where $s'$ is a network statistics of given network.
+Although there are lots of other statistics related to ERGMs, I omit them and end this subsection here.
+If you need, see \cite{RN107} for the ERGMs, and \cite{RN125}'s chapter 2.1 for temporal the ERGMs.
 
-Suppose that a sequence of graph samples $y_1,...,y_T$ is observed
-\\ (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined,) 
-\\ and we want to fit the observation using BSTERGM.
+\section{Bayesian STERGMs}
+A goal is to convert the STERGMs to the Bayesian setting.
+With priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$,
+the Bayesian model for the STERGMs can be made. I call this model Bayesian Separable Temporal Exponential family Random Graph Models,
+or the BSTERGMs. More explicitly, for given the number of nodes $n$, let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (of $\mathcal{Y}$).
+As the STERGMs, define $\mathcal{Y}^+|_t$, $\mathcal{Y}^-|_t$ be subsets of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$,
+and which have equal or sparse edges comparing to $y_{t-1}$ respectively.
+Then, for $Y_t^+$ of $\mathcal{Y}^+|_t$ and $Y_t^-$ of $\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
+\[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+\[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
+for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
+and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
+and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
-How do we find the posterior distribution of $\theta^+,\theta^-$?
+Then, with the edgewise boolean algebra, set the combined network $y_t$ to
+\[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
 
-Note that an ordinary MCMC algorithm does not work in our situation, because the constant part remains when calculating the MCMC ratio.
-To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
+and assume that the first-order Markov property
+\[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
+and the conditional independence(separability) between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$.
+Then we get
+\[\theta^+ \sim p(\theta^+), \theta^- \sim p(\theta^-)\]
+\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
 
-Here is the algorithm of our main chain for the exchange algorithm.
+The purpose of Bayesian inference is to get information of the posterior distribution of $\theta^+,\theta^-$.
+By the Bayes rule, the posterior can be expressed by
+\[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})p(y_t, y_{t-1})} \]
+where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$ as set above.
+
+Note that, the BSTERGMs also have similar kinds of problems of the BERGMs.
+The normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically cannot be computed because we need to sum up too many terms.
+The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
+Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample in this model,
+because the constant part remains when calculating the MCMC ratio.
+
+\section{Estimation}
+
+Suppose that a sequence of graph samples $y_1,...,y_T$ is observed.
+(then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined.) 
+Then how can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using the BSTERGMs?
+To solve the doubly-intractable constant problem, I propose an novel algorithm using the exchange algorithm,
+one of the special MCMC techniques by \cite{RN127}, extending the algorithm of the BERGMs'(\cite{RN115}).
+Although MCMC types of approach cannot give an analytic expression of posterior distribution,
+it gives posterior samples as much as we want, and using the samples we can conduct statistical inference on given data.
+
+The key idea is running two MCMC chain to make independent sample to cancel the constant parts
+when calculating the MCMC accept-reject ratio.
+As a motivation, consider the ordinary MCMC ratio for $m$-th MCMC iteration in the BSTERGMs context 
+to show doubly intractable constant problem explicitly.
+
+With symmetric proposal distribution,
+\[\pi_{ordinary} = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)}
+{P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+where $\theta^*$ is proposed parameter in an iteration of MCMC.
+Plugging specific model form to above equation,
+\[\pi_{ordinary} = \frac{\frac{exp(\theta_*^+ s(y_t^+,y_{t-1}))}{c(\theta_*^+, y_{t-1})} \frac{exp(\theta_*^- s(y_t^-, y_{t-1}))}{c(\theta_*^-, y_{t-1})} p(\theta_*^+)p(\theta_*^-)}
+    {\frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))}{c(\theta_{m-1}^+, y_{t-1})} \frac{exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}{c(\theta_{m-1}^-,y_{t-1})} p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+Since each c depends on unique subscript of $\theta$ respectively, the constants are not canceled out.
+
+If we can generate simulated samples $y_{ex}|y_{t-1},\theta_{m-1}$ and $y_{ex}|y_{t-1},\theta_*$ and
+calculate their probability following the BSTERGMs, this situation can be resolved.
+At this point, applying the key idea mentioned above, multiply the probability terms in order to exclude the constant terms.
+Then we get the exchange algorithm's ratio,
+\[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}
+{P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
+
+\[=\frac{\frac{exp(\theta_*^+ s(y_t^+,y_{t-1}))}{c(\theta_*^+, y_{t-1})} \frac{exp(\theta_*^- s(y_t^-, y_{t-1}))}{c(\theta_*^-, y_{t-1})} p(\theta_*^+)p(\theta_*^-)}
+{\frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))}{c(\theta_{m-1}^+, y_{t-1})} \frac{exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}{c(\theta_{m-1}^-,y_{t-1})} p(\theta_{m-1}^+)p(\theta_{m-1}^-)}
+\frac{\frac{exp(\theta_{m-1}^+)s(y_{ex}^+,y_{t-1})}{c(\theta_{m-1}^+, y_{t-1})}\frac{exp(\theta_{m-1}^- s(y_{ex}^-,y_{t-1}))}{c(\theta_{m-1}^-, y_{t-1})}}
+{\frac{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))}{c(\theta_*^+,y_{t-1})} \frac{exp(\theta_*^- s(y_{ex}^-,y_{t-1}))}{c(\theta_*^-,y_{t-1})}}\]
+
+\[=\frac{exp(\theta_*^+ s(y_t^+,y_{t-1})) exp(\theta_*^- s(y_t^-, y_{t-1}))}{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}
+\frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_{ex}^-,y_{t-1}))}{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))exp(\theta_*^- s(y_{ex}^-,y_{t-1}))}
+\frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+
+% \[=exp((\theta_*^+ - \theta_{m-1}^+)s(y_t^+,y_{t-1}) + (\theta_*^- - \theta_{m-1}^-)s(y_t^-,y_{t-1})
+% -(\theta_*^+ - \theta_{m-1}^+)s(y_{ex}^+,y_{t-1}) - (\theta_*^- - \theta_{m-1}^-)s(y_{ex}^-,y_{t-1}))
+% \frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+
+\[=exp((\theta_*^+ - \theta_{m-1}^+)(s(y_t^+,y_{t-1}) - s(y_{ex}^+,y_{t-1}))(\theta_*^- - \theta_{m-1}^-)(s(y_t^-,y_{t-1}) - s(y_{ex}^-,y_{t-1})))
+\frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+
+The last expression shows that all constants are canceled out.
+This finding leads us to the fitting algorithm for the BSTERGMs using the exchange algorithm.
+
 \begin{algo}[the main chain]
@@
 \begin{enumerate}
-    \item Propose candidates $\theta_*^+,\theta_*^-$ from $\epsilon(.|\theta_{m-1}^+,\theta_{m-1}^-)$.
+    \item Propose candidates $\theta_*^+,\theta_*^-$ from a proposal distribution $\epsilon(.|\theta_{m-1}^+,\theta_{m-1}^-)$.
     \item Select a lag $(t-1,t)$ randomly on $2 \leq t \leq T$.
     \item Generate an exchange graph $y_{ex} \in\mathcal{Y}|_t$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
+    \item Calculate network statistics $s(y_t^+,y_{t-1}), s(y_{ex}^+,y_{t-1}), s(y_t^-,y_{t-1}), s(y_{ex}^-,y_{t-1})$.
     \item Calculate the exchange MCMC ratio $\pi$ at the lag,
-        \[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+,\theta_*^-)}
-            {P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+,\theta_{m-1}^-)}
-            \frac{P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}{P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
+        \[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)\epsilon(\theta_{m-1}|\theta_*)}
+            {P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)\epsilon(\theta_*|\theta_{m-1})} \]
+        \[* \frac{P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}{P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
     \item With probability $min(\pi,1)$, accept the proposal and put $(\theta_m^+,\theta_m^-) = (\theta_*^+,\theta_*^-)$.\\
@@
 \end{algo}
-Observe that, the $\pi$ has no normalizing constant terms because they are canceled out by added exchange terms.
-\(log\pi = (\theta_*^+-\theta_{m-1}^+)(s(y_{t}^+,y_{t-1})-s(y_{ex}^+,y_{t-1}))
-+(\theta_*^- -\theta_{m-1}^-)(s(y_{t}^-,y_{t-1})-s(y_{ex}^-,y_{t-1}))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
-\( = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
-+(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\)
 
-We need one more thing. To generate the exchange graph $y_{ex}$ for time $t$ (of the second part of main chain), 
-we should have a generative algorithm at a given parameter points.
-However, we still do not know the normalizing constant. Thus, I use one more MCMC chain.
+If $s(.,.)$ and $\theta$ can be multi-dimension (say, of $\mathbb{R}^p$,) replace the multiplication of the exponent parts to dot product between two vectors.
+In addition, when implementing the algorithm, I strongly recommend to use $log\pi$ instead of $\pi$ to avoid an underflow problem.
+\[log\pi = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
++(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\]
+
+To generate the exchange graph $y_{ex}$ for time $t$ at the second part of main chain,
+a generative algorithm at a given parameter points is needed.
+However, since the normalizing constant is still unknown, we cannot generate the exchange algorithm 
+directly using the model's expression of the distribution.
+Thus, I use one more MCMC chain to get the exchange samples without calculating the constants.
+
 \begin{algo}[the auxiliary chain]
@@
 \subsection{Model diagnostics}
-Since we run two kinds of MCMC chains, we should proceed two diagnostic task.
+The ERGMs and all derived models of the ERGMs are suffered degeneracy problem:
+all nodes of exchange sample become fully connected or totally isolated as the MCMC chain runs. 
+If the situation is worse, the problem cause an explosion to $\infty$ or $-\infty$ of the fitted parameters.
+The BSTERGMs are not exception. In other words, the diagnostic process is very important part for the BSTERGMs.
 
-The main chain produces the posterior samples of parameter, so the procedure is same as ordinary MCMC case.
-\begin{itemize}
-    \item Cut burn-in period. Do thinning if it is needed.
-    \item Depict traceplots of each parameter chain to check the convergence and the mixing.
-    \item Depict autocorrelation plot. Calculate ESS if it is needed.
-\end{itemize}
+Since two MCMC chains are run in the fitting algorithm of the BSTERGMs, two diagnostic tasks should be proceeded.
 
-Next, checking all auxiliary chains is practically irritating one. 
-In general, it is suffice to check the auxiliary chain of the last iteration (of the main chain)
-with statistics included in the model.
+The main chain produces the posterior samples of parameter
+$\{\theta_i^+\}_{i=1}^M, \{\theta_i^-\}_{i=1}^M$ in $\mathbb{R}^p, p\in\mathbb{N}$,
+which are not different with running ordinary MCMC algorithm.
+Depict traceplots of each parameter chain to check the convergence and the mixing.
+If needed, cut samples of burn-in period and do thinning.
+Plotting sample autocorrelation for each lag is also important. Calculate ESS if it is needed.
+
+However, diagnostics of auxiliary chains is challenging.
+One reason of the difficulty is that there are $M$ auxiliary chains in our whole procedure.
+Practically, it is not possible to all the chains as the number of iteration $M$ of main chain is large.
+Thus here is a guideline as a compromise: it suffices to check only one auxiliary chain of the last iteration of the main chain.
+Second reason comes from the outputs' form of the chains. They are not vectors in $\mathbb{R}^p$ but network structures. 
+Since we cannot draw traceplots of these network sequences, we need other ways for the diagnostics on them.
+One way is calculateing network statistics $s(.,.)$ in the model and drawing traceplots of them.
+On the traceplots, check whether each network statistics converges or not.
+In summary, for auxiliary chain,
 \begin{itemize}
@@
 \end{itemize}
+If the network statistics do not converges, raise the number of iteration $K$ of each auxiliary chain and
+run the fitting algorithm again.
+
+Here are some tips for remedy when facing a problem.
+If parameters explode to infinity the number of iteration of main chain becomes large,
+check whether the number of iteration of exchange chain is too small.
+Moreover, take account of the model itself. 
+If you put linearly dependent network statistics simultaneously in the model,
+the parameter may not be identifiable. If you set the model with network statistics of hugh different scale but
+similar MCMC proposal distribution, the MCMC chain tends to fail to converge. For example,
+if you put k-star of specific order and edge count to the model at the same time but use
+same proposal like $N(0,\sigma^2)$ for some $\sigma^2>0$, MCMC chain does not converge easily
+because the parameter of edge count should be changed by relatively wide range even if the parameter of k-star moves slightly.
+Finally, the convergence of the BSTERGMs is very sensitive to initial point of main chain.
+If the main chain starts at the region of degeneracy in the parameter space, it is a hopeless one to converge the chain properly.
 
@@
 
-Since we have posterior samples by running the main chain as many as we want, we can do basic inference procedure.
+Following standard inference methods of the Bayesian statistics,
+we can conduct the statistical inference on the parameters of the BSTERGMs.
+From the posterior samples generated by the exchange algorithm for the BSTERGMs, we can get
 \begin{itemize}
     \item A outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
-    \item An approximated summary statistics: mean, mode, variance, ...
-    \item An approximated quantile and probability interval
+    \item An approximated summary statistics: mean, mode, variance and so on.
+    \item An approximated $p$-th quantile and probability interval
 \end{itemize}
 
-Moreover, we already have a generative algorithm at the specific parameter point,
-we can predict the form of network at $T+1,T+2,...$ using posterior sample following standard Bayesian method.
-For example, for predicting $T+1$,
+Moreover, we already have a generative algorithm for a network at the specific parameter point.
+In fact, it is same to the algorithm used in the auxiliary chain.
+Thus, we can predict the form of network at $T+1,T+2,...$ using posterior sample following the standard Bayesian prediction method.
+For example, for predicting networks structure at $T+1$,
 \begin{itemize}
@@
 \end{itemize}
-If you need, calculate some network statistics for the results and get a summary statistics of them.
+If you need, calculate some network statistics of the generated networks and get some summary statistics on them.
 
 \subsection{Goodness of Fit}
-To evaluate the goodness of fit for the posterior $\theta^+,\theta^-$,
+To evaluate the goodness of fit for the BSTERGMs and to make sure the posterior $\theta^+,\theta^-$ is fitted properly,
 we can use the auxiliary chain algorithm once again.
 \begin{algo}[GOF procedure]
-    For t=2,...,T
-    \\For s=1,...,S
+    For $t=2,...,T$ \\
+    and For $j=1,...,J$
     \begin{enumerate}
-        \item sample $\theta_s^+,\theta_s^-$ from the estimate of posterior.
-        \item simulate $y_s$ using the auxiliary chain under $y_{t-1}$.
-        \item calculate $g(y_s)$, some higher degree statistics (eg. Node-degree dist \& Edgewise Shared Partner dist)
+        \item sample $\theta_j^+,\theta_j^-$ from the estimate of posterior.
+        \item simulate $y_j$ using the auxiliary chain under $y_{t-1}$.
+        \item calculate $s(y_j, y_{t-1})$, some higher degree statistics (eg. Node-degree dist \& Edgewise Shared Partner dist)
     \end{enumerate}
-    Draw the box-plot of $g(y_s)$ and compare with $g(y_t)$.
+    Take $p$-th and $(1-p)$-th quantiles of each element of $s(y_j, y_{t-1})$s respectively and check 
+    whether the network statistics of observed data is covered by the range of te quantiles or not.
+    In addition, draw the box-plot of calculated $s(y_j, y_{t-1})$ and add points or a line with the network statistics of observed data.
 \end{algo}
 
-\section{Example: Scholars and faculty members' experiences
-in online social networks}
+\section{Example: UCLA medical working project - team science data}
+%data description
+As an application, I consider the Bayesian separable temporal exponential family model with UCLA medical working project's team science data.
+These data consist of three sequence,
+and each sequence has three time points and for each time,
+there is a network with 52 nodes corresponding 29 scholar members and 23 faculty members.
+
+% \begin{figure}
+%     \begin{center}
+%     \includegraphics[scale=0.5]{LS_ggplot_2.png}
+%     \caption{Latent positions of politicians and bills.}
+%     \label{fig:ls}
+%     \end{center}
+% \end{figure}
+
+I consider 2 dimentional model as follows.
+\[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+\[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
+and
+\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+where
+\(s(z_1,z_2)^T=(|z_1|-|z_2|)\)
+for $z_1,z_2\in\mathcal{Y}$ with 52 nodes. In other words, the network statistic in the model is
+difference of the number of edges for both the formation model and the disolution model.
+
+I use the ordinary normal prior $p(\theta^+)~\mathcal{N}(0,100)$ and $p(\theta^-)~\mathcal{N}(0,100)$,
+and set the proposal distribution of main chain $\epsilon~\mathcal{N}(0,0.02)$.
+The auxiliary chains consist of 1000 iterations and the main chain consists of 50000 iteration.
+After generating the chain, I cut first 10000 samples for a burn-in period and thin the sample by 10.
+
+The posterior estimates for each network and histograms are follows.
+%mean, sd
+%hist
+
+For the diagnostics, I attach traceplots and autocorrelation plots.
+%traceplot
+%acf
+
+In addition, here are traceplots of the number of edges in last auxiliary chain.
+They shows that all chains are not degenerated at last parameter point.
+%lastsampler
+
+Finally, I append the results of goodness of fit.
+%GOF table
+
 \section{Summary and Discussion}
+In this paper, I have proposed the Bayesian STERGMs for a sequence of networks
+and a framework for statistical estimation and inference using the standard Bayesian approach for the model.
+I also have shown a application of the BSTERGMs. %more add
+
+Since the models are based on ERGMs, it can be extended to more general networks
+including a variety of network statistics.
+Moreover, the model can be applied to weighted graph and use the weights as covariates of the model.
+Plus, the BSTERGMs are using the exchange algorithm, so it is flexible enough to be integrated
+with other methods to help the MCMC chains to converge and run faster.
+For example, the 'tie no tie'(TNT) algorithm used by ergm package (\cite{RN100}) can be applied to
+enhance the performance of the auxiliary chain of the BSTERGMs.
+
+A modeling by the BSTERGMs for large networks is possible,
+but a problem of computational inefficiency rises.
+If given observed network has lots of nodes, the estimation algorithm for the BSTERGM needs 
+huge number of auxiliary chain itertation to sample a proper exchange network at a given parameter point.
+In addition, the calculation of network statistics in the model cause another problem for large networks,
+especially when the model has network statistics that high complexity of computation.
+For example, statistics based on edgewise shared partner distributions needs $O(n^3)$,
+so computation cost increases rapidly as the number of nodes $n$ grows and
+the auxiliary chains become bottlenecks of the whole computation procedure.
+
+Another important issue related in the BSTERGMs is degeneracy as I have mentioned at the section of model diagnostics.
+The BSTERGMs are very sensitive to the initial point of main MCMC chain, so
+searching a region for non-degenerate initial parameter point in the parameter space is very important.
+Often, however, the work finding the region is too difficult for the BSTERGMs,
+partly because of the high number of parameters in the models.
+Since the BSTERGMs has two parts for $\theta^-, \theta^+$ in the setting of the models by the their definition,
+the dimension of parameter space increases easily although each formation and dissolution model does not have
+too many terms respectively.
+What makes the situation worse is not only the high dimension of parameter space but also the shape of convergent region of the parameter.
+In many case, the region is geometrically curved due to an implicit relation between several network statistics in the models
+and interaction between the formation model and the dissolution model.
+Thus, the MCMC algorithm may fail to exploring the whole parameter space, especially when the space has high dimension,
+and suffer non-convergence problem.
+Imposing some model level constraint or using special prior may be a remedy,
+but finding the proper constraint or the prior distribution will often be a more difficult problem than original one.
+Thus, further research is needed to resolve these issues.
 
@@
 
-testcite \cite{RN100}
-
 \nocite{*}
-\bibliographystyle{plain}
+\bibliographystyle{apalike}
 \bibliography{doc.bib}

### 2020-12-10 doc/doc.tex
tex1211-2

@@
 for $z_1,z_2\in\mathcal{Y}$ with 52 nodes. In other words, the network statistic in the model is
-difference of the number of edges for both the formation model and the disolution model.
+difference of the number of edges for both the formation model and the dissolution model.
 
-I use the ordinary normal prior $p(\theta^+)~\mathcal{N}(0,100)$ and $p(\theta^-)~\mathcal{N}(0,100)$,
-and set the proposal distribution of main chain $\epsilon~\mathcal{N}(0,0.02)$.
+I use the ordinary normal prior $p(\theta^+)\sim\mathcal{N}(0,100)$ and $p(\theta^-)\sim\mathcal{N}(0,100)$,
+and set the proposal distribution of main chain $\epsilon\sim\mathcal{N}(0,0.0004)$.
 The auxiliary chains consist of 1000 iterations and the main chain consists of 50000 iteration.
+In this setting, the BSTERGM fitting algorithm run 250 iteration of main chain per a minute on the intel i7-8700(3.20GHz) CPU.
+
 After generating the chain, I cut first 10000 samples for a burn-in period and thin the sample by 10.
-
 The posterior estimates for each network and histograms are follows.

### 2020-12-10 BayesianSTERGM/NetMCMCSampler_STERGM.h
candidate and pics

@@ -54,7 +54,7 @@ private:
     double log_r(Network& last, Network& proposed, bool isDissolution) {
         //NOW MODEL
         Col<double> model_delta = { (double)proposed.get_n_Edge() - last.get_n_Edge(),
-                                     (double)proposed.get_undirected_geoWeightedESP(0.5) - last.get_undirected_geoWeightedESP (0.5) 
+                                     // (double)proposed.get_undirected_geoWeightedESP(0.5) - last.get_undirected_geoWeightedESP (0.5) 
         }; // <-model specify
         Col<double> param;
         if (isDissolution) {

### 2020-12-10 BayesianSTERGM/data/make_network_pic.R

@@ -1,0 +1,14 @@
+library(statnet)
+
+name_vec = c("f1.19", "f2.19", "f3.19", "m1.19", "m2.19", "m3.19", "w1.19", "w2.19", "w3.19")
+
+
+for(i in 1:length(name_vec)){
+    name = name_vec[i]
+    net.structure = read.table(paste("BayesianSTERGM/data/",name,".network.txt",sep=''), sep=",")
+    net = as.network(as.matrix(net.structure), directed=FALSE)
+    plot(net)
+    filename = paste("BayesianSTERGM/data/",name,"_nework.png",sep="")
+    dev.copy(png,filename=filename)
+    dev.off()
+}

### 2020-12-10 BayesianSTERGM/main.cpp

@@ -669,11 +669,11 @@ int main()
     //=================================================================================================
     //BSTERGM test
     
-    Col<double> initFormationParam = { -5, 1.5 };
-    Col<double> initDissolutionParam = { -5, 5 };
+    Col<double> initFormationParam = { -0.9 };
+    Col<double> initDissolutionParam = { -0.3 };
     BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_3_seq);
     //아래 GOF seq도 바꿀 것!
-    Bstergm.generateSample(20000, 700, 0.002);
+    Bstergm.generateSample(50000, 10000, 0.0004); //50000,1000,0.0004
     //Bstergm.cutBurnIn(1000);
     //Bstergm.thinning();
 
@@ -683,17 +683,17 @@ int main()
     Col<double> quantilePts = { 0.1, 0.25, 0.5, 0.75, 0.9 };
     BstergmDiag1.print_quantile(0, quantilePts);
     BstergmDiag1.print_autoCorr(0, 30);
-    BstergmDiag1.print_mean(1);
+    /*BstergmDiag1.print_mean(1);
     BstergmDiag1.print_quantile(1, quantilePts);
-    BstergmDiag1.print_autoCorr(1, 30);
+    BstergmDiag1.print_autoCorr(1, 30);*/
 
     Diagnostics_MCParamSample BstergmDiag2(Bstergm.getPosteriorSample_dissolution());
     BstergmDiag2.print_mean(0);
     BstergmDiag2.print_quantile(0, quantilePts);
     BstergmDiag2.print_autoCorr(0, 30);
-    BstergmDiag2.print_mean(1);
+    /*BstergmDiag2.print_mean(1);
     BstergmDiag2.print_quantile(1, quantilePts);
-    BstergmDiag2.print_autoCorr(1, 30);
+    BstergmDiag2.print_autoCorr(1, 30);*/
 
     BstergmDiag1.writeToCsv_Sample("BSTERGM_formation.csv");
     BstergmDiag2.writeToCsv_Sample("BSTERGM_dissolution.csv");
@@ -708,6 +708,7 @@ int main()
     cout << "\n\nt=1 to t=2" << endl;
     BstergmGoF1.undirected_run(1, 100, 100);
     BstergmGoF1.undirected_printResult(1);
+
 
     //BERGM LAST Exchange Sampler diag
[…]

### 2020-12-10 doc/doc.tex

--- 2020-12-06 doc/doc.tex
@@
 because the size of $\mathcal{Y}$ grows exponentially as $n$ grows.
-This situation is said to doubly-intractable constant problem, and it causes serious challenge to estimate the parameter vector $\theta$
-in real data analysis. For more detail on fitting algorithm using MCMCMLE, see Hunter and Handcock(2006).% ref
+Thus in real data analysis, the calculation of $c(\theta)$ should be avoided.
+For example, see the detail on fitting algorithm using MCMCMLE. %see Hunter and Handcock(2006).% ref
 
@@
 \[p(\theta|y)=\frac{p(y|\theta)p(\theta)}{p(y)}=\frac{exp(\theta^T s(y))}{c(\theta)}\frac{p(\theta)}{p(y)}\]
-Since not only the doubly-intractable constant problem still exists but also the $p(y)$ cannot be evaluated directly,
-the estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
-The solution of Caimo and Friel is using the exchange algorithm.
+
+Observe that not only $c(\theta)$ but also the $p(y)$ cannot be evaluated directly
+because of the difficulty of summation when $n$ grows and of integration of $p(y)$.
+Moreover, the constant depends on the parameter $\theta$, so ordinary MCMC technique cannot be used and
+This situation is said to doubly-intractable constant problem, and causes serious challenge to estimate the parameter vector $\theta$..
+Thus, The estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
+The solution of Caimo and Friel is using the exchange algorithm. See %ref
 
@@
 \subsection{STERGM: Separable-Temporal Exponential family Random Graphs Models}
-\begin{defn}[STERGM: Separable-Temporal Exponential family Random Graphs Models]
-    Let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-    Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs (on $\mathcal{Y}$).
-    Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) by following way:
+One solution to remedy the problems of TERGM is 
+the Separable-Temporal Exponential family Random Graphs Models, or simply STERGM. %ref
+The key idea is to separate the formation models and dissolution model
+for decoupling an effect on edge's duration of model parameter from edge's incidence
+when setting a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds, $T\in\mathbb{N}$),
+where $\mathcal{Y}$ is a set of graphs with given $n$ nodes.
 
+Precisely,
+Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (on $\mathcal{Y}$).
 Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
@@
 for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
-    and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1})), c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
+and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
+and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
-    Then, defining operations $+,-$ on $\mathcal{Y}$ following the boolean algebra edgewise-ly, set $y_t$ to
+Then, defining operations $+,-$ on $\mathcal{Y}$ following the edgewise boolean algebra,
+(For detail, Define binary operations on the edge values $\{0,1\}$ by $+$ and $-$ 
+by $1+1=1, 1+0=0+1=1, 0+0=0$ and $1-1=0, 1-0=1, 0-1=0, 0-0=0$) 
+set the combined network $y_t$ to
 \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
 
-    Additionally, assume that
-    \begin{itemize}
-        \item The first-order Markov assumption: $P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})$
-        \item The separability: the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$;
-            thus, \(P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+)P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\)
-    \end{itemize}
+Additionally, assume the first-order Markov assumption
+\[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
+and the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$, called the separability.
+Then we get
+\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
 Models which have a such form called the STERGM.
-\end{defn}
 
-\subsection{Network Statistics for BSTERGM}
-As a usual, $s(y_{t}^.,y_{t-1})$ would be chosen by a difference of network statistics, $s'(y_t^.)-s'(y_{t-1})$.
-A common candidates of $s'$ are:
-\begin{itemize}
-    \item the number of edges
-    \item node degree distribution (each order)
-    \item edgewise shared partner distribution (each order)
-    \item dyadwise shared partner distribution (each order)
-    \item k-star distribution
-    \item triangle distribution
-\end{itemize}
-In fact, a bundle of all order of node degree distribution and all order of edgewise shared partner distribution form sufficient statistics of a graph.
-Also, we can set $s'$ by a function of these statistics.
+\subsection{Network Statistics for ERGM and its derivative models}
+Now, I briefly discuss the terms of network statistics.
+As a usual, $s(y)$ of ERGM or $s(y_{t}^.,y_{t-1})$ of TERGM and STERGM would 
+reflect the structure of given network without identifying each node index.
+Common candidates of the statistics are here.
+
+Let $y\in\mathcal{Y}$ and $n\in\mathbb{N}$ be the number of nodes of $y$.
+The simplest candidate is the number of edges, denoted by $|y|$.
+Another common candidates are distributions defined by some sense of structure of given graph.
+For a given $i, 1\leq i \leq n-1$, $D_i(y)$ is defined to be a number of nodes in y whose the number of edges incident to the node equals $i$,
+and called $i$-th order node degree distribution. 
+%David Hunter, Mark Handcock 2006 %ref
+For a given $i, 1\leq i \leq n-2$, $P_i(y)$ is defined to be a number of dyads $(j,k), 1\leq j,k \leq n$
+in y whose $j$ and $k$ have nodes connected by an edge each other and they share exactly $i$ connected nodes in common,
+and called $i$-th order edgewise shared partner distribution.
+Note that, the bundle of all order of node degree distribution and all order of edgewise shared partner distribution,
+\[(D_1(y),D_2(y),...,D_{n-1}(y),P_1(y),P_2(y),...,P_{n-2}(y))\] 
+forms sufficient statistics of a graph.
+And there are relation that
+\[n=D_0(y)+D_1(y)+...+D_{n-1}(y)\]
+and
+\[|y|=P_0(y)+P_1(y)+...+P_{n-2}(y)=\frac{1}{2}\sum_{i=1}^{n-1} iD_i(y)\]
+
+A function of the sufficient statistics can be candidates for the model statistics.
+%Snijders et al 2006 %ref
+They proposed graph statistics including $k$-star $S_k(y)$ for $1\leq k \leq n-1$ and $k$-triangles $T_k$ for $1\leq k \leq n-2$.
+The $k$-star consists of a node together with a set of $k$ of its connected nodes,
+and the $k$-triangle consists of $k$ triangles that share one common edge.
+They can be expressed by the definition of node degree distributions and edgewise shared partner distributions,
+\[S_1 = |y|, 
+S_k = \sum_{i=1}^{n-1} \binom{i}{k} D_i(y), k\geq 2\]
+and
+\[T_1 = \frac{1}{3}\sum_{i=1}^{n-2} iP_i(y),
+T_k = \sum_{i=1}^{n-2} \binom{i}{k} P_i(y), k\geq 2\]
+
+Another interesting statistics are geometrically weighted value of distributions defined earlier.
+%ref needed. adaptive exchange [17][19][42]
+For example, geometrically weighted (node) degree with parameter $\tau$ is defined by
+\[GWD(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)D_k(y)\]
+and geometrically weighted shared partner distribution with parameter $\tau$ is defined by
+\[GWESP(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)P_k(y)\]
+
+In the context of temporal models, parameters in the model take two network as a arguments like $s(y_t,y_{t-1})$.
+One of common choices for these statistics is the bundle of sufficient statistics of first argument $y_t$, or
+Although there are lots of other statistics related to ERGM, I omit them and end this subsection here.
 
 \section{Bayesian STERGM}
+A goal is to convert the STERGM to the Bayesian setting.
+With priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$,
+the bayesian model for STERGM can be made. I call this model Bayesian Separable Temporal Exponential family Random Graph Model,
+or BSTERGM. More explicitly, for given the number of nodes $n$, let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (of $\mathcal{Y}$).
+As STERGM, define $\mathcal{Y}^+|_t$, $\mathcal{Y}^-|_t$ be subsets of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$,
+and which have equal or sparse edges comparing to $y_{t-1}$ respectively.
+Then, for $Y_t^+$ of $\mathcal{Y}^+|_t$ and $Y_t^-$ of $\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
+\[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+\[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
+for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
+and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
+and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
-To convert the STERGM to the Bayesian setting, put priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$ and take the Bayes theorem as a inference rule.
-Then, the posterior of $\theta^+,\theta^-$ becomes
-\[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})} \]
-where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$.
+Then, with the edgewise boolean algebra, set the combined network $y_t$ to
+\[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
 
-Remarks:
-\begin{itemize}
-    \item We cannot compute the normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically because we need to sum up too many terms.
-    \item The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
-        Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample.
-\end{itemize}
+and assume that the first-order Markov property
+\[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
+and the conditional independence(separability) between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$.
+Then we get
+\[\theta^+ \sim p(\theta^+), \theta^- \sim p(\theta^-)\]
+\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+
+The purpose of Bayesian inference is to get information of the posterior distribution of $\theta^+,\theta^-$.
+By the Bayes rule, the posterior can be expressed by
+\[P(\theta^+,\theta^-|y_t, y_{t-1}) = \frac{P(Y_t^+=y_t^+|y_{t-1},\theta^+) P(Y_t^-=y_t^-|y_{t-1},\theta^-)P(\theta^+),P(\theta^-)}{c(\theta^+,y_{t-1})c(\theta^-,y_{t-1})p(y_t|y_{t-1})} \]
+where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$ as set above.
+
+Note that, BSTERGM also have similar kinds of problems of BERGM.
+The normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically cannot be computed because we need to sum up too many terms.
+The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
+Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample in this model,
+because the constant part remains when calculating the MCMC ratio.
 
@@
 
-Suppose that a sequence of graph samples $y_1,...,y_T$ is observed
-\\ (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined,) 
-\\ and we want to fit the observation using BSTERGM.
+Suppose that a sequence of graph samples $y_1,...,y_T$ is observed.
+(then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined.) 
+Then how can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using BSTERGM?
+To solve this doubly-intractable constant problem, I propose an algorithm using the exchange MCMC algorithm,
+one of the special MCMC techniques.
 
-How do we find the posterior distribution of $\theta^+,\theta^-$?
+The idea is running two MCMC chain to make independent sample to cancel the constant parts
+when calculating the MCMC accept-reject ratio.
+To begin with, consider the MCMC ratio in the BSTERGM context.
+With symmetric proposal distribution,
+\[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)}
+{P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+where $\theta^*$ is proposed parameter in an iteration of MCMC.
+Plugging specific model form to above equation,
+\[\pi=\] %start 12/10 here!!
 
-Note that an ordinary MCMC algorithm does not work in our situation, because the constant part remains when calculating the MCMC ratio.
-To solve this doubly-intractable constant problem, we should use more special MCMC technique, the exchange MCMC algorithm.
-
-Here is the algorithm of our main chain for the exchange algorithm.
 \begin{algo}[the main chain]
@@
 \nocite{*}
-\bibliographystyle{plain}
+\bibliographystyle{apalike}
 \bibliography{doc.bib}

### 2020-12-10 paper_format/CSY.tex

@@
 \end{equation}
-
+\end{itemize}
 %--------------------------------------------------------------------------------%
@@
 \section{Estimated Latent Space}
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.5]{LS_ggplot_2.png}
-\caption{Latent positions of politicians and bills.}
-\label{fig:ls}
-\end{center}
-\end{figure}
 
@@
 As for politicians comparatively located in the mid-area, we can say that they did not follow their party blindly.
-
-\begin{figure}
-  \begin{center}
-    \subfigure[2016]{\label{fig:annual-a}\includegraphics[scale=0.25]{LS_2016.png}}
-    \subfigure[2018]{\label{fig:annual-b}\includegraphics[scale=0.25]{LS_2018.png}}
-    \subfigure[2020]{\label{fig:annual-c}\includegraphics[scale=0.25]{LS_2020.png}}
-  \end{center}
-  \caption{Latent positions in 2016, 2018, and 2020. }
-  \label{fig:annual}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{districts.png}
-  \end{center}
-  \caption{Congressional districts of U.S. }
-  \label{fig:districts}
-\end{figure}
-
-\quad In the analysis, we accept several beliefs in political tendency.
-First, Republican is conservative, and Democrat is progressive.
-Second, an urban region, including the big city, is more politically progressive.
-Third, each state excepting the swing state supports a certain political party.
-Taking it into consideration, the analysis is undertaken.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{mn.png}
-  \end{center}
-  \caption{Congressional districts of Minnesota }
-  \label{fig:mn}
-\end{figure}
-
-\quad Minnesota, having Minneapolis as a capital city, is a relatively progressive state.
-Among the districts of Minnesota, the 7th district covers almost all of the western side of Minnesota.
-In this region, the proportion of the rural area is larger than that of the urban area.
-Moreover, the district borders with North Dakota and South Dakota, which are conservative states.
-Considering this geographic characteristic, the 7th district is likely to be conservative.
-Therefore, it is natural that the district's representative, Collin Peterson, is a conservative Democrat, reflecting both the state's progressive personality and the conservative aspect of the district.
-Peterson is one of the founders of the Blue Dog Coalition, a group of moderate and conservative Democrats.
-Peterson has split from his party on issues such as gay marriage, healthcare, the estate tax, gun control, the environment, and abortion.
-It is no doubt that he successfully represented the district's opinion considering that he was first elected in 1990 and kept re-elected until 2018.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{il.png}
-  \end{center}
-  \caption{Congressional districts of Illinois}
-  \label{fig:il}
-\end{figure}
-
-\quad Illinois is known as one of the progressive states, placing the large city of Chicago.
-Unlike our anticipation, the Republican politician was elected in the district of the progressive state.
-Bob Dold, the representative for Illinois's 10th congressional district in 2016, is a moderate Republican who supports gun control measures.
-Additionally, his predecessor, Brad Schneider, is also known as moderate Democrat.
-From the characteristic of representatives, it is possible to say that 10th district of Illinois prefers a moderate politician.
-We can deduce the reasons for the political position of this district from geographic factors as well.
-10th district lies in the northeast corner of the state, and mostly comprises northern suburbs of Chicago.
-In this region, the district covers Lake county which is the second most wealthy county in Illinois.
-When it is believed that wealthy people tend to be more conservative in common sense, it is natural thinking that the 10th district is politically moderate despite it is near Chicago.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{nj.png}
-  \end{center}
-  \caption{Congressional districts of New Jersey }
-  \label{fig:nj}
-\end{figure}
-
-\quad New Jersey is also a progressive state influenced by an adjacent state, New York.
-However, the 2nd district elected Republican representative.
-The district, based in Southern New Jersey, is a swing region with a history of supporting different parties.
-The proportion of the rural area is relatively high, and there is no metropolis in this district.
-Moreover, it is located most far from New York and less influenced by New York City.
-For this reason, the 2nd district seems to be more conservative than the average of New Jersey.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{ok.png}
-  \end{center}
-  \caption{Congressional districts of Oklahoma }
-  \label{fig:ok}
-\end{figure}
-
-\quad Oklahoma is a state in the South Central region of the United States and it is a conservative state.
-However, the 5th district includes Oklahoma City, and it is likely to have a moderate political stance.
-The result of the election of Kendra Horn, the representative of Democrat, demonstrates this enough.
-The district's preference is obvious, seeing the fact that Horn was joining the Blue Dog Coalition, a group of moderate and conservative Democrats.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{sc.png}
-  \end{center}
-  \caption{Congressional districts of South Carolina }
-  \label{fig:sc}
-\end{figure}
-
-\quad South Carolina is known as a conservative state.
-Conversely, the 1st district elected the Democrat politician.
-The 1st district has historically been based in Charleston, the largest city in South Carolina.
-The urban characteristic may influence the political position.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.5]{utah.png}
-  \end{center}
-  \caption{Congressional districts of Utah }
-  \label{fig:utah}
-\end{figure}
-
-\quad Utah is famous as conservative state because a little more than half of all Utahns are Mormons.
-The 4th district representative, Ben McAdams is the only Democratic member of Utah's congressional delegation, but a member of the Blue Dog Coalition, a group of a moderate politician.
-Despite the Mormons' headquarter, Salt Lake City, the most populous city in Utah, tends to affect progressive in a political position.
-
-\quad Interpreting the patterns found above, the commonality can be seen that Bipartisan is elected in the district of a moderate political position compared with that of the whole state.
-Moreover, we infer that the degree of the district's political orientation varies with the geographical feature.
-For the districts of the progressive state, the district with the broad rural area or covering wealthy county tends to be more conservative than other districts in the state.
-As for the districts of the conservative state, the district containing a large city is more progressive than others.
-In addition, the political atmosphere of neighboring states affects the residents' political ideas of the district.
-
-\section{The Bills in the Middle Area}
-
-\begin{figure}[h]
-  \begin{center}
-    \subfigure[2016]{\label{fig:bill-a}\includegraphics[scale=0.18]{vote_2016.png}}
-    \subfigure[2018]{\label{fig:bill-b}\includegraphics[scale=0.18]{vote_2018.png}}
-    \subfigure[2020]{\label{fig:bill-c}\includegraphics[scale=0.18]{vote_2020.png}}
-  \end{center}
-  \caption{Latent positions of bills and their numbers in 2016, 2018, and 2020. }
-  \label{fig:bill}
-\end{figure}
-
-\section{Outliers}
-
-\end{itemize}
-
 %--------------------------------------------------------------------------------%
@@
 
-\section{Trace Plots}
-\subsection{$\theta_{t,k}$}
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[scale=0.4]{ts_theta_1.png}
-  \end{center}
-  \caption{Trace plot of $\theta_{t,1}$}
-  \label{fig:ts_theta}
-\end{figure}
+% %----------------------------------------------------------
+% \newpage
 
-%----------------------------------------------------------
-\newpage
+% \section*{국 문 요 약}
+% \setstretch{1.8}
+% \par
+% \bigskip
 
-\section*{국 문 요 약}
-\setstretch{1.8}
-\par
-\bigskip
+% \begin{center}
+% \large
+% 종단 잠재 공간 항목 반응 모형의 응용,\\
+% 미 하원 의회 투표 자료 분석을 중심으로
+% \end{center}
+% \setstretch{1.8}
+% \par
+% \bigskip
+% \noindent
 
-\begin{center}
-\large
-종단 잠재 공간 항목 반응 모형의 응용,\\
-미 하원 의회 투표 자료 분석을 중심으로
-\end{center}
-\setstretch{1.8}
-\par
-\bigskip
-\noindent
+% \bigskip
+% \noindent \hrulefill\\
+% 핵심 용어 :  \par
 
-\bigskip
-\noindent \hrulefill\\
-핵심 용어 :  \par
+% %\pagebreak
 
-%\pagebreak
+% \newpage
 
-\newpage
+% %--------------------------------------------------------------------------------%
+% %--------------------------------------------------------------------------------%
 
-%--------------------------------------------------------------------------------%
-%--------------------------------------------------------------------------------%
-
-%\nocite{*}
-\bibliographystyle{apalike}
-\bibliography{References}
+% %\nocite{*}
+% \bibliographystyle{apalike}
+% \bibliography{References}
 

### 2020-12-10 BayesianSTERGM/candidate_run/makePlots.R
net3seq

@@ -1,5 +1,5 @@
-namevec = c("net1seq_chain1", "net2seq_chain1")
-name = namevec[2]
+namevec = c("net1seq_chain1", "net2seq_chain1", "net3seq_chain1")
+name = namevec[3]
 
 # ==================================================================================
 

### 2020-12-10 BayesianSTERGM/main.cpp

@@ -673,7 +673,7 @@ int main()
     Col<double> initDissolutionParam = { -0.3 };
     BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_3_seq);
     //아래 GOF seq도 바꿀 것!
-    Bstergm.generateSample(50000, 10000, 0.0004); //50000,1000,0.0004
+    Bstergm.generateSample(50000, 1000, 0.0004); //50000,1000,0.0004
     //Bstergm.cutBurnIn(1000);
     //Bstergm.thinning();
 

### 2020-12-10 BayesianSTERGM/candidate_run/string_maker.py
gof res

@@ -1,0 +1,12 @@
+import csv
+
+
+result=[]
+with open('net3seq_chain1_aftercut_BSTERGM_formation.csv', newline='') as csvfile:
+# with open('net3seq_chain1_aftercut_BSTERGM_dissolution.csv', newline='') as csvfile:
+    csv_reader = csv.reader(csvfile, delimiter=' ', quotechar='|')
+    for row in csv_reader:
+        result.append(float(row[0]))
+
+print(result)
+print(len(result))

### 2020-12-10 BayesianSTERGM/main.cpp

@@ -667,54 +667,54 @@ int main()
     //Tsampler.printResult(3);
 
     //=================================================================================================
-    //BSTERGM test
+    ////BSTERGM test
+    //
+    //Col<double> initFormationParam = { -0.9 };
+    //Col<double> initDissolutionParam = { -0.3 };
+    //BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_3_seq);
+    ////아래 GOF seq도 바꿀 것!
+    //Bstergm.generateSample(50000, 1000, 0.0004); //50000,1000,0.0004
+    ////Bstergm.cutBurnIn(1000);
+    ////Bstergm.thinning();
+
+    ////BSTERGM posterior sample diagnostics
+    //Diagnostics_MCParamSample BstergmDiag1(Bstergm.getPosteriorSample_formation());
+    //BstergmDiag1.print_mean(0);
+    //Col<double> quantilePts = { 0.1, 0.25, 0.5, 0.75, 0.9 };
+    //BstergmDiag1.print_quantile(0, quantilePts);
+    //BstergmDiag1.print_autoCorr(0, 30);
+    ///*BstergmDiag1.print_mean(1);
+    //BstergmDiag1.print_quantile(1, quantilePts);
+    //BstergmDiag1.print_autoCorr(1, 30);*/
+
+    //Diagnostics_MCParamSample BstergmDiag2(Bstergm.getPosteriorSample_dissolution());
+    //BstergmDiag2.print_mean(0);
+    //BstergmDiag2.print_quantile(0, quantilePts);
+    //BstergmDiag2.print_autoCorr(0, 30);
+    ///*BstergmDiag2.print_mean(1);
+    //BstergmDiag2.print_quantile(1, quantilePts);
+    //BstergmDiag2.print_autoCorr(1, 30);*/
+
+    //BstergmDiag1.writeToCsv_Sample("BSTERGM_formation.csv");
+    //BstergmDiag2.writeToCsv_Sample("BSTERGM_dissolution.csv");
     
-    Col<double> initFormationParam = { -0.9 };
-    Col<double> initDissolutionParam = { -0.3 };
-    BSTERGM_MCMC_RandomLag Bstergm = BSTERGM_MCMC_RandomLag(initFormationParam, initDissolutionParam, net_3_seq);
-    //아래 GOF seq도 바꿀 것!
[…]

### 2020-12-10 doc/doc.tex
tex 1211_0230

--- 2020-12-10 doc/doc.tex
@@
 
-% \begin{figure}
-%     \begin{center}
-%     \includegraphics[scale=0.5]{LS_ggplot_2.png}
-%     \caption{Latent positions of politicians and bills.}
+\begin{figure}[h]
+    \begin{center}
+    \includegraphics[scale=0.23]{pictures/m1_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/w1_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/f1_19_nework.png}
+    \caption{Sequence 1}
     % \label{fig:ls}
-%     \end{center}
-% \end{figure}
+    \end{center}
+\end{figure}
+\begin{figure}[h]
+    \begin{center}
+    \includegraphics[scale=0.23]{pictures/m2_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/w2_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/f2_19_nework.png}
+    \caption{Sequence 2}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\begin{figure}[h]
+    \begin{center}
+    \includegraphics[scale=0.23]{pictures/m3_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/w3_19_nework.png}
+    \includegraphics[scale=0.23]{pictures/f3_19_nework.png}
+    \caption{Sequence 3}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\clearpage
 
@@
 The auxiliary chains consist of 1000 iterations and the main chain consists of 50000 iteration.
-In this setting, the BSTERGM fitting algorithm run 250 iteration of main chain per a minute on the intel i7-8700(3.20GHz) CPU.
+In this setting, the BSTERGM fitting algorithm run 2400 iterations of main chain per a minute on the intel i7-8700(3.20GHz) CPU.
 
@@
 %mean, sd
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c | c | c | c }
+            Sequence & Parameter & Estimate & Std.Error \\
+            \hline \hline
+            1 & $\theta^+$ & -2.336387 & 0.1274921 \\
+            1 & $\theta^-$ & 1.043077 & 0.1722938 \\
+            2 & $\theta^+$ & -1.435648 & 0.1009999 \\
+            2 & $\theta^-$ & 0.9332358 & 0.1358161 \\
+            3 & $\theta^+$ & -5.02916 & 0.2097569 \\
+            3 & $\theta^-$ & 2.225616 & 0.1904817 \\
+        \end{tabular}
+        \caption{Fitted parameters summary}
+        % \label{tab:data}
+    \end{table}
+\clearpage
 %hist
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 1, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 2, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 3, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\clearpage
 
 For the diagnostics, I attach traceplots and autocorrelation plots.
-%traceplot
-%acf
+The autocorrelation of each chain decrease rapidly as lag increases.
+Moreover, traceplots shows that the MCMC samples are mixed well.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 1, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 2, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 3, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\clearpage
+In addition, as a procedure of diagnostics for auxiliary chains,
+here are traceplots of the number of edges in last auxiliary chain.
+They shows that all chains are not degenerated at last parameter point during the iterations of auxiliary chain.
 
-In addition, here are traceplots of the number of edges in last auxiliary chain.
-They shows that all chains are not degenerated at last parameter point.
 %lastsampler
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.23]{pictures/net1seq_chain1_lastsampler_num_edges.png}
+        \includegraphics[scale=0.23]{pictures/net2seq_chain1_lastsampler_num_edges.png}
+        \includegraphics[scale=0.23]{pictures/net3seq_chain1_lastsampler_num_edges.png}
+    \caption{Sequence 1, left: formation, right: dissolution}
+    % \label{fig:ls}
+    \end{center}
+\end{figure}
+\clearpage
 
 Finally, I append the results of goodness of fit.
+
 %GOF table

### 2020-12-10 doc/doc.tex
tex 1211 0420

@@
 
-Moreover, we already have a generative algorithm for a network at the specific parameter point.
+The interpretation of the parameter is very easy because of the separated formation and dissolution model.
+Since we cannot know the normalizing constant in the models, but we get an odds very easily from the model definition.
+And, from parameter value, we can expect the direction or tendency of network structure with high probability.
+If $\theta^+$ is higher, a network with high $s$ have high probability at the formation model, and vice versa.
+Likewise, If $\theta^-$ is lower, a network with lower $s$ have high probability at the dissolution model, and vice versa.
+Thus, if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $\infty$, the network does not change easily over time.
+If $\theta^+$ go to $\infty$ and $\theta^-$ go to $-\infty$, the network changes drastically over time.
+If $\theta^+$ go to $\infty$ and $\theta^-$ go to $\infty$, the network becomes denser over time.
+Finally, if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $-\infty$, the network becomes sparser over time.
+
+Let us turn to the prediction method. 
+we already have a generative algorithm for a network at the specific parameter point.
 In fact, it is same to the algorithm used in the auxiliary chain.
@@
 Finally, I append the results of goodness of fit.
+Compare observed (true) value with quantiles of simulated values of parts of sufficient network statistics,
+generated by the goodness of fit algorithm.
+Since this example has simply just two edge terms, and since the BSTERGMs share same parameter
+for each network statistics but values of network statistics are different at each time point,
+the sufficient statistics do not fitted perfectly to observed networks' values.
+However, the tendency of simulated values goes with the observed values.
+
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            1 & 1-2 & $D_0$ & 4 &  0& 0& 0& 1& 1 \\
+            1 & 1-2 & $D_1$ & 1 &  0& 0& 1& 2& 3 \\
+            1 & 1-2 & $D_2$ & 0 &  1& 1& 2& 3& 4 \\
+            1 & 1-2 & $D_3$ & 3 &  1& 2& 3& 4& 5 \\
+            1 & 1-2 & $D_4$ & 1 &  1& 2& 3& 4& 6 \\
+            1 & 1-2 & $D_5$ & 3 &  2& 3& 4& 5& 7 \\
+            1 & 1-2 & $D_6$ & 3 &  2& 3& 5& 6& 8 \\
+            1 & 1-2 & $D_7$ & 5 &  3& 4& 5& 7& 9 \\
+            1 & 1-2 & $D_8$ & 9 &  3& 4& 6& 7& 9 \\
+            1 & 1-2 & $D_9$ & 5 &  3& 4& 6& 7& 9 \\
+            1 & 1-2 & $D_{10}$ & 4 &  2& 3& 5& 6& 8 \\
+            1 & 1-2 & $D_{11}$ & 4 &  1& 2& 4& 5& 6 \\
+            1 & 1-2 & $D_{12}$ & 4 &  0& 1& 2& 3& 5 \\
+            1 & 1-2 & $D_{13}$ & 2 &  0& 1& 1& 2& 3 \\
+            1 & 1-2 & $D_{14}$ & 1 &  0& 0& 1& 1& 2 \\
+            1 & 1-2 & $D_{15}$ & 0 &  0& 0& 0& 1& 2 \\
+            1 & 1-2 & $P_0$ & 8 &  25& 28& 33& 37& 42 \\
+            1 & 1-2 & $P_1$ & 10 & 29& 33& 38& 43& 48 \\
+            1 & 1-2 & $P_2$ & 29 & 30& 33& 38& 43& 47 \\
+            1 & 1-2 & $P_3$ & 27 & 28& 32& 36& 40& 44 \\
+            1 & 1-2 & $P_4$ & 13 & 19& 23& 28& 32& 37 \\
+            1 & 1-2 & $P_5$ & 37 &  9& 13& 17& 22& 26 \\
+            1 & 1-2 & $P_6$ & 40 &  3& 5& 8& 11& 15 \\
+            1 & 1-2 & $P_7$ & 20 &  3& 4& 5& 7& 9 \\
+            1 & 1-2 & $P_8$ & 20 &  0& 1& 3& 5& 7 \\
+            1 & 1-2 & $|y_2|$ & 212 &  186& 194& 204& 214& 223 \\
+        \end{tabular}
+        \caption{GOF: sequence 1 (1)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            1 & 2-3 & $D_0$ & 5 &  0& 0& 0& 0& 1 \\
+            1 & 2-3 & $D_1$ & 0 &  0& 0& 0& 1& 2 \\
+            1 & 2-3 & $D_2$ & 2 &  0& 0& 1& 2& 2 \\
+            1 & 2-3 & $D_3$ & 1 &  0& 1& 2& 2& 3 \\
+            1 & 2-3 & $D_4$ & 2 &  1& 1& 2& 3& 4 \\
+            1 & 2-3 & $D_5$ & 5 &  1& 2& 3& 4& 5 \\
+            1 & 2-3 & $D_6$ & 2 &  1& 3& 4& 5& 7 \\
+            1 & 2-3 & $D_7$ & 3 &  2& 4& 5& 7& 8 \\
+            1 & 2-3 & $D_8$ & 8 &  3& 4& 6& 7& 8 \\
+            1 & 2-3 & $D_9$ & 4 &  3& 5& 6& 8& 9 \\
+            1 & 2-3 & $D_{10}$ & 4 &  3& 4& 5& 7& 9 \\
+            1 & 2-3 & $D_{11}$ & 5 &  2& 3& 5& 6& 8 \\
+            1 & 2-3 & $D_{12}$ & 1 &  1& 2& 4& 5& 6 \\
+            1 & 2-3 & $D_{13}$ & 3 &  1& 2& 2& 4& 5 \\
+            1 & 2-3 & $D_{14}$ & 0 &  0& 1& 2& 3& 4 \\
+            1 & 2-3 & $D_{15}$ & 1 &  0& 0& 1& 2& 3 \\
+            1 & 2-3 & $P_0$ & 11 & 20& 24& 28& 32& 37 \\
+            1 & 2-3 & $P_1$ & 12 & 32& 36& 41& 47& 51 \\
+            1 & 2-3 & $P_2$ & 23 & 35& 39& 44& 49& 53 \\
+            1 & 2-3 & $P_3$ & 52 & 32& 35& 40& 45& 49 \\
+            1 & 2-3 & $P_4$ & 39 & 24& 28& 33& 37& 42 \\
+            1 & 2-3 & $P_5$ & 24 & 13& 18& 22& 27& 32 \\
+            1 & 2-3 & $P_6$ & 21 &  6& 9& 12& 17& 21 \\
+            1 & 2-3 & $P_7$ & 20 &  2& 3& 5& 8& 12 \\
+            1 & 2-3 & $P_8$ & 13 &  0& 1& 2& 4& 6 \\
+            1 & 2-3 & $|y_3|$ & 227 &  213& 222& 232& 243& 253 \\
+        \end{tabular}
+        \caption{GOF: sequence 1 (2)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
+
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            2 & 1-2 & $D_0$ & 3 &  0& 0& 0& 0& 0 \\
+            2 & 1-2 & $D_1$ & 2 &  0& 0& 0& 0& 0 \\
+            2 & 1-2 & $D_2$ & 4 &  0& 0& 0& 0& 1 \\
+            2 & 1-2 & $D_3$ & 1 &  0& 0& 0& 1& 1 \\
+            2 & 1-2 & $D_4$ & 3 &  0& 0& 1& 1& 2 \\
+            2 & 1-2 & $D_5$ & 0 &  0& 0& 1& 2& 3 \\
+            2 & 1-2 & $D_6$ & 0 &  0& 1& 2& 3& 4 \\
+            2 & 1-2 & $D_7$ & 1 &  0& 1& 2& 3& 4 \\
+            2 & 1-2 & $D_8$ & 4 &  1& 1& 2& 4& 5 \\
+            2 & 1-2 & $D_9$ & 1 &  1& 1& 2& 3& 5 \\
+            2 & 1-2 & $D_{10}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{11}$ & 1 &  1& 2& 2& 4& 5 \\
+            2 & 1-2 & $D_{12}$ & 1 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{13}$ & 2 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{14}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{15}$ & 1 &  1& 2& 3& 4& 6 \\
+            2 & 1-2 & $D_{16}$ & 1 &  1& 2& 3& 4& 6 \\
+            2 & 1-2 & $D_{17}$ & 3 &  1& 2& 2& 4& 6 \\
+            2 & 1-2 & $D_{18}$ & 2 &  1& 2& 3& 4& 6 \\
+            2 & 1-2 & $D_{19}$ & 2 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{20}$ & 2 &  1& 2& 3& 4& 5 \\
+            2 & 1-2 & $D_{21}$ & 2 &  0& 1& 2& 3& 4 \\
+            2 & 1-2 & $P_4$ & 21 & 33& 38& 43& 48& 52 \\
+            2 & 1-2 & $P_5$ & 28 & 35& 38& 43& 48& 54 \\
+            2 & 1-2 & $P_6$ & 31 & 33& 37& 42& 47& 52 \\
+            2 & 1-2 & $P_7$ & 31 & 28& 32& 37& 43& 48 \\
+            2 & 1-2 & $P_8$ & 32 & 21& 25& 31& 36& 41 \\
+            2 & 1-2 & $P_9$ & 23 & 14& 18& 23& 29& 33 \\
+            2 & 1-2 & $P_{10}$ & 27 & 8& 12& 16& 21& 26 \\
+            2 & 1-2 & $P_{11}$ & 21 & 4& 7& 10& 14& 19 \\
+            2 & 1-2 & $|y_2|$ & 342 &  344& 356& 370& 387& 400 \\
+        \end{tabular}
+        \caption{GOF: sequence 2 (1)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            2 & 2-3 & $D_0$ & 0 &  0& 0& 0& 0& 0 \\
+            2 & 2-3 & $D_1$ & 3 &  0& 0& 0& 0& 0 \\
+            2 & 2-3 & $D_2$ & 1 &  0& 0& 0& 0& 0 \\
+            2 & 2-3 & $D_3$ & 2 &  0& 0& 0& 0& 1 \\
+            2 & 2-3 & $D_4$ & 2 &  0& 0& 0& 1& 2 \\
+            2 & 2-3 & $D_5$ & 1 &  0& 0& 1& 1& 2 \\
+            2 & 2-3 & $D_6$ & 1 &  0& 0& 1& 2& 3 \\
+            2 & 2-3 & $D_7$ & 2 &  0& 1& 2& 3& 4 \\
+            2 & 2-3 & $D_8$ & 1 &  1& 1& 2& 3& 4 \\
+            2 & 2-3 & $D_9$ & 1 &  1& 1& 2& 3& 5 \\
+            2 & 2-3 & $D_{10}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{11}$ & 2 &  1& 1& 2& 4& 5 \\
+            2 & 2-3 & $D_{12}$ & 2 &  1& 2& 4& 5& 6 \\
+            2 & 2-3 & $D_{13}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{14}$ & 2 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{15}$ & 0 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{16}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{17}$ & 1 &  1& 2& 2& 4& 5 \\
+            2 & 2-3 & $D_{18}$ & 0 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{19}$ & 3 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{20}$ & 2 &  1& 2& 3& 4& 5 \\
+            2 & 2-3 & $D_{21}$ & 0 &  1& 1& 2& 3& 4 \\
+            2 & 2-3 & $P_4$ & 6 & 35& 29& 44& 48& 53 \\
+            2 & 2-3 & $P_5$ & 27 & 34& 38& 42& 48& 53 \\
+            2 & 2-3 & $P_6$ & 34 & 35& 39& 44& 49& 53 \\
+            2 & 2-3 & $P_7$ & 25 & 33& 36& 41& 46& 50 \\
+            2 & 2-3 & $P_8$ & 27 & 28& 32& 27& 42& 47 \\
+            2 & 2-3 & $P_9$ & 39 & 22& 26& 31& 36& 40 \\
+            2 & 2-3 & $P_{10}$ & 24 & 15& 19& 24& 29& 34 \\
+            2 & 2-3 & $P_{11}$ & 27 & 10& 14& 18& 23& 27 \\
+            2 & 2-3 & $|y_3|$ & 433 &  375& 387& 402& 417& 430 \\
+        \end{tabular}
+        \caption{GOF: sequence 2 (2)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
+
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            3 & 1-2 & $D_0$ & 4 &  0& 0& 0& 1& 2 \\
+            3 & 1-2 & $D_1$ & 4 &  6& 7& 8& 9& 10 \\
+            3 & 1-2 & $D_2$ & 3 &  1& 2& 3& 4& 5 \\
+            3 & 1-2 & $D_3$ & 1 &  0& 0& 1& 1& 2 \\
+            3 & 1-2 & $D_4$ & 2 &  0& 1& 1& 2& 3 \\
+            3 & 1-2 & $D_5$ & 7 &  1& 2& 3& 4& 5 \\
+            3 & 1-2 & $D_6$ & 4 &  3& 4& 5& 7& 8 \\
+            3 & 1-2 & $D_7$ & 3 &  3& 4& 5& 6& 7 \\
+            3 & 1-2 & $D_8$ & 4 &  2& 3& 4& 5& 6 \\
+            3 & 1-2 & $D_9$ & 6 &  2& 3& 4& 5& 6 \\
+            3 & 1-2 & $D_{10}$ & 0 &  2& 3& 4& 5& 6 \\
+            3 & 1-2 & $D_{11}$ & 6 &  2& 3& 4& 5& 6 \\
+            3 & 1-2 & $D_{12}$ & 2 &  1& 1& 2& 3& 4 \\
+            3 & 1-2 & $D_{13}$ & 1 &  0& 1& 1& 2& 3 \\
+            3 & 1-2 & $D_{14}$ & 0 &  0& 1& 1& 2& 3 \\
+            3 & 1-2 & $D_{15}$ & 0 &  0& 0& 1& 1& 2 \\
+            3 & 1-2 & $P_0$ & 7 & 9& 10& 12& 13& 14 \\
+            3 & 1-2 & $P_1$ & 22 & 12& 14& 17& 20& 23 \\
+            3 & 1-2 & $P_2$ & 21 & 15& 18& 21& 24& 27 \\
+            3 & 1-2 & $P_3$ & 42 & 23& 27& 30& 34& 37 \\
+            3 & 1-2 & $P_4$ & 24 & 25& 29& 32& 36& 40 \\
+            3 & 1-2 & $P_5$ & 32 & 22& 25& 29& 33& 37 \\
+            3 & 1-2 & $P_6$ & 14 & 15& 18& 21& 23& 26 \\
+            3 & 1-2 & $P_7$ & 9 & 12& 14& 17& 19& 22 \\
+            3 & 1-2 & $P_8$ & 14 & 7& 9& 12& 14& 16 \\
+            3 & 1-2 & $P_9$ & 6 &  3& 5& 7& 9& 11 \\
+            3 & 1-2 & $P_10$ & 1 &  1& 2& 3& 4& 5 \\
+            3 & 1-2 & $|y_2|$ & 196 &  198& 201& 206& 209& 213 \\
+        \end{tabular}
+        \caption{GOF: sequence 3 (1)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
+\begin{table}[h!]
+    \centering
+        \begin{tabular}{c| c | c | c | c |c |c |c |c }
+            Sequence& timelag & netstat & obs & q0.1 & q0.25 & q0.5 & q0.75 & q0.9 \\
+            \hline \hline
+            3 & 2-3 & $D_0$ & 11 &  2& 3& 3& 4& 4 \\
+            3 & 2-3 & $D_1$ & 11 &  2& 3& 4& 5& 6 \\
+            3 & 2-3 & $D_2$ & 12 &  2& 2& 3& 4& 5 \\
+            3 & 2-3 & $D_3$ & 9 &  1& 1& 2& 3& 3 \\
+            3 & 2-3 & $D_4$ & 2 &  2& 2& 3& 4& 5 \\
+            3 & 2-3 & $D_5$ & 1 &  4& 5& 6& 7& 8 \\
+            3 & 2-3 & $D_6$ & 0 &  2& 3& 4& 6& 7 \\
+            3 & 2-3 & $D_7$ & 2 &  2& 3& 4& 5& 6 \\
+            3 & 2-3 & $D_8$ & 1 &  2& 3& 4& 5& 6 \\
+            3 & 2-3 & $D_9$ & 0 &  2& 3& 4& 5& 6 \\
+            3 & 2-3 & $D_{10}$ & 1 &  1& 2& 3& 4& 5 \\
+            3 & 2-3 & $D_{11}$ & 0 &  1& 2& 3& 4& 5 \\
+            3 & 2-3 & $D_{12}$ & 0 &  0& 1& 1& 2& 3 \\
+            3 & 2-3 & $D_{13}$ & 0 &  0& 0& 0& 1& 2 \\
+            3 & 2-3 & $D_{14}$ & 1 &  0& 0& 0& 0& 1 \\
+            3 & 2-3 & $D_{15}$ & 1 &  0& 0& 0& 1& 1 \\
+            3 & 2-3 & $P_0$ & 36 & 8& 10& 12& 14& 16 \\
+            3 & 2-3 & $P_1$ & 26 & 20& 22& 25& 28& 31 \\
+            3 & 2-3 & $P_2$ & 4 & 23& 26& 30& 34& 37 \\
+            3 & 2-3 & $P_3$ & 0 & 29& 33& 37& 40& 44 \\
+            3 & 2-3 & $P_4$ & 1 & 21& 24& 27& 31& 34 \\
+            3 & 2-3 & $P_5$ & 0 & 16& 18& 22& 25& 28 \\
+            3 & 2-3 & $P_6$ & 0 &  8& 10& 12& 14& 16 \\
+            3 & 2-3 & $P_7$ & 1 &  6& 7& 9& 11& 13 \\
+            3 & 2-3 & $P_8$ & 0 &  3& 4& 6& 8& 10 \\
+            3 & 2-3 & $P_9$ & 0 &  1& 1& 2& 4& 5 \\
+            3 & 2-3 & $P_10$ & 0 &  0& 1& 1& 2& 3 \\
+            3 & 2-3 & $|y_3|$ & 227 &  180& 183& 187& 190& 193 \\
+        \end{tabular}
+        \caption{GOF: sequence 3 (2)}
+        % \label{tab:data}
+    \end{table}
+\clearpage
 

### 2020-12-10 BayesianSTERGM/main.cpp
main GOF from sample

@@ -617,9 +617,9 @@ int main()
     Network net_w2 = Network(w2_19, false);
     Network net_w3 = Network(w3_19, false);
     
-    //net_m1.undirected_printSummary();
-    //net_w1.undirected_printSummary();
-    //net_f1.undirected_printSummary();
+    // net_m1.undirected_printSummary();
+    net_w3.undirected_printSummary();
+    net_f3.undirected_printSummary();
 
     vector<Network> net_1_seq = { net_m1, net_w1, net_f1 };
     vector<Network> net_2_seq = { net_m2, net_w2, net_f2 };
@@ -802,7 +802,7 @@ int main()
 
     
 
-
+    // =============================================================
     ////BSTERGM GOF from sample
     //vector<double> disSmpl = { 1 };
     //vector<double> forSmpl = {1 };

### 2020-12-11 doc/paper_format.tex
directory adj

--- 2020-12-10 doc/doc.tex
@@
-\ifcase 1  % choose 0=slides, 1=article, 2=refart
-   \documentclass[aspectratio=169,ignorenonframetext,9pt]{beamer}
-\or\documentclass[a4paper,11pt]{article}
-   \usepackage{url,beamerarticle}
-\or\documentclass[a4paper,11pt]{refart}
-   \let\example\relax
-   \usepackage{url,beamerarticle}
-\fi
-
-\ifcase 0  % choose a theme like these
-    % \usetheme{boxes}
-    \usetheme{Boadilla}
-    % \usetheme{Goettingen}% I recommend
-    % \usetheme{Singapore}
-    % \usetheme{Pittsburgh}
-    % \usetheme{Madrid}
-    % \usetheme{Warsaw} % common choice, but often poor
-\fi
-
-\usepackage{algorithm,algorithmicx}
-\usepackage{graphicx,pgfplots,parskip}
-\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,url,array}
-\usepackage{cite}
-
-\theoremstyle{plain}
+\documentclass[a4paper, 11pt]{report}
+\usepackage{times,epsfig,calc,subfigure}
+\usepackage[authoryear,round]{natbib}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{multicol}
+%\usepackage[tablesfirst, nolists]{endfloat}
+\usepackage{amsbsy, epsfig, epsf, psfrag, booktabs, graphicx, amssymb, enumerate}
+\usepackage[hidelinks,bookmarks=false]{hyperref}
+\usepackage{lscape}
+\usepackage{bm}
+\usepackage{float}
+\floatplacement{table}{!htbp}
+\usepackage{adjustbox}
+\usepackage{color, colortbl}
+
+\definecolor{Gray}{gray}{0.9}
+
+\usepackage{makecell}
+\usepackage{changepage}%adjust
+\usepackage{setspace}%stretch
+\usepackage{booktabs} 
+\usepackage{multirow}
+%\usepackage{hypernat}
+\usepackage{authblk}
+\usepackage{rotating}
+\usepackage{multicol}
+\usepackage{epstopdf}
+\usepackage{tabularx, caption}
+\usepackage[flushleft]{threeparttable}
+\usepackage{grad}
+
+\usepackage{kotex}
+\usepackage{placeins}
+
+\theoremstyle{definition}
+\newtheorem{algo}{Algorithm}[section]
+
+\makeatletter
+\newcommand{\distas}[1]{\mathbin{\overset{#1}{\kern\z@\sim}}}%
+\newsavebox{\mybox}\newsavebox{\mysim}
+\newcommand{\distras}[1]{%
+	\savebox{\mybox}{\hbox{\kern3pt$\scriptstyle#1$\kern3pt}}%
+	\savebox{\mysim}{\hbox{$\sim$}}%
+	\mathbin{\overset{#1}{\kern\z@\resizebox{\wd\mybox}{\ht\mysim}{$\sim$}}}%
+}
+\makeatother
+
+\renewcommand{\baselinestretch}{1}
+\linespread{1.4}
+\renewcommand{\arraystretch}{1.3}
+\newcommand{\cov}{\mathop{\rm cov}\nolimits} %covariance
 \newtheorem{thm}{Theorem}[section]
-\newtheorem{lem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
-
-\theoremstyle{definition}
-\newtheorem{defn}{Definition}[section]
-\newtheorem{conj}{Conjecture}[section]
-\newtheorem{exmp}{Example}[section]
-\newtheorem{algo}{Algorithm}[section]
-
-\title{BSTERGM: Bayesian Separable-Temporal Exponential family Graph Models}
-\author{Choi, Seokjun}
-\date{14 Dec. 2020}
+\newtheorem{defn}[thm]{Definition}
+\newcommand\defeq{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
+\DeclareMathOperator*{\argmin}{argmin}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%   TITLE
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\title{BSTERGMs: Bayesian Separable-Temporal Exponential family Graph Models}
+\author{Seokjun Choi}
+\date{December 2020}
 
 \begin{document}	
-
-\begin{frame}
-\maketitle
-\end{frame}
+\maketitle \pagestyle{plain}
+
+\newpage
+\thispagestyle{empty}
+
+  \null
+  \begin{center}
+    \vskip -1.4cm
+    {\fontsize{14pt}{14pt}\selectfont This certifies that Master's thesis of Seokjun Choi is approved.}
+    \vskip 4cm%
+    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
+    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Thesis Supervisor: Prof. Ick Hoon Jin\\
+    \vskip 2cm
+    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
+    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Committee Member: Prof. Chul Eung Kim\\
+    \vskip 2cm
+    {\hskip 5cm\fontsize{12pt}{12pt}\selectfont \underline{\hskip 8cm}}\vspace*{0.2cm}\\
+    \hskip 5cm\fontsize{12pt}{12pt}\selectfont Committee Member: Prof. Jaewoo Park\\
+    \vskip 3cm%
+    {\fontsize{14pt}{14pt}\selectfont {The Graduate School \vskip 1.0em Yonsei University \vskip 1.0em Dec 2020}}%
+  \end{center}
+  \par
+\clearpage
+
+\pagestyle{plain}
+
+\clearpage \pagenumbering{roman} \tableofcontents \clearpage
+
+\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
+	\hspace*{-0.29in} List of Figures}
+\listoffigures
+
+\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
+    \hspace*{-0.29in} List of Tables}
+\listoftables
+
+\clearpage \addcontentsline{toc}{chapter}{\protect\numberline{}
+    \hspace*{-0.29in} Abstracts}
+
+%%%%%%%%%%%%
+% ABSTRACT %
+%%%%%%%%%%%%
 
 \begin{abstract}
-Separable Temporal Exponential Random Graph Models (STERGMs) are useful models to describe changes in dynamic networks' structures evolving over time. 
-The Bayesian-inference based approach for the model, however, had not been developed yet. 
-In this paper, I propose the Bayesian STERGMs (BSTERGMs): STERGMs with the Bayesian setting, and I illustrate the computational algorithm that generates posterior samples from the model dealing with the doubly-intractable constant problem.
-Furthermore, I show the interpretability of the model in the Bayesian context.
-
+\begin{center}
+{\LARGE 
+BSTERGMs: The Bayesian Separable-Temporal Exponential family Graph Models
+}
+\end{center}
+\indent
+\vskip 0.8cm
+\noindent
+\vskip 0.2cm
+
+The separable Temporal Exponential Random Graph Models (STERGMs)(\cite{RN93}) are useful models to describe changes in dynamic networks' structures evolving over time. 
+The Bayesian-inference based approach for the models, however, had not been developed yet. 
+In this paper, I propose the Bayesian Separable Temporal Exponential Random Graph Models (BSTERGMs): STERGMs with the Bayesian setting, 
+and illustrate a computational algorithm that generates posterior samples from the models dealing with the doubly-intractable constant problem. 
+Furthermore, I show an inference method of the models in the Bayesian context.
+
+\vspace{\stretch{1}} \noindent
+\hrulefill\\
+{\bf Key words}
+\parbox[t]{0.8\textwidth}
+{ : Networks, Exponential Random Graph Models, Bayesian inference, Marcov chain Monte Carlo, Longitudinal}
 \end{abstract}
 
-% \begin{frame}{Outline}
-% \tableofcontents
-% \end{frame}
-
-\section{Introduction}
-Nowadays, new kinds of data having special forms emerge in many field.
-These new data are not confined by a series of values of real numbers as traditional data are in the past. 
-The data may have non-number type of value, special structure, or even integrated different type of values with model-specific relation.
-These data are called complex data, and an attention of statistical analysis on the data increases over time.
-
-As one form of the complex data, networks get highlighted attention among complex data for statistical analysis.
-Many phenomena can be represented by some relational structure, and the structure can be expressed effectively
-by a form of a network. For example, social networks including both offline and online can be thought as a sort of networks.
-In this view, some social aspects of relation between actors can be modeled with an object of social science
-including an analysis of power structure and a summary of intimacy structure on given group.
-Not only that, application to a transmitted disease leads a modeling using real data in epidemiology.
-
-Since statistical models on network provide not only lots of possibility of application but also
-interesting and useful features on observed network, statistical models have important role.
-The statistical model gives information about local relations like formation or dissolution between
-two actors as well as global features of the whole network such as probability that a specific structure
-realizes. Not only that, the statistical models investigate a complex behavior of network including
-how the global features of an observed network may be related to local network structure.
-During the investigation, statistical model can capture existing dependency on a given network
-and provide a tool to understand what certain structural features are more common on rare.
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\clearpage
+\chapter{Introduction} \label{Intro}
+\pagenumbering{arabic}
+
+Nowadays, new kinds of data having special forms emerge in many fields.
+A series of real number values do not confine these new data as traditional data are in the past.
+The data may have a non-number type of value, unique structure, or even integrated different types of values with model-specific relation.
+These data are called complex data, and the attention of statistical analysis on the data increases over time.
+
+As one form of complex data, networks, or random graphs get highlighted attention among complex data for statistical analysis.
+A relational structure can represent many phenomena in the real world, and the structure can be expressed effectively by a form of a network. 
+For example, social networks, including offline and online, can be thought of as networks.
+In this view, some social aspects of the relationship between actors can be modeled under an object of social science, 
+including an analysis of power structure and a summary of intimacy structure on a given group.
+Not only that, application to a transmitted disease in epidemiology enables statistical modeling based on real data.
+
+The ability to capture distinct features on the observed network enlarges statistical models' usefulness on networks. 
+The statistical models give information about local relations like formation or dissolution between two actors 
+as well as global features of the whole network, such as the probability that a specific structure realizes. 
+The statistical models investigate the complex behavior of networks, 
+including how an observed network's global features may be related to local network structure. 
+During the investigation, the statistical models can capture existing dependency on a given network and provide a tool 
+to understand what certain structural features are more common on rare.
 
@@
 \cite{RN117}, \cite{RN123}, \cite{RN124} and \cite{RN122} proposed the models and apply them to analyze dependent structures, 
-and the Bayesian approaches of the models were started by \cite{RN115} and \cite{RN116}.
+and \cite{RN115} and \cite{RN116} started the Bayesian approaches of the models.
 In the ERGMs, The dependencies are captured using a series of local features called network statistics from classical graph-theory. 
-Since these models play important role in this paper, I will introduce more precisely.
-
-Next natural question is a way to model a network sequence, or the evolution of network over time.
-For instance, one may want to model social network structures for 1 year, or
-simulation of a disease infection based on the sequential observations in the past.
-The first exploration applying ERGMs was \cite{RN109}. 
+Since these models play an important role throughout this paper, I will introduce them more precisely in the following chapter.
+
+The next natural question is a way to model a network sequence or network evolution over time. 
+For instance, one may want to model social network structures for one year or 
+simulate a disease infection based on the sequential observations in the past. 
+The first exploration applying ERGMs to a network sequence was \cite{RN109}. 
 After them, a notable contributions were of \cite{RN125} and \cite{RN126}, the Temporal ERGMs(TERGMs). 
-In the TERGM, ERGMs are used by modelling the transition probability between times.
-Moreover, \cite{RN93} extend the TERGMs to the Separable Temporal ERGMs(STERGMs), considering an
-separation between duration - how long ties tend to last - and incidence - the rate at which new ties are formed -
-from prevalence - for example, the total number of ties - of the network properties.
-Imposing a conditional independence condition between the duration and incidence,
-the models can be extended effectively using similar fitting methods with TERGMs
+In the TERGMs, ERGMs are used by modeling the transition probability between times. 
+Moreover, \cite{RN93} extend the TERGMs to the Separable Temporal ERGMs(STERGMs), 
+considering the separation between duration - how long ties tend to last - and 
+incidence - the rate at which new ties are formed -from 
+prevalence - for example, the total number of ties - of the network properties. 
+By imposing a conditional independence condition between the duration and incidence on the model, 
+the models can be extended to more effectively interpretable ones using similar fitting methods with TERGMs 
 without adding much complexity.
 
-Although the STERGMs give us strength when it comes to interpret the fitted parameter,
-a Bayesian approach for the STERGMs has not been conducted. It is partly because of a hardship
-coming from the huge computation in a parameter estimation method for the model.
-Nonetheless, the Bayesian approach of STERGMs provides the statistical inference method
-following the standard Bayesian way, so the approach has huge advantage of convenience
-for statistical works for fitted parameter and predictions to next time network structure.
-Thus, in this paper, I specify the bayesian separable temporal exponential random graph models(BSTERGMs)
-and propose a fitting algorithm for the model parameter. 
+Although the STERGMs give us strength when interpreting the fitted parameter, 
+a Bayesian approach for the STERGMs has not been conducted. 
+It is partly because of the hardship from the massive computation in a parameter estimation method for the models. 
+Nonetheless, the Bayesian approach of STERGMs provides the statistical inference method following the standard Bayesian way, 
+so the approach has a huge advantage of convenience for statistical works for fitted parameters and predictions to next time network structure.
+Thus, in this paper, I specify the bayesian separable temporal exponential random graph models(BSTERGMs) and propose a fitting algorithm for the model parameter.
 I also discuss ways of inference, the diagnostics procedure, and checking the goodness of fit.
 
-In chapter 2, I review the background meterials including ERGMs, BERGMs, TERGMs, and STERGMs,
-and in chapter 3, I show the detail setting of BSTERGMs. In chapter 4, I propose the fitting method for BSTERGM.
-In chapter 5, I explain ways to conduct diagnostics, statistical inference, and goodness of fit check.
-In chapter 6, I illustrate an application of BSTERGMs to network data of scholars and faculty
-from the UCLA medical working project team. Finally, I discuss some extension and 
-ways to supplement the BSTERGM further for future research.
-
-\section{Background}
-\subsection{Random graphs}
-For statistical analysis on relations between subjects of data,
-I will introduce terminologies and notations for denoting components related to networks, 
-which are a representation of these type of data.
-
-We say each subject on the data to a node. They are same as a vertex in graph theory.
+Chapter 2 reviews the background materials, including the ERGMs, the BERGMs, the TERGMs, and the STERGMs. 
+In chapter 3, I propose the BSTERGMs and show the detailed setting of BSTERGMs. 
+In the following chapter, chapter 4, I offer the estimation method for the BSTERGMs. 
+Chapter 5 explains ways to conduct diagnostics, statistical inference, and goodness of fit checks. 
+Besides, I illustrate an application of BSTERGMs to network data of scholars and faculty
+from the UCLA medical working project team in chapter 6, 
+Finally, I discuss some possible extensions and ways to supplement the BSTERGMs further for future research.
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+
+\chapter{Background} \label{Chapter2}
+\section{Random graphs}
+Networks or random graphs play a central role as a general model 
+describing relations among subjects in the statistical context. 
+To begin with, I will briefly summarize some terminologies and notations for concepts related to the networks.
+
+I will use two words, 'graphs' and 'network,' as the same meaning without distinction. 
+We say each subject on the data to a node. 
+They sometimes are called vertex in graph theory. 
 In addition, we say a tie joining two nodes to an edge. A tie can be directed or not. 
-These two components have fundamental role describing the data.
-Define a graph or a network to $y=(V,E)$, where $V$ is a set of node and $E$ is a set of edges of a network.
+These two components have a fundamental role in describing the relational structure. 
+Define a graph or a network to $y=(V,E)$, where $V$ is a set of nodes, and $E$ is a set of edges of a network.
 
 For simplicity, I will deal with a situation without weight on each node or edge.
-Then, a graph $y$ can be represented by simpler form.
+Then, a graph $y$ can be represented by a more straightforward form.
 Assume that the number of node $n\in \mathbb{N}$ on data is given.
@@
 
-If edges of $y$ have directions, then $y$ is called a directed graph. Otherwise, $y$ is called a undirected graph.
+If edges of $y$ have directions, then $y$ is called a directed graph. Otherwise, $y$ is called an undirected graph.
 These are obvious that, for given $n\in \mathbb{N}$,
@@
 and $|\mathcal{Y}|=2^{n(n-1)/2}$ if $\mathcal{Y}$ is the set of all graphs of undirected one.
-Thus, the size of $\mathcal{Y}$ grows exponentially when n increases.
-
-I extend the concept of graphs to be able to describe a situation that connections among nodes are random.
+Thus, the size of $\mathcal{Y}$ exponentially grows when n increases.
+
+I extend the concept of graphs to describe a situation that connections among nodes are random.
 Let $\Omega$ be an event set of all possible connected situations among given nodes.
@@
 
-\subsection{ERGMs: Exponential random graph models}
-From the definition of random graph, one of the natural questions is 
-how we can model a distribution on $\mathcal{Y}$.
+\section{ERGMs: Exponential random graph models}
+One of the natural questions from the definition of the random graph is 
+how we can model a probability distribution on $\mathcal{Y}$.
 As a answer, the Exponential random graph models (ERGMs) (\cite{RN123} and \cite{RN122})
-gives a classical but 
-flexible scheme to model for time-invariant networks represented as a single directed or undirected graph.
+gives a classical but flexible scheme to model for time-invariant networks represented as a single directed or undirected graph.
 Moreover, ERGMs provide a statistical approach to the systematic exploration of several local features simultaneously and 
@@
 
-Specifically, following the notation of previous section,
+Specifically, following the notation of the previous section,
 let $\mathcal{Y}$ be a set of random graphs with $n$ nodes and $Y \in \mathcal{Y}$ be a random graph.
@@
 and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
-Models which have such a form called the ERGMs.
-
-Note that the form of the right-hand side is the canonical form of exponential family distributions.
-That is why the model is called exponential. Moreover, observe that the normalizing constant cannot be directly calculated
+Models that have such a form are called the ERGMs.
+
+Note that the form of the right-hand side is a canonical form of exponential family distributions.
+That is why the model is called exponential. 
+Although the form is familiar, directly calculating the normalizing constant is not possible in practical data analysis 
 because the size of $\mathcal{Y}$ grows exponentially as $n$ grows. 
-Thus in real data analysis, the calculation of $c(\theta)$ should be avoided.
-For more detail of fitting algorithm, see \cite{RN104} the detail on fitting algorithm using MCMCMLE.
-You can also see \cite{RN100} addressing the ergm package in R with the MCMCMLE or the classic method using psudolikelihood.
-
-\subsection{BERGMs: Bayesian ERGMs}
-\cite{RN115} and \cite{RN116} extend ERGMs to bayesian setting. Putting prior $p(\theta)$ to the parameter $\theta$, the model becomes
+Thus for the application to the real data, the calculation of $c(\theta)$ should be avoided.
+Luckily, some fitting algorithms have been developed for this issue. 
+If you need, see \cite{RN100} addressing the ergm package in R with the MCMCMLE or the classic method using pseudolikelihood. 
+For more detail of the MCMCMLE algorithm, see \cite{RN104}
+
+\section{BERGMs: Bayesian ERGMs}
+\cite{RN115} and \cite{RN116} extend ERGMs to bayesian setting. 
+Putting prior $p(\theta)$ to the parameter $\theta$, the model becomes
 \[p(\theta|y)=\frac{p(y|\theta)p(\theta)}{p(y)}=\frac{exp(\theta^T s(y))}{c(\theta)}\frac{p(\theta)}{p(y)}\]
@@
 because of the difficulty of summation when $n$ grows and of integration of $p(y)$.
-Moreover, the constant depends on the parameter $\theta$, so ordinary MCMC technique cannot be used.
-This situation is said to doubly-intractable constant problem(\cite{RN119}), and causes serious challenge to estimate the parameter vector $\theta$.
-Thus, The estimation of $\theta$ needs more advanced algorithm than the standard MCMC procedure.
-The solution of Caimo and Friel is using the exchange algorithm. See \cite{RN115} for details of fitting methods of BERGMs
+Moreover, the constant depends on the parameter $\theta$, so the ordinary MCMC technique cannot be used.
+This situation is said to the doubly intractable constant problem(\cite{RN119}), 
+and causes a serious challenge to estimate the parameter vector $\theta$.
+Thus, The estimation of $\theta$ needs an advanced algorithm than the standard MCMC procedure.
+The solution of Caimo and Friel is using the exchange algorithm. 
+You can see \cite{RN115} for details of fitting methods of BERGMs
 and \cite{RN127} for the exchange algorithm.
 
-\subsection{TERGMs: Temporal Exponential family Random Graphs Models}
+\section{TERGMs: Temporal Exponential family Random Graphs Models}
 Let's consider a sequence of random graph $\{Y_1, Y_2, ..., Y_T\} (T\in\mathbb{N})$ and assume that
@@
 and focus on the transition probability along index of the sequence.
-For convenience, I refer the index as a time.
+For convenience, I refer to the index as a time.
 
@@
 and a normalizing constant $c(\theta, y_{t-1})=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y, y_{t-1}))$.
-Models which have a such form called the TERGMs.
-
-The estimation of $\theta$ can be performed by MCMCMLE,
-but the estimation suffers severe degeneracy problem.
-Furthermore, When we interpret the estimated TERGMs parameters,
+Models that have such a form are called the TERGMs.
+
+The estimation of $\theta$ can be performed by MCMCMLE.
+Nevertheless, the estimation suffers severe degeneracy problem.
+Furthermore, When interpreting the estimated TERGMs parameters,
 we cannot separate the generation and duration because
-the higher value of $\theta$ increases both.
+the parameter value of $\theta$ affects both.
 These raise demand for modified models to remedy the degeneracy and enable separable interpretation.
 
-\subsection{STERGMs: Separable-Temporal Exponential family Random Graphs Models}
+\section{STERGMs: Separable-Temporal Exponential family Random Graphs Models}
 One solution to remedy the problems of TERGMs is 
@@
 The key idea is to separate the formation models and dissolution model
-for decoupling an effect on edge's duration of model parameter from edge's incidence
+for decoupling an effect on edge's duration of the model parameter from edge's incidence
 when setting a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds, $T\in\mathbb{N}$),
@@
 Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (on $\mathcal{Y}$).
-Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$.
+Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting of all graphs which have equal or additional edges comparing to $y_{t-1}$.
 Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
@@
 
-Additionally, assume the first-order Markov assumption
+Additionally, assume the first-order Markov assumption,
 \[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
@@
 \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
-Models which have a such form called the STERGMs.
-
-\subsection{Network Statistics for ERGMs and derived models}
+Models that have such a form are called the STERGMs.
+
+\section{Network Statistics for ERGMs and derived models}
 Now, I briefly discuss the terms of network statistics.
@@
 reflect the structure of given network without identifying each node index.
-Common candidates of the statistics are here.
+Typical candidates of the statistics are here.
 
 Let $y\in\mathcal{Y}$ and $n\in\mathbb{N}$ be the number of nodes of $y$.
-The simplest candidate is the number of edges, denoted by $|y|$.
-Following \cite{RN104}, common candidates are distributions defined by some sense of structure of given graph.
-For a given $i, 1\leq i \leq n-1$, $D_i(y)$ is defined to be a number of nodes in y whose the number of edges incident to the node equals $i$,
+The most straightforward candidate is the number of edges, denoted by $|y|$.
+
+Following \cite{RN104}, usual candidates are distributions defined by some sense of the structure of a given graph. 
+For a given $i, 1\leq i \leq n-1$, $D_i(y)$ is defined as the number of nodes in y whose the number of edges incident to the node equals $i$, 
 and called $i$-th order node degree distribution.
-For a given $i, 1\leq i \leq n-2$, $P_i(y)$ is defined to be a number of dyads $(j,k), 1\leq j,k \leq n$
+
+For a given $i, 1\leq i \leq n-2$, $P_i(y)$ is defined to be the number of dyads $(j,k), 1\leq j,k \leq n$
 in y whose $j$ and $k$ have nodes connected by an edge each other and they share exactly $i$ connected nodes in common,
 and called $i$-th order edgewise shared partner distribution.
+
 Note that, the bundle of all order of node degree distribution and all order of edgewise shared partner distribution,
@@
 \cite{RN128} proposed graph statistics including $k$-star $S_k(y)$ for $1\leq k \leq n-1$ and $k$-triangles $T_k$ for $1\leq k \leq n-2$.
-The $k$-star consists of a node together with a set of $k$ of its connected nodes,
-and the $k$-triangle consists of $k$ triangles that share one common edge.
+The $k$-star consists of a node together with a set of $k$ of its connected nodes.
+The $k$-triangle consists of $k$ triangles that share one common edge.
 They can be expressed by the definition of node degree distributions and edgewise shared partner distributions,
@@
 Another interesting statistics are geometrically weighted value of distributions defined by
-\cite{RN118},\cite{RN104},\cite{RN128} and \cite{RN103}.
+\cite{RN118}, \cite{RN104}, \cite{RN128} and \cite{RN103}.
 For example, geometrically weighted (node) degree with parameter $\tau$ is defined by
@@
 In the context of temporal models, parameters in the model take two network as a arguments like $s(y_t,y_{t-1})$.
-One of common choices for these statistics is the bundle of sufficient statistics of first argument $y_t$.
+One of the common choices for these statistics is the bundle of sufficient statistics of first argument $y_t$.
 Another common choice is using a difference, $s(y_t,y_{t-1})=s'(y_t)-s'(y_{t-1})$ where $s'$ is a network statistics of given network.
-Although there are lots of other statistics related to ERGMs, I omit them and end this subsection here.
-If you need, see \cite{RN107} for the ERGMs, and \cite{RN125}'s chapter 2.1 for temporal the ERGMs.
-
-\section{Bayesian STERGMs}
-A goal is to convert the STERGMs to the Bayesian setting.
+Although there are many other statistics related to ERGMs, I omit them and end this section here.
+For additional example, see \cite{RN107} for the ERGMs, and \cite{RN125}'s chapter 2.1 for temporal the ERGMs.
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\chapter{Bayesian STERGMs} \label{Cahpter3}
+
+The goal of this chapter is to convert the STERGMs to the Bayesian setting.
 With priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$,
 the Bayesian model for the STERGMs can be made. I call this model Bayesian Separable Temporal Exponential family Random Graph Models,
-or the BSTERGMs. More explicitly, for given the number of nodes $n$, let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (of $\mathcal{Y}$).
+or the BSTERGMs.
+
+More explicitly, for given the number of nodes $n$, let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (of $\mathcal{Y}$).
 As the STERGMs, define $\mathcal{Y}^+|_t$, $\mathcal{Y}^-|_t$ be subsets of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$,
@@
 where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$ as set above.
-
-Note that, the BSTERGMs also have similar kinds of problems of the BERGMs.
+I define models that have such forms as the STERGMs.
+
+Note that the BSTERGMs also have similar kinds of problems to the BERGMs.
 The normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically cannot be computed because we need to sum up too many terms.
 The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
-Thus, we cannot use an ordinary MCMC algorithm to get the posterior sample in this model,
+Thus, we cannot use a standard MCMC algorithm to get the posterior sample in this model,
 because the constant part remains when calculating the MCMC ratio.
-
-\section{Estimation}
+I will address this issue in the next chapter.
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\chapter{Estimation} \label{Chapter4}
 
@@
 (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined.) 
-Then how can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using the BSTERGMs?
-To solve the doubly-intractable constant problem, I propose an novel algorithm using the exchange algorithm,
+How can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using the BSTERGMs?
+To solve the doubly intractable constant problem, I propose a novel algorithm using the exchange algorithm,
 one of the special MCMC techniques by \cite{RN127}, extending the algorithm of the BERGMs'(\cite{RN115}).
-Although MCMC types of approach cannot give an analytic expression of posterior distribution,
-it gives posterior samples as much as we want, and using the samples we can conduct statistical inference on given data.
-
-The key idea is running two MCMC chain to make independent sample to cancel the constant parts
+Although an approach using MCMC cannot give an analytic expression of the posterior distribution, 
+it gives posterior samples as much as we want, and we can conduct statistical inference on given data using the samples.
+
+The key idea is running two MCMC chains to draw independent samples to cancel the constant parts
 when calculating the MCMC accept-reject ratio.
-As a motivation, consider the ordinary MCMC ratio for $m$-th MCMC iteration in the BSTERGMs context 
-to show doubly intractable constant problem explicitly.
+As a motivation before introducing the new algorithm, consider the ordinary MCMC ratio for $m$-th MCMC iteration in the BSTERGMs context 
+to show the doubly intractable constant problem explicitly.
 
@@
 
-The last expression shows that all constants are canceled out.
+The last expression has no constant term depending on the parameters.
 This finding leads us to the fitting algorithm for the BSTERGMs using the exchange algorithm.
@@
 
-If $s(.,.)$ and $\theta$ can be multi-dimension (say, of $\mathbb{R}^p$,) replace the multiplication of the exponent parts to dot product between two vectors.
+If $s(.,.)$ and $\theta$ can be multi-dimension, (say, of $\mathbb{R}^p$,)
+replace the multiplication of the exponent parts to dot product between two vectors.
 In addition, when implementing the algorithm, I strongly recommend to use $log\pi$ instead of $\pi$ to avoid an underflow problem.
@@
 
-To generate the exchange graph $y_{ex}$ for time $t$ at the second part of main chain,
-a generative algorithm at a given parameter points is needed.
-However, since the normalizing constant is still unknown, we cannot generate the exchange algorithm 
-directly using the model's expression of the distribution.
-Thus, I use one more MCMC chain to get the exchange samples without calculating the constants.
+To get the exchange network $y_{ex}$ for time $t$ in the second part of the main chain, 
+we need a generative algorithm at a given parameter points. 
+However, since the normalizing constant is still unknown, 
+I cannot get the generative algorithm directly using the model's expression. 
+Thus, I use one more MCMC chain.
 
@@
 \end{algo}
+
 After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
 
-\section{Diagnostics, inference, and Bayesian GOF}
-
-\subsection{Model diagnostics}
-The ERGMs and all derived models of the ERGMs are suffered degeneracy problem:
-all nodes of exchange sample become fully connected or totally isolated as the MCMC chain runs. 
-If the situation is worse, the problem cause an explosion to $\infty$ or $-\infty$ of the fitted parameters.
-The BSTERGMs are not exception. In other words, the diagnostic process is very important part for the BSTERGMs.
-
-Since two MCMC chains are run in the fitting algorithm of the BSTERGMs, two diagnostic tasks should be proceeded.
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\chapter{Diagnostics, inference, and the Bayesian GOF} \label{Chapter5}
+
+\section{Model diagnostics}
+The ERGMs and all derived models of the ERGMs have suffered the degeneracy problem: 
+all exchange samples' nodes become fully connected or totally isolated as the MCMC chain runs. 
+The problem may cause an explosion to $\infty$ or $-\infty$ of the fitted parameters, 
+failing to converge of MCMC chains. The BSTERGMs are no exception. 
+In other words, the diagnostic process is a very important part of using the BSTERGMs.
+
+Since two MCMC chains run in the fitting algorithm of the BSTERGMs, two diagnostic tasks should proceed.
 
@@
 which are not different with running ordinary MCMC algorithm.
-Depict traceplots of each parameter chain to check the convergence and the mixing.
-If needed, cut samples of burn-in period and do thinning.
-Plotting sample autocorrelation for each lag is also important. Calculate ESS if it is needed.
+Depict trace plots of each parameter chain to check the convergence and the mixing.
+If needed, cut samples of the burn-in period and thin the samples.
+Plotting sample autocorrelation for each lag is also essential. Calculate ESS if it is needed.
 
 However, diagnostics of auxiliary chains is challenging. 
-One reason of the difficulty is that there are $M$ auxiliary chains in our whole procedure.
-Practically, it is not possible to all the chains as the number of iteration $M$ of main chain is large.
-Thus here is a guideline as a compromise: it suffices to check only one auxiliary chain of the last iteration of the main chain.
-Second reason comes from the outputs' form of the chains. They are not vectors in $\mathbb{R}^p$ but network structures. 
-Since we cannot draw traceplots of these network sequences, we need other ways for the diagnostics on them.
-One way is calculateing network statistics $s(.,.)$ in the model and drawing traceplots of them.
-On the traceplots, check whether each network statistics converges or not.
-In summary, for auxiliary chain,
+One reason for the difficulty is that there are too many auxiliary chains in our whole procedure. 
+It is practically impossible to check all the auxiliary chains as the number of main chain iterations $M$ becomes large. 
+Thus here is a guideline as a compromise: it suffices to check only one auxiliary chain of the main chain's last iteration.
+
+The second reason comes from the outputs' form of the auxiliary chains. 
+They are not vectors in $\mathbb{R}^p$ but network structures. 
+Since we cannot draw trace plots of these network sequences, we need other ways to diagnostics. 
+One way is to calculate network statistics $s(.,.)$ in the model and 
+draw trace plots of them instead of the network sequence itself. 
+On the trace plots, we can check whether each network statistic in the model converges or not. 
+The result can be regarded as the corresponding check result of the convergence of the networks' sequence.
+
+In summary, for the auxiliary chains,
 \begin{itemize}
     \item Calculate network statistics of all graphs produced by the last auxiliary chain.
-    \item Depict traceplots of each statistics to check the convergence and the mixing.
+    \item Depict trace plots of statistics in the model to check the convergence and the mixing.
 \end{itemize}
-If the network statistics do not converges, raise the number of iteration $K$ of each auxiliary chain and
-run the fitting algorithm again.
-
-Here are some tips for remedy when facing a problem.
-If parameters explode to infinity the number of iteration of main chain becomes large,
-check whether the number of iteration of exchange chain is too small.
-Moreover, take account of the model itself. 
-If you put linearly dependent network statistics simultaneously in the model,
-the parameter may not be identifiable. If you set the model with network statistics of hugh different scale but
-similar MCMC proposal distribution, the MCMC chain tends to fail to converge. For example,
-if you put k-star of specific order and edge count to the model at the same time but use
-same proposal like $N(0,\sigma^2)$ for some $\sigma^2>0$, MCMC chain does not converge easily
-because the parameter of edge count should be changed by relatively wide range even if the parameter of k-star moves slightly.
-Finally, the convergence of the BSTERGMs is very sensitive to initial point of main chain.
-If the main chain starts at the region of degeneracy in the parameter space, it is a hopeless one to converge the chain properly.
-
-\subsection{Model inference}
+If the network statistics do not converge, raise the number of iterations $K$ of each auxiliary chain,
+and rerun the fitting algorithm.
+
+Here are some remedy tips when facing a problem.
+If parameters explode to infinity as the number of iteration of the main chain becomes large, 
+check whether the number of iteration of the exchange chain is too small. 
+Picking the exchange sample before an auxiliary chain converges may disturb the main chain. 
+Moreover, you have to take account of the combination of network statistics in the model. 
+If you put a set of linearly dependent network statistics simultaneously in the model, 
+the parameter may not be identifiable. 
+If you set the model with network statistics with very different sensitivity for adding or erasing an edge but similar MCMC proposal distribution, 
+the MCMC chain tends to fail to converge. 
+For example, if you put k-star of specific order and edge count to the model at the same time 
+but use the same gaussian proposal with the same variance for both parameter, 
+the MCMC chain does not converge easily because the parameter of edge count should be changed by relatively wide range 
+even if the parameter of k-star moves slightly. 
+Finally, the convergence of the BSTERGMs is very sensitive to the initial point of the main chain. 
+If the main chain starts at the region of degeneracy in the parameter space, 
+it is hopeless for the chain to converge appropriately.
+Thus, I recommend running many test chains with low iterations of the main chain 
+but lots of iterations of the auxiliary chains 
+to determine the number of the auxiliary chain iterations and good initial points.
+
+\section{Model inference}
 
@@
 \begin{itemize}
-    \item A outlining shape of the posterior of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histogram.
-    \item An approximated summary statistics: mean, mode, variance and so on.
-    \item An approximated $p$-th quantile and probability interval
+    \item a outlining shape of the posteriors of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histograms.
+    \item an approximated summary statistics: mean, mode, variance and so on.
+    \item an approximated $p$-th quantile and probability interval
 \end{itemize}
 
-The interpretation of the parameter is very easy because of the separated formation and dissolution model.
-Since we cannot know the normalizing constant in the models, but we get an odds very easily from the model definition.
-And, from parameter value, we can expect the direction or tendency of network structure with high probability.
-If $\theta^+$ is higher, a network with high $s$ have high probability at the formation model, and vice versa.
-Likewise, If $\theta^-$ is lower, a network with lower $s$ have high probability at the dissolution model, and vice versa.
-Thus, if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $\infty$, the network does not change easily over time.
+The interpretation of the parameter is straightforward because of the separated formation and dissolution model.
+Since we cannot know the normalizing constant in the models, but we get odds quickly from the model definition.
+From parameter value, we can expect the direction or tendency of network structure with high probability.
+If $\theta^+$ is high, a network with high $s$ has high probability at the formation model, and vice versa.
+Likewise, If $\theta^-$ is low, a network with low $s$ has high probability at the dissolution model, and vice versa.
+Thus, if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $\infty$, the network does not change quickly over time.
 If $\theta^+$ go to $\infty$ and $\theta^-$ go to $-\infty$, the network changes drastically over time.
@@
 
-Let us turn to the prediction method. 
-we already have a generative algorithm for a network at the specific parameter point.
-In fact, it is same to the algorithm used in the auxiliary chain.
-Thus, we can predict the form of network at $T+1,T+2,...$ using posterior sample following the standard Bayesian prediction method.
+Let us turn to the prediction method. We already have a generative algorithm for a network at the specific parameter point.
+It is the same as the algorithm used in the auxiliary chain.
+Thus, we can predict the network's form at $T+1,T+2,...$ by the standard Bayesian prediction method using the MCMC posterior samples.
 For example, for predicting networks structure at $T+1$,
 \begin{itemize}
-    \item Run K iteration using auxiliary chain at $y_T$ with each posterior sample points.
-    \item Take the each network as a predicted result.
+    \item Run K iteration using an auxiliary chain at $y_T$ with each posterior sample points.
+    \item Take each network as a predicted result.
 \end{itemize}
@@
 
-\subsection{Goodness of Fit}
+\section{Goodness of Fit}
 To evaluate the goodness of fit for the BSTERGMs and to make sure the posterior $\theta^+,\theta^-$ is fitted properly,
@@
 
-\section{Example: UCLA medical working project - team science data}
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\chapter{Example: UCLA medical working project - team science data} \label{Chapter6}
 %data description
-As an application, I consider the Bayesian separable temporal exponential family model with UCLA medical working project's team science data.
-These data consist of three sequence,
-and each sequence has three time points and for each time,
-there is a network with 52 nodes corresponding 29 scholar members and 23 faculty members.
+As an application, I consider the Bayesian separable temporal exponential family model 
+with UCLA medical working project's team science data.
+These data consist of three sequences, and each sequence has three-time points, 
+and for each time, there is a network with 52 nodes corresponding 29 scholar members and 23 faculty members.
 
@@
     \begin{center}
-    \includegraphics[scale=0.23]{pictures/m1_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/w1_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/f1_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/m1_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/w1_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/f1_19_nework.png}
     \caption{Sequence 1}
@@
     \begin{center}
-    \includegraphics[scale=0.23]{pictures/m2_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/w2_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/f2_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/m2_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/w2_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/f2_19_nework.png}
     \caption{Sequence 2}
@@
     \begin{center}
-    \includegraphics[scale=0.23]{pictures/m3_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/w3_19_nework.png}
-    \includegraphics[scale=0.23]{pictures/f3_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/m3_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/w3_19_nework.png}
+    \includegraphics[scale=0.26]{pictures/f3_19_nework.png}
     \caption{Sequence 3}
@@
 
-I consider 2 dimentional model as follows.
+I consider the 2-dimensional BSTERGM as follows.
 \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
@@
 and set the proposal distribution of main chain $\epsilon\sim\mathcal{N}(0,0.0004)$.
-The auxiliary chains consist of 1000 iterations and the main chain consists of 50000 iteration.
-In this setting, the BSTERGM fitting algorithm run 2400 iterations of main chain per a minute on the intel i7-8700(3.20GHz) CPU.
-
-After generating the chain, I cut first 10000 samples for a burn-in period and thin the sample by 10.
-The posterior estimates for each network and histograms are follows.
+The auxiliary chains consist of 1000 iterations, and the main chain consists of 50000 iterations.
+In this setting, the BSTERGM fitting algorithm runs 2400 iterations of the main chain per minute on the intel i7-8700(3.20GHz) CPU.
+
+After generating the chain, I cut the first 10000 samples for a burn-in period and thin the sample by 10.
+The posterior estimates for each network and histograms are as follows.
+
 %mean, sd
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_histogram.png}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_histogram.png}
-    \caption{Sequence 1, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 1, posterior histogram, left: formation, right: dissolution}
     % \label{fig:ls}
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_histogram.png}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_histogram.png}
-    \caption{Sequence 2, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 2, posterior histogram, left: formation, right: dissolution}
     % \label{fig:ls}
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_histogram.png}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_histogram.png}
-    \caption{Sequence 3, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_formation_histogram.png}
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_dissolution_histogram.png}
+    \caption{Sequence 3, posterior histogram, left: formation, right: dissolution}
     % \label{fig:ls}
@@
 
-For the diagnostics, I attach traceplots and autocorrelation plots.
-The autocorrelation of each chain decrease rapidly as lag increases.
-Moreover, traceplots shows that the MCMC samples are mixed well.
+For the diagnostics, I attach trace plots and autocorrelation plots.
+The autocorrelation of each chain decreases rapidly as lag increases.
+Moreover, trace plots show that the MCMC samples are mixed well.
+
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_traceplot.png}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_traceplot.png} \\
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_formation_acf.png}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_BSTERGM_dissolution_acf.png}
-    \caption{Sequence 1, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.395]{pictures/net1seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 1, diagnostics plots, left: formation, right: dissolution}
     % \label{fig:ls}
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_traceplot.png}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_traceplot.png} \\
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_formation_acf.png}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_BSTERGM_dissolution_acf.png}
-    \caption{Sequence 2, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.395]{pictures/net2seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 2, diagnostics plots, left: formation, right: dissolution}
     % \label{fig:ls}
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_traceplot.png}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_traceplot.png} \\
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_formation_acf.png}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_BSTERGM_dissolution_acf.png}
-    \caption{Sequence 3, left: formation, right: dissolution}
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_formation_traceplot.png}
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_dissolution_traceplot.png} \\
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_formation_acf.png}
+        \includegraphics[scale=0.395]{pictures/net3seq_chain1_BSTERGM_dissolution_acf.png}
+    \caption{Sequence 3, diagnostics plots, left: formation, right: dissolution}
     % \label{fig:ls}
@@
 \clearpage
+
 In addition, as a procedure of diagnostics for auxiliary chains,
-here are traceplots of the number of edges in last auxiliary chain.
-They shows that all chains are not degenerated at last parameter point during the iterations of auxiliary chain.
+here are trace plots of the number of edges in the last auxiliary chain.
+They show that all chains do not degenerate at the last parameter point during auxiliary chain iterations.
 
@@
     \begin{center}
-        \includegraphics[scale=0.23]{pictures/net1seq_chain1_lastsampler_num_edges.png}
-        \includegraphics[scale=0.23]{pictures/net2seq_chain1_lastsampler_num_edges.png}
-        \includegraphics[scale=0.23]{pictures/net3seq_chain1_lastsampler_num_edges.png}
-    \caption{Sequence 1, left: formation, right: dissolution}
+        \includegraphics[scale=0.26]{pictures/net1seq_chain1_lastsampler_num_edges.png}
+        \includegraphics[scale=0.26]{pictures/net2seq_chain1_lastsampler_num_edges.png}
+        \includegraphics[scale=0.26]{pictures/net3seq_chain1_lastsampler_num_edges.png}
+    \caption{The number of nodes of exchange samples in the last auxiliary chains,\\left: Sequence 1, middle: Sequence 2, right: Sequence 3}
     % \label{fig:ls}
@@
 
-Finally, I append the results of goodness of fit.
-Compare observed (true) value with quantiles of simulated values of parts of sufficient network statistics,
-generated by the goodness of fit algorithm.
-Since this example has simply just two edge terms, and since the BSTERGMs share same parameter
-for each network statistics but values of network statistics are different at each time point,
-the sufficient statistics do not fitted perfectly to observed networks' values.
+Finally, I append the results of the goodness of fit check.
+Compare the observed value on the data with simulated values' quantiles generated by the goodness-of-fit algorithm. 
+Parts of sufficient statistics are selected as the network statistics in the goodness of fit procedure.
+Since this example model merely has just two edge terms, 
+and since the BSTERGMs share the same parameter for each network statistics but values of network statistics are different at each time point, 
+the sufficient statistics do not fit perfectly to observed networks' values. 
 However, the tendency of simulated values goes with the observed values.
@@
 
-%GOF table
-
-\section{Summary and Discussion}
-In this paper, I have proposed the Bayesian STERGMs for a sequence of networks
-and a framework for statistical estimation and inference using the standard Bayesian approach for the model.
-I also have shown a application of the BSTERGMs. %more add
-
-Since the models are based on ERGMs, it can be extended to more general networks
-including a variety of network statistics.
-Moreover, the model can be applied to weighted graph and use the weights as covariates of the model.
-Plus, the BSTERGMs are using the exchange algorithm, so it is flexible enough to be integrated
-with other methods to help the MCMC chains to converge and run faster.
-For example, the 'tie no tie'(TNT) algorithm used by ergm package (\cite{RN100}) can be applied to
-enhance the performance of the auxiliary chain of the BSTERGMs.
-
-A modeling by the BSTERGMs for large networks is possible,
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+\chapter{Summary and Discussion}\label{Chapter7}
+\section{Summary and discussion}
+
+In this paper, I have proposed the Bayesian STERGMs for statistical analysis on a sequence of networks. 
+I also provide the algorithm to estimate the model's parameters and the inference method 
+following the standard Bayesian approach. 
+Moreover, an application to real data is shown using BSTERGMs.
+
+Since the BSTERGMs have their base on ERGMs, the models are flexible and can include various network statistics. 
+Moreover, the BSTERGMs can be applied to more general networks, including weighted graphs. 
+Moreover, by combining these features, the BSTERGM can take the weights as covariates of the model. 
+On the other hand, since the BSTERGMs are using the exchange algorithm, 
+the models can be integrated with other methods to help the MCMC chains to converge and run faster. 
+For example, the 'tie no tie'(TNT) algorithm used by ergm package (\cite{RN100}) can be applied 
+to enhance the performance of the auxiliary chain of the BSTERGMs.
+
+Modeling by the BSTERGMs for large networks is possible, 
 but a problem of computational inefficiency rises. 
-If given observed network has lots of nodes, the estimation algorithm for the BSTERGM needs 
-huge number of auxiliary chain itertation to sample a proper exchange network at a given parameter point.
-In addition, the calculation of network statistics in the model cause another problem for large networks,
-especially when the model has network statistics that high complexity of computation.
-For example, statistics based on edgewise shared partner distributions needs $O(n^3)$,
+If the observed networks in the given sequence have lots of nodes, 
+the estimation algorithm for the BSTERGM needs a vast number of iterations in the auxiliary chains 
+to draw a proper exchange sample at a given parameter point. 
+In addition, the calculation of network statistics in the model causes another problem for large networks, 
+especially when the network statistics demand a high complexity of computation. 
+For example, calculating the edgewise shared partner distribution of one network 
+needs computation of complexity $O(n^3)$, 
 so computation cost increases rapidly as the number of nodes $n$ grows and 
-the auxiliary chains become bottlenecks of the whole computation procedure.
-
-Another important issue related in the BSTERGMs is degeneracy as I have mentioned at the section of model diagnostics.
-The BSTERGMs are very sensitive to the initial point of main MCMC chain, so
-searching a region for non-degenerate initial parameter point in the parameter space is very important.
-Often, however, the work finding the region is too difficult for the BSTERGMs,
-partly because of the high number of parameters in the models.
-Since the BSTERGMs has two parts for $\theta^-, \theta^+$ in the setting of the models by the their definition,
-the dimension of parameter space increases easily although each formation and dissolution model does not have
-too many terms respectively.
-What makes the situation worse is not only the high dimension of parameter space but also the shape of convergent region of the parameter.
-In many case, the region is geometrically curved due to an implicit relation between several network statistics in the models
-and interaction between the formation model and the dissolution model.
-Thus, the MCMC algorithm may fail to exploring the whole parameter space, especially when the space has high dimension,
-and suffer non-convergence problem.
-Imposing some model level constraint or using special prior may be a remedy,
-but finding the proper constraint or the prior distribution will often be a more difficult problem than original one.
+the calculating $\pi$ become bottlenecks of the whole computation procedure.
+
+Another critical issue related to the BSTERGMs is degeneracy, as I have mentioned in chapter 5. 
+The BSTERGMs are very sensitive to the main MCMC chain's initial point, 
+so searching a region for non-degenerate initial parameter point in the parameter space is very important. 
+Often, however, finding the region is tricky work for the BSTERGMs, partly because of the too many parameters. 
+Since the BSTERGMs have two parts for $\theta^-, \theta^+$ in the setting of the models by definition, 
+the parameter space dimension tends to increase, 
+although each formation and dissolution model does not have too many terms, respectively. 
+What makes the situation worse is not only the high dimension of parameter space but also the shape of the convergent region in the space. 
+In many cases, the region is geometrically curved 
+due to an implicit relation between several network statistics in the models and interaction between the formation and dissolution models. 
+Thus, the MCMC algorithm may fail to explore the whole parameter space, 
+especially when the parameter space has a high dimension.  
+As a result, the MCMC chain suffers a non-convergence problem. 
+Imposing some model level constraint or using a special prior may be a remedy, 
+but finding the proper constraint or the prior distribution will often be a more difficult problem than the original one. 
 Thus, further research is needed to resolve these issues.
 
-\begin{frame}{Supplements}
+\section{Supplements}
 You can find the C++ implementation (using Armadilo: see http://arma.sourceforge.net/)
-    \\ of BSTERGM fitting, diagnostic, and GOF algorithms
-    \\ at my Github page: https://github.com/letsjdosth/BayesianSTERGM.
-\end{frame}
+of BSTERGM fitting, diagnostic, and GOF algorithms
+at my Github page: \\
+https://github.com/letsjdosth/BayesianSTERGM.
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
+
+%--------------------------------------------------------------------------------%
+%--------------------------------------------------------------------------------%
 
@@
 \bibliographystyle{apalike}
-\bibliography{doc.bib}
+\bibliography{paper_format}
+
+% %----------------------------------------------------------
+\newpage
+
+\section*{국문 초록}
+\setstretch{1.8}
+\par
+\bigskip
+
+\begin{center}
+\large
+BSTERGMs: 베이지안 분해 가능-시간 지수 랜덤 그래프 모형 
+\end{center}
+\setstretch{1.8}
+\par
+\bigskip
+\noindent
+분해 가능-시간 지수 랜덤 그래프 모형(STERGMs)은 시간에 따라 변화하는
+동적 네트워크의 구조 변화를 묘사하는 데에 유용한 모형이나,
+이 모형에 대한 베이지안 접근은 지금가지 이루어지지 않았었다.
+이 논문에서는, STERGMs에 대한 베이즈 통계학적 접근인 베이지안 분해 가능-시간 지수 랜덤 그래프 모형(BSTERGMs)을 제안하고,
+이중적으로 추적 불가능한 정규화 상수 문제를 고려하며 사후 분포로부터 표본을 얻는
+계산 알고리즘을 설명한다. 나아가, 이 모형에 대해 베이지안 방법을 통한 통계적 추론 과정을 보인다.
+
+\bigskip
+\noindent \hrulefill\\
+핵심 용어 :  \par 네트워크, 지수 랜덤 그래프 모형, 베이지안 추론, 마르코프 연쇄 몬테 카를로, 시간 데이터
+
+%\pagebreak
+
+\newpage
+
+% %--------------------------------------------------------------------------------%
+% %--------------------------------------------------------------------------------%
 

### 2020-12-30 doc/paper_format.tex
revise doc

@@
 
-Nowadays, new kinds of data having special forms emerge in many fields.
-A series of real number values do not confine these new data as traditional data are in the past.
+Nowadays, new kinds of data with special forms emerge in many fields.
+A sequence of real number values do not confine these new data as traditional data are in the past.
 The data may have a non-number type of value, unique structure, or even integrated different types of values with model-specific relation.
@@
 
-As one form of complex data, networks, or random graphs get highlighted attention among complex data for statistical analysis.
-A relational structure can represent many phenomena in the real world, and the structure can be expressed effectively by a form of a network. 
+As one form of complex data, network, or random graph gets highlighted attention as a flexible modeling tool for statistical analysis.
+Many phenomena in the real world can be represented effectively by a form of networks. 
 For example, social networks, including offline and online, can be thought of as networks.
-In this view, some social aspects of the relationship between actors can be modeled under an object of social science, 
-including an analysis of power structure and a summary of intimacy structure on a given group.
-Not only that, application to a transmitted disease in epidemiology enables statistical modeling based on real data.
-
-The ability to capture distinct features on the observed network enlarges statistical models' usefulness on networks. 
+Under an specific object of social science, some aspects of the relationship among actors 
+can be modeled by networks, such as an analysis of power structure or a summary of intimacy structure on a given group. 
+Another example is an application to a transmitted disease in epidemiology.
+Modeling an infection using networks enables statistical analysis based on real data.
+
+The ability to capture distinct features on an observed network enlarges statistical models' usefulness on networks. 
 The statistical models give information about local relations like formation or dissolution between two actors 
@@
 The statistical models investigate the complex behavior of networks, 
-including how an observed network's global features may be related to local network structure. 
+including how the observed network's global features may be related to local network structure. 
 During the investigation, the statistical models can capture existing dependency on a given network and provide a tool 
@@
 and \cite{RN115} and \cite{RN116} started the Bayesian approaches of the models.
-In the ERGMs, The dependencies are captured using a series of local features called network statistics from classical graph-theory. 
+In the ERGMs, the dependencies are captured by a series of local features called network statistics from classical graph theory. 
 Since these models play an important role throughout this paper, I will introduce them more precisely in the following chapter.
@@
 The first exploration applying ERGMs to a network sequence was \cite{RN109}. 
-After them, a notable contributions were of \cite{RN125} and \cite{RN126}, the Temporal ERGMs(TERGMs). 
-In the TERGMs, ERGMs are used by modeling the transition probability between times. 
-Moreover, \cite{RN93} extend the TERGMs to the Separable Temporal ERGMs(STERGMs), 
+After them, a notable contributions were of \cite{RN125} and \cite{RN126}, the Temporal ERGMs (TERGMs). 
+In the TERGMs, the ERGMs are used by modeling the transition probability between times. 
+Moreover, \cite{RN93} extend the TERGMs to the Separable Temporal ERGMs (STERGMs), 
 considering the separation between duration - how long ties tend to last - and 
-incidence - the rate at which new ties are formed -from 
+incidence - the rate at which new ties are formed - from 
 prevalence - for example, the total number of ties - of the network properties. 
 By imposing a conditional independence condition between the duration and incidence on the model, 
-the models can be extended to more effectively interpretable ones using similar fitting methods with TERGMs 
+the models can be extended to more effectively interpretable ones than the TERGMs 
 without adding much complexity.
@@
 a Bayesian approach for the STERGMs has not been conducted. 
-It is partly because of the hardship from the massive computation in a parameter estimation method for the models. 
+It is partly because of the hardship from the massive computation in the parameter estimation method for the models. 
 Nonetheless, the Bayesian approach of STERGMs provides the statistical inference method following the standard Bayesian way, 
 so the approach has a huge advantage of convenience for statistical works for fitted parameters and predictions to next time network structure. 
-Thus, in this paper, I specify the bayesian separable temporal exponential random graph models(BSTERGMs) and propose a fitting algorithm for the model parameter.
-I also discuss ways of inference, the diagnostics procedure, and checking the goodness of fit.
+Thus, in this paper, I specify the Bayesian separable temporal exponential random graph models (BSTERGMs) and propose a fitting algorithm for the model parameters.
+I also discuss the inference methods, the diagnostics and goodness of fit checking procedures.
 
@@
 from the UCLA medical working project team in chapter 6, 
-Finally, I discuss some possible extensions and ways to supplement the BSTERGMs further for future research.
+Finally, in chapter 7, I discuss some possible extensions and ways to supplement the BSTERGMs further for future research.
 
@@
 describing relations among subjects in the statistical context. 
-To begin with, I will briefly summarize some terminologies and notations for concepts related to the networks.
-
-I will use two words, 'graphs' and 'network,' as the same meaning without distinction. 
+To begin with, I briefly summarize some terminologies and notations for concepts related to the networks.
+
+In this paper, I will use two words, 'graphs' and 'network,' as the same meaning without distinction.
 We say each subject on the data to a node. 
@@
 
-For simplicity, I will deal with a situation without weight on each node or edge.
+For simplicity, I will deal with a situation without weights on each node or edge.
 Then, a graph $y$ can be represented by a more straightforward form.
@@
 Then, for $\mathcal{B}=\{0,1\}$,
-the set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes.
+a set $\mathcal{Y} \subset \mathcal{B}^{n^2}$ is a set of graphs of $n$ nodes.
 Now, a graph $y \in \mathcal{Y}$ can be represented by a $n \times n$ binary matrix
@@
 We say $Y: \Omega \to \mathcal{Y}$ is a random variable for a graph, or a random graph.
-For a random graph $Y \in \mathcal{Y}$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$
+For a random graph $Y$, denote the edge between i-th node and j-th node by $Y_{ij}$ for $i,j=1,2,...,n$
 satisfying $Y_{ij}=1$ if the edge is connected. Otherwise, $Y_{ij}=0$.
-Note that the $Y_{ij}$s are also random variables.
+Note that the $Y_{ij}$s are also random variables in our context.
 
@@
 \section{ERGMs: Exponential random graph models}
-One of the natural questions from the definition of the random graph is 
-how we can model a probability distribution on $\mathcal{Y}$.
+One of the natural questions from the definition of the random graphs is 
+how we can model a probability distribution to the graphs.
 As a answer, the Exponential random graph models (ERGMs) (\cite{RN123} and \cite{RN122})
 gives a classical but flexible scheme to model for time-invariant networks represented as a single directed or undirected graph.
-Moreover, ERGMs provide a statistical approach to the systematic exploration of several local features simultaneously and 
+Moreover, the ERGMs provide a statistical approach to the systematic exploration of several local features simultaneously and 
 how these local structures interact to form a global network.
@@
 Specifically, following the notation of the previous section,
-let $\mathcal{Y}$ be a set of random graphs with $n$ nodes and $Y \in \mathcal{Y}$ be a random graph.
-Set a distribution on $\mathcal{Y}$ to
+let $\mathcal{Y}$ be a set of networks with $n$ nodes and $Y$ be a random graph.
+Set a distribution of $Y$ to
     \[P(Y=y;\theta) = \frac{exp(\theta^{T}s(y))}{c(\theta)}\]
 for some $\theta\in\mathbb{R}^p$,
-where $s(y)\in\mathbb{R}^p$ is a vector which is part of y's sufficient network statistics,
+where $s(y)\in\mathbb{R}^p$ is a vector which is part of sufficient network statistics of $y\in\mathcal{Y}$,
 and $c(\theta)\in\mathbb{R}$ is a normalizing constant satisfying $c(\theta)=\sum_{y\in\mathcal{Y}}exp(\theta^{T}s(y))$.
@@
 we are interested in the dynamics or evolution of the network sequence.
-The temporal Temporal Exponential family Random Graphs Models(TERGMs)
+The Temporal Exponential family Random Graphs Models(TERGMs)
 proposed by \cite{RN125} and \cite{RN126}
-is a model for our interest if we permit a Markov assumption
+are models for our interest if we permit a Markov assumption
 \[P(Y_2,Y_3,...,Y_T|Y_1)=P(Y_2|Y_1)P(Y_3|Y_2)...P(Y_T|Y_{T-1})\]
 and focus on the transition probability along index of the sequence.
-For convenience, I refer to the index as a time.
+For convenience, I refer to the index $1,2,...,T$ as a time.
 
 Specifically, let $\mathcal{Y}$ be a set of graphs with $n$ nodes. 
-Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs in $\mathcal{Y}$.
-Set a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds) to
+Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs.
+Set a distribution of $Y_2,...,Y_T$ by
 \[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta) = \frac{exp(\theta^{T}s(y_t, y_{t-1}))}{c(\theta, y_{t-1})}\]
-where $t=2,...,T$, $\theta\in\mathbb{R}^p$,
+where $t=2,...,T$, $y_t\in\mathcal{Y}$, $\theta\in\mathbb{R}^p$,
 sufficient network statistics $s(y_t, y_{t-1})\in\mathbb{R}^n$,
@@
 One solution to remedy the problems of TERGMs is 
-the Separable-Temporal Exponential family Random Graphs Models, or simply STERGMs. %ref
+the Separable-Temporal Exponential family Random Graphs Models, or simply STERGMs (\cite{RN93}.)
 The key idea is to separate the formation models and dissolution model
 for decoupling an effect on edge's duration of the model parameter from edge's incidence
-when setting a distribution on $\mathcal{Y}\times ... \times \mathcal{Y}$ ($T-1$ folds, $T\in\mathbb{N}$),
-where $\mathcal{Y}$ is a set of graphs with given $n$ nodes.
+when setting a distribution on a sequence of random graph.
 
 Precisely,
-Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (on $\mathcal{Y}$).
+let $\mathcal{Y}$ is a set of graphs with given $n\in\mathbb{N}$ nodes. 
+Suppose $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T (T\in\mathbb{N})$ be random graphs.
 Let $\mathcal{Y}^+|_t$ be a subset of $\mathcal{Y}$ consisting of all graphs which have equal or additional edges comparing to $y_{t-1}$.
 Likewise, let $\mathcal{Y}^-|_t$ be a subset of $\mathcal{Y}$ consisting all graphs which have equal or sparse edges comparing to $y_{t-1}$.
-Next, for $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
+Next, for random graphs $Y_t^+: \Omega \to\mathcal{Y}^+|_t$, $Y_t^-: \Omega \to\mathcal{Y}^-|_t$,
+and graphs $y_t^+ \in \mathcal{Y}^+|_t$, $y_t^- \in \mathcal{Y}^-|_t$, set
 \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
 \[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
-for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
+for parameters $\theta^+,\theta^-\in\mathbb{R}^p$, parts of sufficient network statistics $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^p$,
 and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
-and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
+$c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
 
@@
 by $1+1=1, 1+0=0+1=1, 0+0=0$ and $1-1=0, 1-0=1, 0-1=0, 0-0=0$) 
-set the combined network $y_t$ to
+and set the combined network $y_t$ to
 \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
@@
 \[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
-and the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$, called the separability.
+and the conditional independence between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$.
+Note that the last condition is called the separability.
 Then we get
-\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+\begin{multline}
+    P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-) \\
+    = P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)
+\end{multline}
 Models that have such a form are called the STERGMs.
@@
 reflect the structure of given network without identifying each node index.
-Typical candidates of the statistics are here.
+Typical candidates are here.
 
@@
 For a given $i, 1\leq i \leq n-1$, $D_i(y)$ is defined as the number of nodes in y whose the number of edges incident to the node equals $i$, 
-and called $i$-th order node degree distribution.
+and called the $i$-th order node degree distribution.
 
@@
 in y whose $j$ and $k$ have nodes connected by an edge each other and they share exactly $i$ connected nodes in common,
-and called $i$-th order edgewise shared partner distribution.
-
-Note that, the bundle of all order of node degree distribution and all order of edgewise shared partner distribution,
+and called the $i$-th order edgewise shared partner distribution.
+
+Note that the bundle of all order of the node degree distribution and all order of the edgewise shared partner distribution,
 \[(D_1(y),D_2(y),...,D_{n-1}(y),P_1(y),P_2(y),...,P_{n-2}(y))\] 
 forms sufficient statistics of a graph.
-And there are relation that
+Moreover, there are relation that
 \[n=D_0(y)+D_1(y)+...+D_{n-1}(y)\]
@@
 A function of the sufficient statistics can be candidates for the model statistics.
-\cite{RN128} proposed graph statistics including $k$-star $S_k(y)$ for $1\leq k \leq n-1$ and $k$-triangles $T_k$ for $1\leq k \leq n-2$.
+\cite{RN128} proposed graph statistics including the $k$-star $S_k(y)$ for $1\leq k \leq n-1$ and the $k$-triangles $T_k$ for $1\leq k \leq n-2$.
 The $k$-star consists of a node together with a set of $k$ of its connected nodes.
@@
 
-Another interesting statistics are geometrically weighted value of distributions defined by
+Another interesting statistics are the geometrically weighted value of distributions defined by
 \cite{RN118}, \cite{RN104}, \cite{RN128} and \cite{RN103}.
-For example, geometrically weighted (node) degree with parameter $\tau$ is defined by
+For example, the geometrically weighted (node) degree with parameter $\tau$ is defined by
 \[GWD(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)D_k(y)\]
-and geometrically weighted shared partner distribution with parameter $\tau$ is defined by
+and the geometrically weighted shared partner distribution with parameter $\tau$ is defined by
 \[GWESP(y;\tau)=e^{\tau} \sum_{k=1}^{n-2} (1-(1-e^{-\tau})^k)P_k(y)\]
@@
 Another common choice is using a difference, $s(y_t,y_{t-1})=s'(y_t)-s'(y_{t-1})$ where $s'$ is a network statistics of given network.
-Although there are many other statistics related to ERGMs, I omit them and end this section here.
-For additional example, see \cite{RN107} for the ERGMs, and \cite{RN125}'s chapter 2.1 for temporal the ERGMs.
+Although there are many other statistics related to the ERGMs, I omit them and end this section here.
+For additional example, see \cite{RN107} for the ERGMs, and \cite{RN125}'s chapter 2.1 for the TERGMs.
 
@@
 With priors $p(\theta^+),p(\theta^-)$ over $\theta^+,\theta^-$,
-the Bayesian model for the STERGMs can be made. I call this model Bayesian Separable Temporal Exponential family Random Graph Models,
+the Bayesian model for the STERGMs can be made. I call this model the Bayesian Separable Temporal Exponential family Random Graph Models,
 or the BSTERGMs.
 
-More explicitly, for given the number of nodes $n$, let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T$ be random graphs (of $\mathcal{Y}$).
+More explicitly, let $\mathcal{Y}$ be the set of all networks with $n\in \mathbb{N}$ nodes. Let $Y_1=y_1 \in \mathcal{Y}$ be given and $Y_2,...,Y_T, T\in\mathbb{N}$ be random graphs.
 As the STERGMs, define $\mathcal{Y}^+|_t$, $\mathcal{Y}^-|_t$ be subsets of $\mathcal{Y}$ consisting all graphs which have equal or additional edges comparing to $y_{t-1}$,
-and which have equal or sparse edges comparing to $y_{t-1}$ respectively.
-Then, for $Y_t^+$ of $\mathcal{Y}^+|_t$ and $Y_t^-$ of $\mathcal{Y}^-|_t$, $y_t^+ \in \mathcal{Y}^+|_t$ and $y_t^- \in \mathcal{Y}^-|_t$, set
+and which have equal or sparse edges comparing to $y_{t-1}$, respectively.
+Then, for random graphs $Y_t^+: \Omega \to \mathcal{Y}^+|_t$, $Y_t^-: \Omega \to \mathcal{Y}^-|_t$,
+and graphs $y_t^+ \in \mathcal{Y}^+|_t$, $y_t^- \in \mathcal{Y}^-|_t$, set formation model
 \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+and dissolution model
 \[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
-for some $\theta^+,\theta^-\in\mathbb{R}^p$, $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^n$, which are parts of sufficient network statistics,
-and normalizers $c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))$, 
-and $c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))$.
-
-Then, with the edgewise boolean algebra, set the combined network $y_t$ to
+for parameters $\theta^+,\theta^-\in\mathbb{R}^p$, network statistics $s(y_t^+, y_{t-1}), s(y_t^-, y_{t-1})\in\mathbb{R}^p$,
+and normalizers 
+\[c(\theta^+, y_{t-1})=\sum_{y^+\in\mathcal{Y}^+}exp((\theta^+)^{T}s(y^+, y_{t-1}))\]
+\[c(\theta^-, y_{t-1})=\sum_{y^-\in\mathcal{Y}^-}exp((\theta^-)^{T}s(y^-, y_{t-1}))\]
+
+Then, with the edgewise boolean algebra defined in chapter 2.5, set the combined network $y_t$ to
 \[y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})\]
@@
 \[P(Y_2,...,Y_T|Y_1)=P(Y_2|Y_1)...P(Y_T|Y_{T-1})\]
-and the conditional independence(separability) between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$.
+and separability (or, conditional independence) between $Y_t^+$ and $Y_t^-$ for all $t=2,...,T$.
+
 Then we get
-\[\theta^+ \sim p(\theta^+), \theta^- \sim p(\theta^-)\]
-\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+\[\theta^+ \sim p(\theta^+)\]
+\[\theta^- \sim p(\theta^-)\]
+and
+\begin{multline}
+    P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-) \\ 
+    =P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)
+\end{multline}
 
@@
 where $y_t=y_t^+ - (y_{t-1} - y_t^-) = y_t^- + (y_t^+ - y_{t-1})$ as set above.
-I define models that have such forms as the STERGMs.
+I define models that have such forms as the BSTERGMs.
 
 Note that the BSTERGMs also have similar kinds of problems to the BERGMs.
-The normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically cannot be computed because we need to sum up too many terms.
-The constants are doubly intractable: they depend on $\theta^+$,$\theta^-$, the parameters of a model.
-Thus, we cannot use a standard MCMC algorithm to get the posterior sample in this model,
+The normalizing constants $c(\theta^+,y_{t-1})$, $c(\theta^-,y_{t-1})$ practically cannot be computed 
+because there are too many terms to sum up.
+The normalizing constants are doubly intractable since they depend on $\theta^+$ and $\theta^-$, the parameters of a model.
+Thus, we cannot use a standard MCMC algorithm to get the posterior sample of the parameters
 because the constant part remains when calculating the MCMC ratio.
-I will address this issue in the next chapter.
+I will address this issue with more details in the next chapter.
 
@@
 
-Suppose that a sequence of graph samples $y_1,...,y_T$ is observed.
+Suppose that a sequence of graphs $y_1,...,y_T$ in $\mathcal{Y}$ is observed.
 (then, $y_2^+,...,y_T^+$ and $y_2^-,...,y_T^-$ are uniquely determined.) 
-How can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using the BSTERGMs?
-To solve the doubly intractable constant problem, I propose a novel algorithm using the exchange algorithm,
-one of the special MCMC techniques by \cite{RN127}, extending the algorithm of the BERGMs'(\cite{RN115}).
+Then how can we find the posterior distribution of $\theta^+,\theta^-$ with the observation using the BSTERGMs?
+In this chapter, I propose a novel algorithm to estimate parameters of the BSTERGMs using the exchange algorithm, 
+one of the special MCMC techniques by \cite{RN127}.
+
+The motivation of using the exchange algorithm comes from the algorithm of the BERGMs' (\cite{RN115})
+to solve the doubly intractable constant problem. By extending the algorithm of the BERGMs,
+I build the fitting algorithm for the BSTERGMs.
 Although an approach using MCMC cannot give an analytic expression of the posterior distribution, 
-it gives posterior samples as much as we want, and we can conduct statistical inference on given data using the samples.
-
-The key idea is running two MCMC chains to draw independent samples to cancel the constant parts
-when calculating the MCMC accept-reject ratio.
-As a motivation before introducing the new algorithm, consider the ordinary MCMC ratio for $m$-th MCMC iteration in the BSTERGMs context 
+it gives posterior samples as much as we want, so we can conduct statistical inference on given data using the generated samples.
+
+Before introducing the new algorithm,
+let us consider the ordinary MCMC ratio for $m$-th MCMC iteration in the BSTERGMs context
 to show the doubly intractable constant problem explicitly.
 
+For convenience, let the parameters of formation model and dissolution model are in $\mathbb{R}$.
 With symmetric proposal distribution,
@@
 where $\theta^*$ is proposed parameter in an iteration of MCMC.
-Plugging specific model form to above equation,
+Plugging the model form,
 \[\pi_{ordinary} = \frac{\frac{exp(\theta_*^+ s(y_t^+,y_{t-1}))}{c(\theta_*^+, y_{t-1})} \frac{exp(\theta_*^- s(y_t^-, y_{t-1}))}{c(\theta_*^-, y_{t-1})} p(\theta_*^+)p(\theta_*^-)}
@@
 
-If we can generate simulated samples $y_{ex}|y_{t-1},\theta_{m-1}$ and $y_{ex}|y_{t-1},\theta_*$ and
-calculate their probability following the BSTERGMs, this situation can be resolved.
-At this point, applying the key idea mentioned above, multiply the probability terms in order to exclude the constant terms.
-Then we get the exchange algorithm's ratio,
-\[\pi = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}
-{P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)}\]
-
-\[=\frac{\frac{exp(\theta_*^+ s(y_t^+,y_{t-1}))}{c(\theta_*^+, y_{t-1})} \frac{exp(\theta_*^- s(y_t^-, y_{t-1}))}{c(\theta_*^-, y_{t-1})} p(\theta_*^+)p(\theta_*^-)}
+The key idea of new algorithm to solve this problem is running two MCMC chains to draw independent samples to cancel the constant parts
+when calculating the MCMC accept-reject ratio. If we can generate simulated samples $y_{ex}|y_{t-1},\theta_{m-1}$ and $y_{ex}|y_{t-1},\theta_*$ and
+consider their probability following the BSTERGMs, this situation can be resolved.
+By multiplying the probability terms in order to exclude the constant terms,
+we get the exchange algorithm's ratio,
+
+\begin{equation}
+    \begin{aligned}
+        \pi & = \frac{P(y_t^+|y_{t-1},\theta_*^+)P(y_t^-|y_{t-1},\theta_*^-)p(\theta_*^+)p(\theta_*^-)P(y_{ex}^+|y_{t-1},\theta_{m-1}^+)P(y_{ex}^-|y_{t-1},\theta_{m-1}^-)}
+        {P(y_t^+|y_{t-1},\theta_{m-1}^+)P(y_t^-|y_{t-1},\theta_{m-1}^-)p(\theta_{m-1}^+)p(\theta_{m-1}^-)P(y_{ex}^+|y_{t-1},\theta_*^+)P(y_{ex}^-|y_{t-1},\theta_*^-)} \\
+        & = \frac{\frac{exp(\theta_*^+ s(y_t^+,y_{t-1}))}{c(\theta_*^+, y_{t-1})} \frac{exp(\theta_*^- s(y_t^-, y_{t-1}))}{c(\theta_*^-, y_{t-1})} p(\theta_*^+)p(\theta_*^-)}
         {\frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))}{c(\theta_{m-1}^+, y_{t-1})} \frac{exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}{c(\theta_{m-1}^-,y_{t-1})} p(\theta_{m-1}^+)p(\theta_{m-1}^-)}
         \frac{\frac{exp(\theta_{m-1}^+)s(y_{ex}^+,y_{t-1})}{c(\theta_{m-1}^+, y_{t-1})}\frac{exp(\theta_{m-1}^- s(y_{ex}^-,y_{t-1}))}{c(\theta_{m-1}^-, y_{t-1})}}
-{\frac{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))}{c(\theta_*^+,y_{t-1})} \frac{exp(\theta_*^- s(y_{ex}^-,y_{t-1}))}{c(\theta_*^-,y_{t-1})}}\]
-
-\[=\frac{exp(\theta_*^+ s(y_t^+,y_{t-1})) exp(\theta_*^- s(y_t^-, y_{t-1}))}{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}
-\frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_{ex}^-,y_{t-1}))}{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))exp(\theta_*^- s(y_{ex}^-,y_{t-1}))}
-\frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
+        {\frac{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))}{c(\theta_*^+,y_{t-1})} \frac{exp(\theta_*^- s(y_{ex}^-,y_{t-1}))}{c(\theta_*^-,y_{t-1})}} \\
+        & = \frac{exp(\theta_*^+ s(y_t^+,y_{t-1})) exp(\theta_*^- s(y_t^-, y_{t-1}))}{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_t^-, y_{t-1}))}
+        \frac{exp(\theta_{m-1}^+ s(y_t^+, y_{t-1}))exp(\theta_{m-1}^- s(y_{ex}^-,y_{t-1}))}{exp(\theta_*^+ s(y_{ex}^+,y_{t-1}))exp(\theta_*^- s(y_{ex}^-,y_{t-1}))} \\
+        & \quad * \frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)} \\
+        & = exp((\theta_*^+ - \theta_{m-1}^+)(s(y_t^+,y_{t-1}) - s(y_{ex}^+,y_{t-1}))(\theta_*^- - \theta_{m-1}^-)(s(y_t^-,y_{t-1}) - s(y_{ex}^-,y_{t-1}))) \\
+        & \quad * \frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}
+    \end{aligned}
+  \end{equation}
 
@@
 
-\[=exp((\theta_*^+ - \theta_{m-1}^+)(s(y_t^+,y_{t-1}) - s(y_{ex}^+,y_{t-1}))(\theta_*^- - \theta_{m-1}^-)(s(y_t^-,y_{t-1}) - s(y_{ex}^-,y_{t-1})))
-\frac{p(\theta_*^+)p(\theta_*^-)}{p(\theta_{m-1}^+)p(\theta_{m-1}^-)}\]
-
 The last expression has no constant term depending on the parameters.
-This finding leads us to the fitting algorithm for the BSTERGMs using the exchange algorithm.
-
+This result leads us to the fitting algorithm for the BSTERGMs using the exchange algorithm.
+
+\clearpage
 \begin{algo}[the main chain]
@@
 
-If $s(.,.)$ and $\theta$ can be multi-dimension, (say, of $\mathbb{R}^p$,)
-replace the multiplication of the exponent parts to dot product between two vectors.
+For step 5, you can use the result of equation (4.1). If $s(.,.)$ and $\theta$ are multi-dimension (in $\mathbb{R}^p, p>1$,) 
+replace the multiplications between $\theta$ and $s$ in the exponent parts to the dot product between two vectors.
 In addition, when implementing the algorithm, I strongly recommend to use $log\pi$ instead of $\pi$ to avoid an underflow problem.
-\[log\pi = (\theta_*^+-\theta_{m-1}^+)(s'(y_{t}^+)-s'(y_{ex}^+))
-+(\theta_*^- -\theta_{m-1}^-)(s'(y_{t}^-)-s'(y_{ex}^-))+log \frac{P(\theta_*^+,\theta_*^-)}{P(\theta_{m-1}^+,\theta_{m-1}^-)}\]
 
@@
 However, since the normalizing constant is still unknown, 
-I cannot get the generative algorithm directly using the model's expression. 
+we cannot get the generative algorithm directly using the model's expression. 
 Thus, I use one more MCMC chain.
@@
     \item Propose a new graph $y_*$ with switching the $y_{ij}$ value from $y_{k-1}$: 
-    \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$(dissolution case.) Oterwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$(formation case.)
+    \\ If $y_{k-1;ij}=1$, then set $y_{*;ij}=0$ (dissolution case.) 
+    \\ Otherwise, If $y_{k-1;ij}=0$, then set $y_{*;ij}=1$ (formation case.)
     \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
     \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
-    \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= \frac{exp(\theta^T s(y_*,y_{t-1}))}{exp(\theta^T s(y_{k-1},y_{t-1}))} = exp(\theta^T (s'(y*)-s'(y_{k-1})))\]
+    \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= exp(\theta^T (s(y_{*},y_{t-1}) - s(y_{k-1},y_{t-1})))\]
     \item With probability $min(\phi,1)$, accept the proposal and put $y_k=y_*$.\\
@@
 
-After $K$ iteration, use the last network as the exchange sample in the second part of the main algorithm.
+After $K$ iteration, use the last sample as the exchange sample in the second part of the main algorithm.
 
@@
 \section{Model diagnostics}
-The ERGMs and all derived models of the ERGMs have suffered the degeneracy problem: 
+The ERGMs and all derived models of the ERGMs suffer the degeneracy problem: 
 all exchange samples' nodes become fully connected or totally isolated as the MCMC chain runs. 
-The problem may cause an explosion to $\infty$ or $-\infty$ of the fitted parameters, 
-failing to converge of MCMC chains. The BSTERGMs are no exception. 
-In other words, the diagnostic process is a very important part of using the BSTERGMs.
+A degenerated chain may cause an explosion of estimated parameter to $\infty$ or $-\infty$,
+failing to converge of MCMC chains. In fact, the BSTERGMs are no exception. 
+Thus, the diagnostic process is a very important part of using the BSTERGMs.
 
@@
 $\{\theta_i^+\}_{i=1}^M, \{\theta_i^-\}_{i=1}^M$ in $\mathbb{R}^p, p\in\mathbb{N}$,
-which are not different with running ordinary MCMC algorithm.
-Depict trace plots of each parameter chain to check the convergence and the mixing.
-If needed, cut samples of the burn-in period and thin the samples.
-Plotting sample autocorrelation for each lag is also essential. Calculate ESS if it is needed.
-
-However, diagnostics of auxiliary chains is challenging. 
+so the diagnostic procedure is not different with an ordinary MCMC chain.
+We can check the convergence and the mixing by depicting trace plots of each parameter.
+We should cut samples of the burn-in period and thin the samples.
+Plotting sample autocorrelation for each lag is also essential. In addition, we can calculate ESS if it is needed.
+
+However, the diagnostic of auxiliary chains is challenging. 
 One reason for the difficulty is that there are too many auxiliary chains in our whole procedure. 
@@
 They are not vectors in $\mathbb{R}^p$ but network structures. 
-Since we cannot draw trace plots of these network sequences, we need other ways to diagnostics. 
-One way is to calculate network statistics $s(.,.)$ in the model and 
+Since we cannot draw trace plots of these network sequences, we need an alternative way for the diagnostic. 
+One possible way is to calculate network statistics $s(.,.)$ in the model and 
 draw trace plots of them instead of the network sequence itself. 
@@
     \item Calculate network statistics of all graphs produced by the last auxiliary chain.
-    \item Depict trace plots of statistics in the model to check the convergence and the mixing.
+    \item Depict trace plots of statistics for the calculated network statistics to check the convergence and the mixing.
 \end{itemize}
@@
 
-Here are some remedy tips when facing a problem.
-If parameters explode to infinity as the number of iteration of the main chain becomes large, 
+Here are some remedy tips when you are facing a convergence problem.
+If parameters explode to the infinity as the number of iteration of the main chain becomes large, 
 check whether the number of iteration of the exchange chain is too small. 
-Picking the exchange sample before an auxiliary chain converges may disturb the main chain. 
+Picking an exchange sample before an auxiliary chain converges may disturb the main chain. 
 Moreover, you have to take account of the combination of network statistics in the model. 
@@
 the parameter may not be identifiable. 
-If you set the model with network statistics with very different sensitivity for adding or erasing an edge but similar MCMC proposal distribution, 
-the MCMC chain tends to fail to converge. 
+If you set the model with several network statistics having very different scale,
+then the fluctuations of each value become differently sensitive when adding or erasing an edge, 
+and the MCMC chain tends to fail to converge if you set similar MCMC proposal distributions over the parameters.
 For example, if you put k-star of specific order and edge count to the model at the same time 
@@
 it is hopeless for the chain to converge appropriately.
-Thus, I recommend running many test chains with low iterations of the main chain 
+Thus, I recommend running many test chains with small iterations of the main chain 
 but lots of iterations of the auxiliary chains 
-to determine the number of the auxiliary chain iterations and good initial points.
+to determine the number of the auxiliary chain iterations and the initial point.
 
@@
 \begin{itemize}
-    \item a outlining shape of the posteriors of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histograms.
+    \item a shape of the posteriors of $\theta^+,\theta^-|y_1,y_2,...,y_T$ by histograms.
     \item an approximated summary statistics: mean, mode, variance and so on.
@@
 
-The interpretation of the parameter is straightforward because of the separated formation and dissolution model.
-Since we cannot know the normalizing constant in the models, but we get odds quickly from the model definition.
-From parameter value, we can expect the direction or tendency of network structure with high probability.
+The interpretation of the fitted parameters is straightforward because of the separated formation and dissolution model in the BSTERGMs.
+Although we cannot know the normalizing constant in the models, we get odds quickly from the model definition and
+expect the direction or tendency of network structures' future transition.
+Let us assume that there exist $\theta^+, \theta^-$ corresponding network statistic $s$ in the model.
 If $\theta^+$ is high, a network with high $s$ has high probability at the formation model, and vice versa.
 Likewise, If $\theta^-$ is low, a network with low $s$ has high probability at the dissolution model, and vice versa.
-Thus, if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $\infty$, the network does not change quickly over time.
+Additionally assume that $s$ becomes higher when the number of edges increases.
+Then if $\theta^+$ go to $-\infty$ and $\theta^-$ go to $\infty$, the network does not change quickly over time.
 If $\theta^+$ go to $\infty$ and $\theta^-$ go to $-\infty$, the network changes drastically over time.
@@
 \begin{itemize}
-    \item Run K iteration using an auxiliary chain at $y_T$ with each posterior sample points.
+    \item Run $K\in \mathbb{N}$ iteration using an auxiliary chain at $y_T$ with $K$ posterior sample points.
     \item Take each network as a predicted result.
@@
 we can use the auxiliary chain algorithm once again.
-\begin{algo}[GOF procedure]
-    For $t=2,...,T$ \\
-    and For $j=1,...,J$
+\begin{algo}[Goodness of Fit checking procedure]
+    For $t=2,...,T$ and for $j=1,...,J$
     \begin{enumerate}
@@
         \item simulate $y_j$ using the auxiliary chain under $y_{t-1}$.
-        \item calculate $s(y_j, y_{t-1})$, some higher degree statistics (eg. Node-degree dist \& Edgewise Shared Partner dist)
+        \item calculate $s(y_j, y_{t-1})$, some higher degree statistics (For example, Node-degree dist \& Edgewise Shared Partner dist)
     \end{enumerate}
-    Take $p$-th and $(1-p)$-th quantiles of each element of $s(y_j, y_{t-1})$s respectively and check 
-    whether the network statistics of observed data is covered by the range of te quantiles or not.
+    Take $p$-th and $(1-p)$-th quantiles of each element of $s(y_j, y_{t-1})$ respectively and check 
+    whether the network statistics of observed data are covered by the range of te quantiles or not.
     In addition, draw the box-plot of calculated $s(y_j, y_{t-1})$ and add points or a line with the network statistics of observed data.
@@
 with UCLA medical working project's team science data.
-These data consist of three sequences, and each sequence has three-time points, 
-and for each time, there is a network with 52 nodes corresponding 29 scholar members and 23 faculty members.
+These data consist of three sequences, and each sequence has three-time points.
+For each time, there is a network with 52 nodes corresponding 29 scholar members and 23 faculty members.
 
@@
 I consider the 2-dimensional BSTERGM as follows.
+For the formation model,
 \[P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) = \frac{exp((\theta^+)^{T}s(y_t^+, y_{t-1}))}{c(\theta^+, y_{t-1})}\]
+and for the dissolution model,
 \[P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-) = \frac{exp((\theta^-)^{T}s(y_t^-, y_{t-1}))}{c(\theta^-, y_{t-1})}\]
-and
-\[P(Y_t=y_t|Y_{t-1}=y_{t-1};\theta^+,\theta^-)=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)\]
+and by assumptions of the BSTERGMs,
+\begin{align*}
+    P(Y_t=y_t|Y_{t-1} &= y_{t-1};\theta^+,\theta^-)\\
+    &=P(Y_t^+=y_t^+|Y_{t-1}=y_{t-1};\theta^+) P(Y_t^-=y_t^-|Y_{t-1}=y_{t-1};\theta^-)
+\end{align*}
 where
@@
 I use the ordinary normal prior $p(\theta^+)\sim\mathcal{N}(0,100)$ and $p(\theta^-)\sim\mathcal{N}(0,100)$,
-and set the proposal distribution of main chain $\epsilon\sim\mathcal{N}(0,0.0004)$.
+and set the proposal distribution of main chain to $\epsilon\sim\mathcal{N}(0,0.0004)$.
 The auxiliary chains consist of 1000 iterations, and the main chain consists of 50000 iterations.
@@
 After generating the chain, I cut the first 10000 samples for a burn-in period and thin the sample by 10.
-The posterior estimates for each network and histograms are as follows.
+The summary estimates and histograms for the posterior distributions of each parameter are as follows.
 
@@
         \begin{tabular}{c | c | c | c }
-            Sequence & Parameter & Estimate & Std.Error \\
+            Sequence & Parameter & Estimated Mean & Estimated Std.Error \\
             \hline \hline
@@
     \end{table}
-\clearpage
+
 %hist
@@
 Finally, I append the results of the goodness of fit check.
-Compare the observed value on the data with simulated values' quantiles generated by the goodness-of-fit algorithm. 
-Parts of sufficient statistics are selected as the network statistics in the goodness of fit procedure.
-Since this example model merely has just two edge terms, 
+By comparing the observed value on the data with simulated values' quantiles generated by the goodness-of-fit algorithm,
+we can check whether the fitting results are proper or not.
+Some parts of sufficient statistics are selected as the network statistics in the goodness of fit procedure.
+Since the example model merely has just two edge terms, 
 and since the BSTERGMs share the same parameter for each network statistics but values of network statistics are different at each time point, 
@@
 
-Since the BSTERGMs have their base on ERGMs, the models are flexible and can include various network statistics. 
+Since the BSTERGMs have their base on the ERGMs, the models are flexible and can include various network statistics. 
 Moreover, the BSTERGMs can be applied to more general networks, including weighted graphs. 
-Moreover, by combining these features, the BSTERGM can take the weights as covariates of the model. 
+Moreover, by combining these features, the BSTERGM can take the weights as covariates of the model
+by a little of modification as the ERGM can.
+
 On the other hand, since the BSTERGMs are using the exchange algorithm, 
@@
 In addition, the calculation of network statistics in the model causes another problem for large networks, 
-especially when the network statistics demand a high complexity of computation. 
+especially when the network statistics demand computation with high complexity.
 For example, calculating the edgewise shared partner distribution of one network 
 needs computation of complexity $O(n^3)$, 
-so computation cost increases rapidly as the number of nodes $n$ grows and 
-the calculating $\pi$ become bottlenecks of the whole computation procedure.
-
-Another critical issue related to the BSTERGMs is degeneracy, as I have mentioned in chapter 5. 
-The BSTERGMs are very sensitive to the main MCMC chain's initial point, 
+so computation cost increases rapidly as the number of nodes $n$ grows, and 
+the calculating $\pi$ may become bottlenecks of the whole computation procedure.
+
+Another critical issue related to the BSTERGMs is degeneracy. 
+As I have mentioned in chapter 5, 
+the convergence of BSTERGMs fitting algorithm is very sensitive to the main MCMC chain's initial point, 
 so searching a region for non-degenerate initial parameter point in the parameter space is very important. 
-Often, however, finding the region is tricky work for the BSTERGMs, partly because of the too many parameters. 
-Since the BSTERGMs have two parts for $\theta^-, \theta^+$ in the setting of the models by definition, 
-the parameter space dimension tends to increase, 
+Often, however, finding the region is tricky work for the BSTERGMs, partly because there are too many parameters. 
+Since the BSTERGMs have two parts - a formation model and a dissolution model - in the setting of the models by definition, 
+the dimension of the parameter space tends to increase 
 although each formation and dissolution model does not have too many terms, respectively. 
@@
 In many cases, the region is geometrically curved 
-due to an implicit relation between several network statistics in the models and interaction between the formation and dissolution models. 
-Thus, the MCMC algorithm may fail to explore the whole parameter space, 
-especially when the parameter space has a high dimension.  
-As a result, the MCMC chain suffers a non-convergence problem. 
+due to an implicit relation between several network statistics in the models and interaction between the formation model and dissolution model. 
+Then, the fitting algorithm's main chain may fail to explore the whole parameter space.
+As a result, the fitting algorithm suffers a non-convergence problem. 
 Imposing some model level constraint or using a special prior may be a remedy, 
@@
 You can find the C++ implementation (using Armadilo: see http://arma.sourceforge.net/)
-of BSTERGM fitting, diagnostic, and GOF algorithms
+of the BSTERGM fitting, diagnostic, and GOF algorithms
 at my Github page: \\
@@
 동적 네트워크의 구조 변화를 묘사하는 데에 유용한 모형이나,
-이 모형에 대한 베이지안 접근은 지금가지 이루어지지 않았었다.
+이 모형에 대한 베이지안 접근은 지금까지 이루어지지 않았었다.
 이 논문에서는, STERGMs에 대한 베이즈 통계학적 접근인 베이지안 분해 가능-시간 지수 랜덤 그래프 모형(BSTERGMs)을 제안하고,

### 2020-12-31 doc/paper_format.tex
doc final
doc final2

@@
 
-The separable Temporal Exponential Random Graph Models (STERGMs)(\cite{RN93}) are useful models to describe changes in dynamic networks' structures evolving over time. 
+The Separable Temporal Exponential Random Graph Models (STERGMs)(\cite{RN93}) are useful models to describe changes in dynamic networks' structures evolving over time. 
 The Bayesian-inference based approach for the models, however, had not been developed yet. 
@@
 \end{figure}
-\clearpage
 

### 2020-12-31 doc/paper_format.tex
mod abstract

@@
 In this paper, I propose the Bayesian Separable Temporal Exponential Random Graph Models (BSTERGMs): STERGMs with the Bayesian setting, 
-and illustrate a computational algorithm that generates posterior samples from the models dealing with the doubly-intractable constant problem. 
+and I illustrate a computational algorithm that generates posterior samples from the models dealing with the doubly-intractable constant problem. 
 Furthermore, I show an inference method of the models in the Bayesian context.

### 2021-02-01 doc/paper_format.tex
doc modicification 0202

@@
 
-\title{BSTERGMs: Bayesian Separable-Temporal Exponential family Graph Models}
+\title{BSTERGMs: Bayesian Separable-Temporal Exponential family Random Graph Models}
 \author{Seokjun Choi}
@@
     \item Select a lag $(t-1,t)$ randomly on $2 \leq t \leq T$.
-    \item Generate an exchange graph $y_{ex} \in\mathcal{Y}|_t$ (with $y_{ex}^+, y_{ex}^-$) at the $\theta_*^+,\theta_*^-$.
+    \item Generate exchange graphs $y_{ex}^+, y_{ex}^-$ in $\mathcal{Y}^+|_t, \mathcal{Y}^-|_t$ at the $\theta_*^+,\theta_*^-$, respectively.
     \item Calculate network statistics $s(y_t^+,y_{t-1}), s(y_{ex}^+,y_{t-1}), s(y_t^-,y_{t-1}), s(y_{ex}^-,y_{t-1})$.
@@
 \begin{algo}[the auxiliary chain]
-Let $\theta^+,\theta^-,y_{t-1}$ be given. For $k=1,...,K$,
+Let $\theta_*$ be given. For $k=1,...,K$,
 \begin{enumerate}
@@
     \\ (If sample graphs are undirected, switch $y_{ji}$ simultaneously.)
-    \item Take $\theta$ as $\theta^+$ or $\theta^-$ according to the case. Calculate the MCMC ratio $\phi$,
+    \item Calculate the MCMC ratio $\phi$,
     \[\phi = \frac{P(Y_t=y_*|y_{t-1},\theta)}{P(Y_t=y_{k-1}|y_{t-1},\theta)}= exp(\theta^T (s(y_{*},y_{t-1}) - s(y_{k-1},y_{t-1})))\]

### 2021-02-01 pyBSTERGM/results/quick_plots.R

@@ -1,11 +1,11 @@
-posterior_chain0 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_0chain.csv", header=FALSE)
-posterior_chain1 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_1chain.csv", header=FALSE)
-posterior_chain2 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_2chain.csv", header=FALSE)
-posterior_chain3 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_3chain.csv", header=FALSE)
-posterior_chain4 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_4chain.csv", header=FALSE)
-posterior_chain5 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_5chain.csv", header=FALSE)
-posterior_chain6 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_6chain.csv", header=FALSE)
-posterior_chain7 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq1_7chain.csv", header=FALSE)
+posterior_chain0 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_0chain.csv", header=FALSE)
+posterior_chain1 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_1chain.csv", header=FALSE)
+posterior_chain2 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_2chain.csv", header=FALSE)
+posterior_chain3 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_3chain.csv", header=FALSE)
+posterior_chain4 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_4chain.csv", header=FALSE)
+posterior_chain5 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_5chain.csv", header=FALSE)
+posterior_chain6 = read.csv("C:/gitProject/BayesianSTERGM/pyBSTERGM/results/seq2_6chain.csv", header=FALSE)
[…]

### 2021-02-01 pyBSTERGM/results/statnet_stergm_fit.R

@@ -1,7 +1,7 @@
 library(statnet)
 
 
-name_vec = c("m2.19", "w2.19", "f2.19")
+name_vec = c("m3.19", "w3.19", "f3.19")
 
 sequence = list()
 for(i in 1:length(name_vec)){

### 2021-02-01 pyBSTERGM/test_gof.r

@@ -1,0 +1,19 @@
+library(tergm)
+
+data(samplk)
+
+# Fit a transition from Time 1 to Time 2
+samplk12 <- stergm(list(samplk1, samplk2, samplk3), #<-여기를 바꿔가며 돌릴 것
+                    formation=~edges+mutual+transitiveties+cyclicalties,
+                    dissolution=~edges+mutual+transitiveties+cyclicalties,
+                    estimate="CMLE")
+
+samplk12.gof <- gof(samplk12)
+
+samplk12.gof
+
+summary(samplk12.gof)
+par(mfrow=c(2,5))
+plot(samplk12.gof)
+
+plot(samplk12.gof, plotlogodds=TRUE)

### 2021-02-01 pyBSTERGM/BSTERGM.py
pyBSTERGM: directed graph support

@@ -4,7 +4,7 @@ import csv
 import numpy as np
 import matplotlib.pyplot as plt
 
-from network import UndirectedNetwork
+from network import UndirectedNetwork, DirectedNetwork
 from network_sampler import NetworkSampler
 
 class BSTERGM:
@@ -16,7 +16,8 @@ class BSTERGM:
         self.MC_formation_samples = []
         self.MC_dissolution_samples = []
         self.node_num = 0
-        
+        self.isDirected = False
+
         self.random_seed = 2021
         self.random_gen = 0
         self.obs_network_formation_seq = []
@@ -39,10 +40,16 @@ class BSTERGM:
 
         self.model = model_fn
         self.node_num = obs_network_seq[0].node_num
+        if isinstance(obs_network_seq[0], DirectedNetwork):
+            self.isDirected = True
+        else:
+            self.isDirected = False
+        
         self.random_seed = rng_seed
         self.random_gen = np.random.default_rng(seed=rng_seed)
         if pid is not None:
             self.pid = pid
+
 
     
     def __str__(self):
@@ -60,7 +67,12 @@ class BSTERGM:
                     y_plus[row,col]=1
                 if y_now_structure[row,col]==0:
                     y_minus[row,col]=0
-        return (UndirectedNetwork(y_plus), UndirectedNetwork(y_minus))
+        result = 0
+        if self.isDirected:
+            result = (DirectedNetwork(y_plus), DirectedNetwork(y_minus))
+        else:
+            result = (UndirectedNetwork(y_plus), UndirectedNetwork(y_minus))
+        return result
 
     def dissociate_obsSeq(self):
         for i in range(1, len(self.obs_network_seq)):
@@ -268,7 +280,7 @@ if __name__=="__main__":
     test_BSTERGM_sampler.run(100, exchange_iter=30)
     # print(test_BSTERGM_sampler.MC_formation_samples)
     # print(test_BSTERGM_sampler.MC_dissolution_samples)
-    # test_BSTERGM_sampler.show_traceplot()
-    # test_BSTERGM_sampler.show_latest_exchangeSampler_netStat_traceplot()
+    test_BSTERGM_sampler.show_traceplot()
+    test_BSTERGM_sampler.show_latest_exchangeSampler_netStat_traceplot()
     # test_BSTERGM_sampler.write_posterior_samples("test")
     # test_BSTERGM_sampler.write_latest_exchangeSampler_netStat("test_netStat")

### 2021-02-01 pyBSTERGM/BSTERGM_GOF.py

@@ -1,7 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-from network import UndirectedNetwork
+from network import UndirectedNetwork, DirectedNetwork
 from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
 
@@ -16,6 +16,11 @@ class BSTERGM_GOF:
         self.random_seed = rng_seed
         self.random_gen = np.random.default_rng(seed=rng_seed)
         self.netstat_list = []
+        self.isDirected = False
+        if isinstance(init_network, DirectedNetwork):
+            self.isDirected = True
+        else:
+            self.isDirected = False
 
     def choose_samples(self):
         num_posterior_sample = len(self.formation_samples)
@@ -38,7 +43,12 @@ class BSTERGM_GOF:
                 if y_last_minus_ydis_structure[row,col]==1:
                     y_now_structure[row,col] = 0
         
-        return UndirectedNetwork(y_now_structure)
+        result_network = 0
+        if self.isDirected:
+            result_network = DirectedNetwork(y_now_structure)
+        else:
+            result_network = UndirectedNetwork(y_now_structure)
+        return result_network
 
 
     def gof_sampler(self, formation_parameter, dissolution_parameter, exchange_iter, rng_seed):
@@ -139,7 +149,7 @@ if __name__ == "__main__":
 
     def gof_netStat(network: UndirectedNetwork):
         gof_netstat = []
-        node_degree_dist = network.statCal_nodeDegreeDist()
+        node_degree_dist = network.statCal_nodeInDegreeDist()
         ESP_dist = network.statCal_EdgewiseSharedPartnerDist()
 
         for val in node_degree_dist:

### 2021-02-01 pyBSTERGM/network.py

@@ -2,8 +2,6 @@ import numpy as np
 
 
 class UndirectedNetwork:
-    
-
     def __init__(self, structure):
         #variables
         self.node_num = 0
@@ -55,6 +53,68 @@ class UndirectedNetwork:
             val += (1 - nested) * ESPdist[k] * np.exp(tau)
         return val
 
+
+
+
+class DirectedNetwork:
+    def __init__(self, structure):
+        #variables
+        self.node_num = 0
+        self.structure = np.array(0)
+        self.stat_nodeDegree = np.array(0)
+
+        #initialize
+        self.structure = structure
+        self.node_num = structure.shape[0]    
+
+    def __str__(self):
+        string_val = "<network.DirectedNetwork object>\n" + self.structure.__str__()
+        return string_val
+
+    def statCal_edgeNum(self):
+        return np.sum(self.structure)
+    def statCal_nodeOutDegree(self):
+        return np.sum(self.structure, axis=1) #rowsum
+    def statCal_nodeInDegree(self):
+        return np.sum(self.structure, axis=0) #rowsum
+    
+    def statCal_nodeOutDegreeDist(self):
+        node_degree_dist = np.zeros(self.node_num)
+        node_degree = self.statCal_nodeOutDegree()
+        for i in node_degree:
+            node_degree_dist[i] += 1
+        return node_degree_dist
+    def statCal_nodeInDegreeDist(self):
+        node_degree_dist = np.zeros(self.node_num)
+        node_degree = self.statCal_nodeInDegree()
+        for i in node_degree:
+            node_degree_dist[i] += 1
+        return node_degree_dist
+    
+    def statCal_EdgewiseSharedPartner(self):
+        ESP = np.zeros((self.node_num, self.node_num))
+        for row_ind in range(self.node_num):
+            for col_ind in range(self.node_num):
+                if self.structure[row_ind, col_ind] == 1:
+                    ESP[row_ind, col_ind] = np.dot(self.structure[row_ind,:].T, self.structure[col_ind,:])
+        return ESP
+    def statCal_EdgewiseSharedPartnerDist(self):
+        fullESP= self.statCal_EdgewiseSharedPartner()
+        ESP_dist = np.zeros(self.node_num-1)
[…]

### 2021-02-01 pyBSTERGM/network_sampler.py

@@ -1,12 +1,12 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-from network import UndirectedNetwork
+from network import UndirectedNetwork, DirectedNetwork
 
 
 
 class NetworkSampler:
-    def __init__(self, model_fn, param, init_network: UndirectedNetwork, rng_seed=2021):
+    def __init__(self, model_fn, param, init_network, rng_seed=2021):
         #variables
         self.initial_network = 0
         self.param = np.array(0)
@@ -28,18 +28,23 @@ class NetworkSampler:
             edge_idx = self.random_gen.integers(low=0, high=self.node_num - 1, size=2)
         return edge_idx
 
-    def propose_network(self, last_network: UndirectedNetwork):
+    def propose_network(self, last_network):
         proposed_structure = last_network.structure.copy()
         edge_idx = self.choose_edge()
-        proposed_structure[edge_idx[0], edge_idx[1]] = 1 - proposed_structure[edge_idx[0], edge_idx[1]]
-        proposed_structure[edge_idx[1], edge_idx[0]] = proposed_structure[edge_idx[0], edge_idx[1]]
-        return UndirectedNetwork(proposed_structure)
+        result_network = 0
+        if isinstance(last_network, UndirectedNetwork):
+            proposed_structure[edge_idx[0], edge_idx[1]] = 1 - proposed_structure[edge_idx[0], edge_idx[1]]
+            proposed_structure[edge_idx[1], edge_idx[0]] = proposed_structure[edge_idx[0], edge_idx[1]]
+            result_network = UndirectedNetwork(proposed_structure)
+        elif isinstance(last_network, DirectedNetwork):
+            proposed_structure[edge_idx[0], edge_idx[1]] = 1 - proposed_structure[edge_idx[0], edge_idx[1]]
+            result_network = DirectedNetwork(proposed_structure)
+        return result_network
 
     def log_r(self, last_network, proposed_network):
         proposed_network_netStat = self.model(proposed_network)
         last_network_netStat = self.model(last_network)
         return np.dot(proposed_network_netStat - last_network_netStat, self.param)
-
 
     def sampler(self):
         last_network = self.network_samples[-1]
@@ -106,6 +111,12 @@ if __name__ == "__main__":
         ]
         )
     test_initnet = UndirectedNetwork(test_structure)
+    test_netSampler = NetworkSampler(model_netStat, np.array([0, 0]), test_initnet)
[…]

### 2021-02-01 pyBSTERGM/testr.r

@@ -1,57 +1,61 @@
 library(statnet)
 
-test_structure = matrix(c(0,1,1,0,0,1,0,1,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,1,0), 5,5)
-test_network = as.network(test_structure, directed=FALSE)
+# test_structure = matrix(c(0,1,1,0,0,1,0,1,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,1,0), 5,5)
+# test_network = as.network(test_structure, directed=FALSE)
+# plot(test_network)
+
+
+test_structure = matrix(c(0,1,1,0,0,1,0,0,0,1,1,1,0,1,0,0,0,0,0,1,1,0,0,1,0),5,5,byrow=T)
+test_network = as.network(test_structure)
 plot(test_network)
 
-
-summary(test_network~gwesp(0.5, fixed=TRUE)) #5.393469
+summary(test_network~edges+desp(0)+desp(1)+desp(2)+desp(3)+dgwesp(0.5, fixed=TRUE)) #5.393469
 fit = ergm(test_network~gwesp)
+?gwesp
 
 
+# elem1 = matrix(c(
+#         0,1,1,0,0,0,0,0,0,0,
+#         1,0,1,1,0,0,0,0,0,0,
+#         1,1,0,1,0,0,0,0,0,0,
+#         0,1,1,0,1,0,0,0,0,0,
+#         0,0,0,1,0,0,0,0,0,0,
+#         0,0,0,0,0,0,0,0,0,0,
+#         0,0,0,0,0,0,0,0,0,0,
+#         0,0,0,0,0,0,0,0,0,0,
+#         0,0,0,0,0,0,0,0,0,0,
+#         0,0,0,0,0,0,0,0,0,0),10,10)
+# net1 = as.network(elem1, directed=FALSE)
+# fit1 = ergm(net1 ~ edges + gwesp(0.5, fixed=TRUE))
+# mcmc.diagnostics(fit1)
 
-elem1 = matrix(c(
-        0,1,1,0,0,0,0,0,0,0,
-        1,0,1,1,0,0,0,0,0,0,
-        1,1,0,1,0,0,0,0,0,0,
-        0,1,1,0,1,0,0,0,0,0,
-        0,0,0,1,0,0,0,0,0,0,
[…]

### 2021-02-01 pyBSTERGM/data_extractFromRPackages.r
pyBSTERGM: add -.py formed data

@@ -1,0 +1,45 @@
+library(statnet)
+data(samplk)
+as.matrix(samplk1)
+as.matrix(samplk2)
+as.matrix(samplk3)
+
+
+write.table(as.matrix(samplk1), row.names=FALSE, col.names=FALSE, file="pyBSTERGM/samplk1.txt", sep=",")
+write.table(as.matrix(samplk2), row.names=FALSE, col.names=FALSE, file="pyBSTERGM/samplk2.txt", sep=",")
+write.table(as.matrix(samplk3), row.names=FALSE, col.names=FALSE, file="pyBSTERGM/samplk3.txt", sep=",")
+
+?write.table
+
+# ==========================================
+
+library(xergm.common)
+data(knecht)
+?knecht
+# step 1: make sure the network matrices have node labels
+for (i in 1:length(friendship)) {
+  rownames(friendship[[i]]) <- 1:nrow(friendship[[i]])
+  colnames(friendship[[i]]) <- 1:ncol(friendship[[i]])
+}
+rownames(primary) <- rownames(friendship[[1]])
+colnames(primary) <- colnames(friendship[[1]])
+sex <- demographics$sex
+names(sex) <- 1:length(sex)
+
+# step 2: imputation of NAs and removal of absent nodes:
+friendship <- handleMissings(friendship, na = 10, method = "remove")
+friendship <- handleMissings(friendship, na = NA, method = "fillmode")
+
+friendship$t1
+friendship$t2
+friendship$t3
+friendship$t4
+
+write.table(friendship$t1, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t1.txt", sep=",")
+write.table(friendship$t2, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t2.txt", sep=",")
+write.table(friendship$t3, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t3.txt", sep=",")
+write.table(friendship$t4, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t4.txt", sep=",")
+
+
+# ==========================================
+library(RSiena)

### 2021-02-01 pyBSTERGM/data_knecht_friendship.py

@@ -1,0 +1,113 @@
+friendship_t1 = [
+    [0,0,1,0,0,1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0],
+    [0,0,0,0,0,0,1,1,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0],
+    [1,1,0,0,0,1,1,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0],
+    [0,0,0,0,0,0,0,1,0,1,0,0,0,0,0,0,0,1,0,0,0,0,1,0,0,0],
+    [0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,1,0,0,0,0,0,0,0,0,1,0],
+    [0,0,1,0,0,0,0,0,0,0,1,1,0,0,1,0,0,0,0,0,1,0,0,0,0,0],
+    [0,1,1,0,0,0,0,0,0,1,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0],
+    [0,0,0,1,0,0,0,0,0,1,0,0,0,0,0,0,0,1,0,0,0,0,1,0,0,0],
+    [1,1,1,0,0,1,1,1,0,0,1,1,0,0,1,0,0,0,0,0,0,0,0,1,0,0],
+    [0,1,0,0,0,0,0,1,0,0,0,0,0,0,1,0,0,0,0,0,0,0,1,0,0,0],
+    [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,1,0,0,0,0,0],
+    [1,1,1,0,0,1,1,1,0,1,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0],
+    [0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,1,0,0,1],
+    [0,0,0,1,1,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,1,0,1,0],
+    [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0],
+    [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0],
[…]

### 2021-02-01 pyBSTERGM/main.py

@@ -3,7 +3,7 @@ from os import getpid
 
 import numpy as np
 
-import Jdata
+import data_Jdata
 from network import UndirectedNetwork
 from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
@@ -36,17 +36,17 @@ def procedure(result_queue, network_sequence, initial_formation_param, initial_d
 
 
 if __name__=="__main__":
-    # sequence1 = [UndirectedNetwork(np.array(Jdata.m1_19_structure)),
-    #     UndirectedNetwork(np.array(Jdata.w1_19_structure)),
-    #     UndirectedNetwork(np.array(Jdata.f1_19_structure))
+    # sequence1 = [UndirectedNetwork(np.array(data_Jdata.m1_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.w1_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.f1_19_structure))
     # ]
-    # sequence2 = [UndirectedNetwork(np.array(Jdata.m2_19_structure)),
-    #     UndirectedNetwork(np.array(Jdata.w2_19_structure)),
-    #     UndirectedNetwork(np.array(Jdata.f2_19_structure))
+    # sequence2 = [UndirectedNetwork(np.array(data_Jdata.m2_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.w2_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.f2_19_structure))
     # ]
-    sequence3 = [UndirectedNetwork(np.array(Jdata.m3_19_structure)),
-        UndirectedNetwork(np.array(Jdata.w3_19_structure)),
-        UndirectedNetwork(np.array(Jdata.f3_19_structure))
+    sequence3 = [UndirectedNetwork(np.array(data_Jdata.m3_19_structure)),
+        UndirectedNetwork(np.array(data_Jdata.w3_19_structure)),
+        UndirectedNetwork(np.array(data_Jdata.f3_19_structure))
     ]
 
 

### 2021-02-01 pyBSTERGM/testr.r

@@ -9,7 +9,7 @@ test_structure = matrix(c(0,1,1,0,0,1,0,0,0,1,1,1,0,1,0,0,0,0,0,1,1,0,0,1,0),5,5
 test_network = as.network(test_structure)
 plot(test_network)
 
-summary(test_network~edges+desp(0)+desp(1)+desp(2)+desp(3)+dgwesp(0.5, fixed=TRUE)) #5.393469
+summary(test_network~edges+odegree(0)+odegree(1)+odegree(2)+odegree(3)+odegree(4)) #5.393469
 fit = ergm(test_network~gwesp)
 ?gwesp
 

### 2021-02-02 pyBSTERGM/BSTERGM.py
example data chains

@@ -30,6 +30,11 @@ class BSTERGM:
         
         #initialize
         self.obs_network_seq = obs_network_seq
+        if isinstance(obs_network_seq[0], DirectedNetwork):
+            self.isDirected = True
+        else:
+            self.isDirected = False
+        
         self.initial_formation_param = initial_formation_param
         self.initial_dissolution_param = initial_dissolution_param
         self.MC_formation_samples.append(initial_formation_param)
@@ -37,13 +42,8 @@ class BSTERGM:
         self.obs_network_formation_seq.append(initial_formation_param)
         self.obs_network_dissolution_seq.append(initial_dissolution_param)
         self.dissociate_obsSeq()
-
         self.model = model_fn
         self.node_num = obs_network_seq[0].node_num
-        if isinstance(obs_network_seq[0], DirectedNetwork):
-            self.isDirected = True
-        else:
-            self.isDirected = False
         
         self.random_seed = rng_seed
         self.random_gen = np.random.default_rng(seed=rng_seed)

### 2021-02-02 pyBSTERGM/main.py

@@ -3,8 +3,8 @@ from os import getpid
 
 import numpy as np
 
-import data_Jdata
-from network import UndirectedNetwork
+import data_samplk, data_knecht_friendship
+from network import UndirectedNetwork, DirectedNetwork
 from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
 
@@ -25,7 +25,7 @@ def procedure(result_queue, network_sequence, initial_formation_param, initial_d
     # print(BSTERGM_sampler.MC_formation_samples)
     # print(BSTERGM_sampler.MC_dissolution_samples)
     BSTERGM_sampler.write_posterior_samples(result_string)
-    BSTERGM_sampler.write_latest_exchangeSampler_netStat(result_string + "_statNet")
+    BSTERGM_sampler.write_latest_exchangeSampler_netStat(result_string + "_NetworkStat")
 
     result_queue.put(BSTERGM_sampler)
 
@@ -36,6 +36,8 @@ def procedure(result_queue, network_sequence, initial_formation_param, initial_d
 
 
 if __name__=="__main__":
+    #Jdata
+
     # sequence1 = [UndirectedNetwork(np.array(data_Jdata.m1_19_structure)),
     #     UndirectedNetwork(np.array(data_Jdata.w1_19_structure)),
     #     UndirectedNetwork(np.array(data_Jdata.f1_19_structure))
@@ -44,9 +46,25 @@ if __name__=="__main__":
     #     UndirectedNetwork(np.array(data_Jdata.w2_19_structure)),
     #     UndirectedNetwork(np.array(data_Jdata.f2_19_structure))
     # ]
-    sequence3 = [UndirectedNetwork(np.array(data_Jdata.m3_19_structure)),
-        UndirectedNetwork(np.array(data_Jdata.w3_19_structure)),
-        UndirectedNetwork(np.array(data_Jdata.f3_19_structure))
+    # sequence3 = [UndirectedNetwork(np.array(data_Jdata.m3_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.w3_19_structure)),
+    #     UndirectedNetwork(np.array(data_Jdata.f3_19_structure))
+    # ]
+
+    # #samplk
+    # samplk_sequence = [
+    #     DirectedNetwork(np.array(data_samplk.samplk1)),
+    #     DirectedNetwork(np.array(data_samplk.samplk2)),
+    #     DirectedNetwork(np.array(data_samplk.samplk3))
+    # ]
+
+
+    # knecht_friendship
[…]

### 2021-02-02 pyBSTERGM/BSTERGM_GOF.py
pyBSTERGM ;example diagonostics and gof procedure

@@ -91,6 +91,9 @@ class BSTERGM_GOF:
 
 
 if __name__ == "__main__":
+
+
+    
     test_structure1 = np.array(
     [
         [0,1,1,0,0,0,0,0,0,0],
@@ -140,14 +143,14 @@ if __name__ == "__main__":
     test_obs_seq = [test_initnet1, test_initnet2, test_initnet3]
 
 
-    def model_netStat(network : UndirectedNetwork):
+    def model_netStat(network):
         model = []
         #define model
         model.append(network.statCal_edgeNum())
         model.append(network.statCal_geoWeightedESP(0.5))
         return np.array(model)
 
-    def gof_netStat(network: UndirectedNetwork):
+    def gof_netStat(network):
         gof_netstat = []
         node_degree_dist = network.statCal_nodeInDegreeDist()
         ESP_dist = network.statCal_EdgewiseSharedPartnerDist()

### 2021-02-02 pyBSTERGM/BSTERGM_diagnosis.py

@@ -162,13 +162,16 @@ class BSTERGM_latest_exchangeSampler_work:
 
 if __name__ == "__main__":
     reader_inst = BSTERGM_posterior_work()
-    reader_inst.read_from_csv("results/seq1_0chain", 2, 2)
+    reader_inst.read_from_csv("friendship_results/friendship_1chain", 2, 2)
     # print(reader_inst.MC_formation_samples[0:10])
+    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[10000::10]
+    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[10000::10]
+    print(np.mean(reader_inst.MC_sample_trace()[0][0]), np.mean(reader_inst.MC_sample_trace()[0][1]),
+        np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]))
     reader_inst.show_traceplot()
     reader_inst.show_histogram()
     reader_inst.show_acfplot()
 
-
     netstat_reader_inst = BSTERGM_latest_exchangeSampler_work()
-    netstat_reader_inst.read_from_csv("results/seq1_0chain_statNet")
+    netstat_reader_inst.read_from_csv("friendship_results/friendship_1chain_NetworkStat")
     netstat_reader_inst.show_traceplot()

### 2021-02-02 pyBSTERGM/data_extractFromRPackages.r

@@ -30,10 +30,10 @@ names(sex) <- 1:length(sex)
 friendship <- handleMissings(friendship, na = 10, method = "remove")
 friendship <- handleMissings(friendship, na = NA, method = "fillmode")
 
-friendship$t1
-friendship$t2
-friendship$t3
-friendship$t4
+length(friendship$t1)
+length(friendship$t2)
+length(friendship$t3)
+length(friendship$t4)
 
 write.table(friendship$t1, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t1.txt", sep=",")
 write.table(friendship$t2, row.names=FALSE, col.names=FALSE, file="pyBSTERGM/knecht_friendship_t2.txt", sep=",")

### 2021-02-02 pyBSTERGM/main_gof.py

@@ -1,0 +1,88 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+from network import UndirectedNetwork, DirectedNetwork
+from network_sampler import NetworkSampler
+from BSTERGM import BSTERGM
+from BSTERGM_diagnosis import BSTERGM_posterior_work
+from BSTERGM_GOF import BSTERGM_GOF
+
+import data_samplk
+import data_knecht_friendship
+
+if __name__ == "__main__":
+    samplk_sequence = [
+        DirectedNetwork(np.array(data_samplk.samplk1)),
+        DirectedNetwork(np.array(data_samplk.samplk2)),
+        DirectedNetwork(np.array(data_samplk.samplk3))
+    ]
+
+    friendship_sequence = [
+        DirectedNetwork(np.array(data_knecht_friendship.friendship_t1)), #node=26
+        DirectedNetwork(np.array(data_knecht_friendship.friendship_t2)), #26
+        DirectedNetwork(np.array(data_knecht_friendship.friendship_t3)), #node=25 (when R controls NAs...)
+        DirectedNetwork(np.array(data_knecht_friendship.friendship_t4)) #25 #<-1/2에서 빠진줄을 빼면 될 듯
+    ]
+
+    reader_inst = BSTERGM_posterior_work()
+    reader_inst.read_from_csv("samplk_results/samplk_1chain", 2, 2)
+    # print(reader_inst.MC_formation_samples[0:10])
+    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[10000::10]
+    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[10000::10]
+    # print(np.mean(reader_inst.MC_sample_trace()[0][0]), np.mean(reader_inst.MC_sample_trace()[0][1]),
+    #     np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]))
+    # reader_inst.show_traceplot()
+    # reader_inst.show_histogram()
+    # reader_inst.show_acfplot()
+
+
+    
+    def model_netStat(network):
+        model = []
+        #define model
+        model.append(network.statCal_edgeNum())
+        model.append(network.statCal_geoWeightedESP(0.5))
+        return np.array(model)
+
+    def gof_netStat(network):
+        gof_netstat = []
[…]

### 2021-02-02 pyBSTERGM/network.py
pyBSTERGM: new network statistics

@@ -53,6 +53,45 @@ class UndirectedNetwork:
             val += (1 - nested) * ESPdist[k] * np.exp(tau)
         return val
 
+    def statCal_MinGeodesic(self):
+        # Dijkstra algorithm
+        def dijkstra(self,source):
+            from math import inf
+            visited_node = [False for _ in range(self.node_num)]
+            dist_vec = [inf for i in range(self.node_num)]
+            prev = [None for i in range(self.node_num)]
+            
+            dist_vec[source]=0
+
+            while not all(visited_node):
+                unvisited = [(i, dist) for i, dist in enumerate(dist_vec) if visited_node[i]==False]
+                chosen_node = min(unvisited, key=lambda tup:tup[1])[0]
+
+                visited_node[chosen_node] = True
+                neighbors = [i for i, val in enumerate(self.structure[chosen_node,:]) if val==1]
+                for nb_node in neighbors:
+                    alt = dist_vec[chosen_node] + 1 #length = 1
+                    if alt < dist_vec[nb_node]:
+                        dist_vec[nb_node] = alt
+                        prev[nb_node] = chosen_node
+
+            return (dist_vec, prev)
+        min_geodesic_vec = []
+        for i in range(self.node_num):
+            dist, _ = dijkstra(self, i)
+            min_geodesic_vec.append(dist)
+        return min_geodesic_vec
+    
+    def statCal_MinGeodesicDist(self):
+        min_geodesic = self.statCal_MinGeodesic()
+        distrib_vec = [0 for _ in range(self.node_num)]
+        for from_source_node in min_geodesic:
+            for val in from_source_node:
+                distrib_vec[val] += 1
+        distrib_vec[0] = 0
+        undirected_distrib_vec = [val/2 for val in distrib_vec]
+        return undirected_distrib_vec[1:] #dist0 = node itself
+
 
 
 
@@ -113,6 +152,103 @@ class DirectedNetwork:
             nested = (-np.expm1(-tau))**k
             val += (1 - nested) * ESPdist[k] * np.exp(tau)
         return val
+    
+    
+    def statCal_MinGeodesic(self):
+        # Dijkstra algorithm
+        def dijkstra(self,source):
+            from math import inf
+            visited_node = [False for _ in range(self.node_num)]
+            dist_vec = [inf for i in range(self.node_num)]
+            prev = [None for i in range(self.node_num)]
+            
+            dist_vec[source]=0
+
[…]

### 2021-02-02 pyBSTERGM/testr.r

@@ -10,8 +10,12 @@ test_network = as.network(test_structure)
 plot(test_network)
 
 summary(test_network~edges+odegree(0)+odegree(1)+odegree(2)+odegree(3)+odegree(4)) #5.393469
+summary(test_network~edges+mutual+transitiveties++cyclicalties)
+
 fit = ergm(test_network~gwesp)
 ?gwesp
+
+
 
 
 # elem1 = matrix(c(

### 2021-02-02 pyBSTERGM/BSTERGM_GOF.py
pyBSTERGM: gof - separate formation/dissolution

@@ -5,82 +5,234 @@ from network import UndirectedNetwork, DirectedNetwork
 from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
 
+def dissociate_network(last_network, now_network):
+    y_last_structure = last_network.structure
+    y_now_structure = now_network.structure
+    y_plus = y_last_structure.copy()
+    y_minus = y_last_structure.copy()
+    for row in range(last_network.node_num):
+        for col in range(last_network.node_num):
+            if y_now_structure[row,col]==1:
+                y_plus[row,col]=1
+            if y_now_structure[row,col]==0:
+                y_minus[row,col]=0
+    result = 0
+    if isinstance(last_network,DirectedNetwork):
+        result = (DirectedNetwork(y_plus), DirectedNetwork(y_minus))
+    else:
+        result = (UndirectedNetwork(y_plus), UndirectedNetwork(y_minus))
+    return result
+
 class BSTERGM_GOF:
-    def __init__(self, model_function, GOFplot_netstat_function, init_network, posterior_formation, posterior_dissolution, rng_seed=2021):
+    def __init__(self, model_function, posterior_parameter_samples, init_network, additional_netstat_function=None , rng_seed=2021):
         self.model = model_function
-        self.netstats = GOFplot_netstat_function
-        self.formation_samples = posterior_formation
-        self.dissolution_samples = posterior_dissolution
+        self.additional_netstats = additional_netstat_function
+        self.posterior_parameter_samples = posterior_parameter_samples
         self.init_network = init_network
 
         self.random_seed = rng_seed
         self.random_gen = np.random.default_rng(seed=rng_seed)
-        self.netstat_list = []
+        self.additional_netstat_list = []
         self.isDirected = False
         if isinstance(init_network, DirectedNetwork):
             self.isDirected = True
+            self.sample_nodeInDegreeDist = []
+            self.sample_nodeOutDegreeDist = []
+            self.sample_ESPDist = []
+            self.sample_minGeodesicDist = []
+            self.sample_modelStats = []
+
         else:
             self.isDirected = False
+            self.sample_nodeDegreeDist = []
+            self.sample_ESPDist = []
+            self.sample_minGeodesicDist = []
+
 
     def choose_samples(self):
-        num_posterior_sample = len(self.formation_samples)
[…]

### 2021-02-02 pyBSTERGM/main_gof.py

@@ -4,8 +4,8 @@ import matplotlib.pyplot as plt
 from network import UndirectedNetwork, DirectedNetwork
 from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
-from BSTERGM_diagnosis import BSTERGM_posterior_work
-from BSTERGM_GOF import BSTERGM_GOF
+from BSTERGM_diagnosis import BSTERGM_posterior_work, BSTERGM_latest_exchangeSampler_work
+from BSTERGM_GOF import BSTERGM_GOF, dissociate_network
 
 import data_samplk
 import data_knecht_friendship
@@ -25,17 +25,20 @@ if __name__ == "__main__":
     ]
 
     reader_inst = BSTERGM_posterior_work()
-    reader_inst.read_from_csv("samplk_results/samplk_1chain", 2, 2)
+    reader_inst.read_from_csv("samplk_results/samplk_6chain", 2, 2)
+
     # print(reader_inst.MC_formation_samples[0:10])
     reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[10000::10]
     reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[10000::10]
     # print(np.mean(reader_inst.MC_sample_trace()[0][0]), np.mean(reader_inst.MC_sample_trace()[0][1]),
     #     np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]))
-    # reader_inst.show_traceplot()
-    # reader_inst.show_histogram()
-    # reader_inst.show_acfplot()
-
-
+    reader_inst.show_traceplot()
+    reader_inst.show_histogram()
+    reader_inst.show_acfplot()
+    
+    netstat_reader_inst = BSTERGM_latest_exchangeSampler_work()
+    netstat_reader_inst.read_from_csv("samplk_results/samplk_6chain_NetworkStat")
+    netstat_reader_inst.show_traceplot()
     
     def model_netStat(network):
         model = []
@@ -44,45 +47,34 @@ if __name__ == "__main__":
         model.append(network.statCal_geoWeightedESP(0.5))
         return np.array(model)
 
-    def gof_netStat(network):
+    def gof_additional_netStat(network):
         gof_netstat = []
-        node_InDegree_dist = network.statCal_nodeInDegreeDist()
[…]

### 2021-02-02 pyBSTERGM/network.py

@@ -83,11 +83,15 @@ class UndirectedNetwork:
         return min_geodesic_vec
     
     def statCal_MinGeodesicDist(self):
+        from math import inf
         min_geodesic = self.statCal_MinGeodesic()
-        distrib_vec = [0 for _ in range(self.node_num)]
+        distrib_vec = [0 for _ in range(self.node_num+1)]
         for from_source_node in min_geodesic:
             for val in from_source_node:
-                distrib_vec[val] += 1
+                if val is inf:
+                    distrib_vec[-1] += 1
+                else:
+                    distrib_vec[val] += 1
         distrib_vec[0] = 0
         undirected_distrib_vec = [val/2 for val in distrib_vec]
         return undirected_distrib_vec[1:] #dist0 = node itself
@@ -184,11 +188,16 @@ class DirectedNetwork:
         return min_geodesic_vec
     
     def statCal_MinGeodesicDist(self):
+        from math import inf
         min_geodesic = self.statCal_MinGeodesic()
-        distrib_vec = [0 for _ in range(self.node_num)]
+        distrib_vec = [0 for _ in range(self.node_num + 1)]
+        # print(min_geodesic)
         for from_source_node in min_geodesic:
             for val in from_source_node:
-                distrib_vec[val] += 1
+                if val is inf:
+                    distrib_vec[-1] += 1
+                else:
+                    distrib_vec[val] += 1
         distrib_vec[0] = 0
         return distrib_vec[1:] #dist0 = node itself
     

### 2021-02-02 pyBSTERGM/main.py
pyBSTERGM: netstat add - ctriple, ttriple

@@ -51,21 +51,21 @@ if __name__=="__main__":
     #     UndirectedNetwork(np.array(data_Jdata.f3_19_structure))
     # ]
 
-    # #samplk
-    # samplk_sequence = [
-    #     DirectedNetwork(np.array(data_samplk.samplk1)),
-    #     DirectedNetwork(np.array(data_samplk.samplk2)),
-    #     DirectedNetwork(np.array(data_samplk.samplk3))
-    # ]
-
-
-    # knecht_friendship
-    friendship_sequence = [
-        DirectedNetwork(np.array(data_knecht_friendship.friendship_t1)),
-        DirectedNetwork(np.array(data_knecht_friendship.friendship_t2)),
-        DirectedNetwork(np.array(data_knecht_friendship.friendship_t3)),
-        DirectedNetwork(np.array(data_knecht_friendship.friendship_t4))
+    #samplk
+    samplk_sequence = [
+        DirectedNetwork(np.array(data_samplk.samplk1)),
+        DirectedNetwork(np.array(data_samplk.samplk2)),
+        DirectedNetwork(np.array(data_samplk.samplk3))
     ]
+
+
+    # # knecht_friendship
+    # friendship_sequence = [
+    #     DirectedNetwork(np.array(data_knecht_friendship.friendship_t1)),
+    #     DirectedNetwork(np.array(data_knecht_friendship.friendship_t2)),
+    #     DirectedNetwork(np.array(data_knecht_friendship.friendship_t3)),
+    #     DirectedNetwork(np.array(data_knecht_friendship.friendship_t4))
+    # ]
 
 
     #core
@@ -96,7 +96,7 @@ if __name__=="__main__":
 
     for i in range(core_num):
         process_unit = mp.Process(target=procedure, 
-        args=(proc_queue, friendship_sequence, initial_formation_vec[i], initial_dissolution_vec[i], "friendship_"+str(i)+"chain", 2021+i*10, 30000, 50))
+        args=(proc_queue, samplk_sequence, initial_formation_vec[i], initial_dissolution_vec[i], "samplk_sequence_iter20000_ex100_"+str(i)+"chain", 2021+i*10, 20000, 100))
         # def procedure(result_queue, network_sequence, initial_formation_param, initial_dissolution_param, result_string, rng_seed=2021, main_iter=30000, ex_iter=50)
         
         process_vec.append(process_unit)

### 2021-02-02 pyBSTERGM/network.py

@@ -220,25 +220,42 @@ class DirectedNetwork:
                     # print(start_node, end_node, idx_set)
                     for inter_node in idx_set:
                         if self.structure[start_node, inter_node]==1 and self.structure[inter_node,end_node]==1:
-                            existance[start_node, end_node] = 1
-                            break
+                            existance[start_node, end_node] += 1
         return existance
 
+    def statCal_transitiveTriples(self):
+        result = 0
+        twoPath = self.statCal_existTwoPath()
+        for start_node in range(self.node_num):
+            for end_node in range(self.node_num):
+                if self.structure[start_node, end_node]==1 and twoPath[start_node, end_node]>0:
+                    result += twoPath[start_node, end_node]
+        return result/3
+    
     def statCal_transitiveTies(self):
         result = 0
         twoPath = self.statCal_existTwoPath()
         for start_node in range(self.node_num):
             for end_node in range(self.node_num):
-                if self.structure[start_node, end_node]==1 and twoPath[start_node, end_node]==1:
+                if self.structure[start_node, end_node]==1 and twoPath[start_node, end_node]>0:
                     result += 1
         return result
+
+    def statCal_cyclicTriples(self):
+        result = 0
+        twoPath = self.statCal_existTwoPath()
+        for start_node in range(self.node_num):
+            for end_node in range(self.node_num):
+                if twoPath[start_node, end_node]>1 and self.structure[end_node, start_node]==1 :
+                    result += twoPath[start_node, end_node]
+        return result/3
     
     def statCal_cyclicalTies(self):
         result = 0
         twoPath = self.statCal_existTwoPath()
         for start_node in range(self.node_num):
             for end_node in range(self.node_num):
-                if twoPath[start_node, end_node]==1 and self.structure[end_node, start_node]==1 :
+                if twoPath[start_node, end_node]>1 and self.structure[end_node, start_node]==1 :
                     result += 1
         return result
 
@@ -300,3 +317,6 @@ if __name__ == "__main__":
     print(test_net_2.statCal_existTwoPath())
     print(test_net_2.statCal_transitiveTies())#true 4
     print(test_net_2.statCal_cyclicalTies()) #true 6
+
+    print(test_net_2.statCal_transitiveTriples())#true 4
+    print(test_net_2.statCal_cyclicTriples())#true 2

### 2021-02-02 pyBSTERGM/testr.r

@@ -11,6 +11,7 @@ plot(test_network)
 
 summary(test_network~edges+odegree(0)+odegree(1)+odegree(2)+odegree(3)+odegree(4)) #5.393469
 summary(test_network~edges+mutual+transitiveties++cyclicalties)
+summary(test_network~edges+ctriple+ttriple)
 
 fit = ergm(test_network~gwesp)
 ?gwesp

### 2021-02-02 pyBSTERGM/network.py
ctriple ttriple modify

@@ -230,7 +230,7 @@ class DirectedNetwork:
             for end_node in range(self.node_num):
                 if self.structure[start_node, end_node]==1 and twoPath[start_node, end_node]>0:
                     result += twoPath[start_node, end_node]
-        return result/3
+        return result
     
     def statCal_transitiveTies(self):
         result = 0
@@ -246,7 +246,7 @@ class DirectedNetwork:
         twoPath = self.statCal_existTwoPath()
         for start_node in range(self.node_num):
             for end_node in range(self.node_num):
-                if twoPath[start_node, end_node]>1 and self.structure[end_node, start_node]==1 :
+                if twoPath[start_node, end_node]>0 and self.structure[end_node, start_node]==1 :
                     result += twoPath[start_node, end_node]
         return result/3
     
@@ -255,7 +255,7 @@ class DirectedNetwork:
         twoPath = self.statCal_existTwoPath()
         for start_node in range(self.node_num):
             for end_node in range(self.node_num):
-                if twoPath[start_node, end_node]>1 and self.structure[end_node, start_node]==1 :
+                if twoPath[start_node, end_node]>0 and self.structure[end_node, start_node]==1 :
                     result += 1
         return result
 

### 2021-02-03 pyBSTERGM/BSTERGM_diagnosis.py
samplk new model

@@ -162,16 +162,20 @@ class BSTERGM_latest_exchangeSampler_work:
 
 if __name__ == "__main__":
     reader_inst = BSTERGM_posterior_work()
-    reader_inst.read_from_csv("friendship_results/friendship_1chain", 2, 2)
+    reader_inst.read_from_csv("samplk_results/samplk_sequence_ex_model_5chain", 4, 4)
     # print(reader_inst.MC_formation_samples[0:10])
-    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[10000::10]
-    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[10000::10]
+    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[2000::10]
+    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[2000::10]
     print(np.mean(reader_inst.MC_sample_trace()[0][0]), np.mean(reader_inst.MC_sample_trace()[0][1]),
-        np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]))
+        np.mean(reader_inst.MC_sample_trace()[0][2]), np.mean(reader_inst.MC_sample_trace()[0][3]))
+    #STERGM formation: -3.5586, 2.2624, -0.4994, 0.2945
+    print(np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]),
+        np.mean(reader_inst.MC_sample_trace()[1][2]), np.mean(reader_inst.MC_sample_trace()[1][3]))
+    #STERGM dissolution: -0.1164, 1.5791, -1.6957, 0.6847
     reader_inst.show_traceplot()
     reader_inst.show_histogram()
     reader_inst.show_acfplot()
 
     netstat_reader_inst = BSTERGM_latest_exchangeSampler_work()
-    netstat_reader_inst.read_from_csv("friendship_results/friendship_1chain_NetworkStat")
+    netstat_reader_inst.read_from_csv("samplk_results/samplk_sequence_ex_model_5chain_NetworkStat")
     netstat_reader_inst.show_traceplot()

### 2021-02-03 pyBSTERGM/data_extractFromRPackages.r

@@ -1,5 +1,6 @@
 library(statnet)
 data(samplk)
+?samplk
 as.matrix(samplk1)
 as.matrix(samplk2)
 as.matrix(samplk3)

### 2021-02-03 pyBSTERGM/main.py

@@ -9,11 +9,13 @@ from network_sampler import NetworkSampler
 from BSTERGM import BSTERGM
 
 
-def model_netStat(network : UndirectedNetwork):
+def model_netStat(network):
     model = []
     #define model
     model.append(network.statCal_edgeNum())
-    model.append(network.statCal_geoWeightedESP(0.5))
+    model.append(network.statCal_mutuality())
+    model.append(network.statCal_cyclicTriples())
+    model.append(network.statCal_transitiveTriples())
     return np.array(model)
 
 def procedure(result_queue, network_sequence, initial_formation_param, initial_dissolution_param, result_string, rng_seed=2021, main_iter=30000, ex_iter=50):
@@ -74,29 +76,29 @@ if __name__=="__main__":
     proc_queue = mp.Queue()
 
     initial_formation_vec = [
-        np.array([0,0]), 
-        np.array([-0.05, -0.05]), 
-        np.array([0.05, 0.05]), 
-        np.array([-0.1, -0.1]), 
-        np.array([0.1, 0.1]), 
-        np.array([-0.2, -0.2]), 
-        np.array([0.2, 0.2]), 
-        np.array([-0.1, 0.1])
+        np.array([0, 0, 0, 0]), 
+        np.array([-0.05, -0.05, -0.05, -0.05]), 
+        np.array([0.05, 0.05, 0.05, 0.05]), 
+        np.array([-0.1, -0.1, -0.1, -0.1]), 
+        np.array([0.1, 0.1, 0.1, 0.1]), 
+        np.array([-0.2, -0.2, -0.2, -0.2]), 
+        np.array([0.2, 0.2, 0.2, 0.2]), 
+        np.array([-0.1, 0.1, -0.1, 0.1])
     ]
     initial_dissolution_vec = [
-        np.array([0,0]), 
-        np.array([-0.05, -0.05]), 
-        np.array([0.05, 0.05]), 
-        np.array([-0.1, -0.1]), 
-        np.array([0.1, 0.1]), 
-        np.array([-0.2, -0.2]), 
-        np.array([0.2, 0.2]), 
-        np.array([-0.1, 0.1])
+        np.array([0, 0, 0, 0]), 
+        np.array([-0.05, -0.05, -0.05, -0.05]), 
+        np.array([0.05, 0.05, 0.05, 0.05]), 
+        np.array([-0.1, -0.1, -0.1, -0.1]), 
+        np.array([0.1, 0.1, 0.1, 0.1]), 
[…]

### 2021-02-03 pyBSTERGM/main_gof.py

@@ -25,11 +25,11 @@ if __name__ == "__main__":
     ]
 
     reader_inst = BSTERGM_posterior_work()
-    reader_inst.read_from_csv("samplk_results/samplk_6chain", 2, 2)
+    reader_inst.read_from_csv("samplk_results/samplk_sequence_ex_model_5chain", 4, 4)
 
     # print(reader_inst.MC_formation_samples[0:10])
-    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[10000::10]
-    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[10000::10]
+    reader_inst.MC_dissolution_samples = reader_inst.MC_dissolution_samples[2000::5]
+    reader_inst.MC_formation_samples = reader_inst.MC_formation_samples[2000::5]
     # print(np.mean(reader_inst.MC_sample_trace()[0][0]), np.mean(reader_inst.MC_sample_trace()[0][1]),
     #     np.mean(reader_inst.MC_sample_trace()[1][0]), np.mean(reader_inst.MC_sample_trace()[1][1]))
     reader_inst.show_traceplot()
@@ -37,14 +37,16 @@ if __name__ == "__main__":
     reader_inst.show_acfplot()
     
     netstat_reader_inst = BSTERGM_latest_exchangeSampler_work()
-    netstat_reader_inst.read_from_csv("samplk_results/samplk_6chain_NetworkStat")
+    netstat_reader_inst.read_from_csv("samplk_results/samplk_sequence_ex_model_5chain_NetworkStat")
     netstat_reader_inst.show_traceplot()
     
     def model_netStat(network):
         model = []
         #define model
         model.append(network.statCal_edgeNum())
-        model.append(network.statCal_geoWeightedESP(0.5))
+        model.append(network.statCal_mutuality())
+        model.append(network.statCal_cyclicTriples())
+        model.append(network.statCal_transitiveTriples())
         return np.array(model)
 
     def gof_additional_netStat(network):
@@ -60,21 +62,21 @@ if __name__ == "__main__":
     net_plus1, net_minus1 = dissociate_network(samplk_sequence[0], samplk_sequence[1])
 
[…]
